---
output: 
  word_document: default
  pdf_document: default
  html_document: default
---

Context section???


## SCIENCE ADVICE {-}
[Mandatory. Maximum length 500 words. Include summary bullets of the following factors affecting the stock that are relevant for decision-making. Please refer to the Guidance Document for standardized language and examples, and what to avoid in this section.] 

### Status {-}

::: {custom-style="fsar-bullet"}
[Mandatory bullet(s)]

The three-year mean exploitable biomass index from the Stratified Random Halibut Survey for 2024 is 51.9 kt which is 1.9B~MSY~. The probability of the stock being in the healthy zone is very high.
:::

### Trends {-}

::: {custom-style="fsar-bullet"}
[Mandatory bullet(s)] 

The 2024 index of exploitable biomass from the Stratified Random Survey and the three-year mean are the highest in the time series. Inter-annual changes in the three-year mean have remained relatively stable through the time series with a small increase in 2024 from the previous year.

Atlantic halibut catch per tow in 2024 from the DFO Maritimes Region Summer Ecosystem Survey (RV Survey) was the second highest value in the time series. Catch from this survey, which provides an index of recruitment for Atlantic halibut, increased between 2000 and 2011, followed by a decline, but has been increasing again since 2020.

:::

### Ecosystem and Climate Change Considerations {-}

::: {custom-style="fsar-bullet"}
[Mandatory bullet(s)]

•	Ecosystem and climate change considerations were taken into account in the most recent stock assessment and were not further updated or reviewed during this interim-year update.

:::

### Stock Advice {-}

::: {custom-style="fsar-bullet"}
[Mandatory bullet(s)]

The three-year mean exploitable biomass of 51.9 kt corresponds to removals of 5,149 t under the Harvest Decision Rule for this stock. - this is word for word from the guidance doc

Removals of 5,149 t is within the maximum annual change of 15% from the 2024-2025 TAC. It is 222 t (4.5%) more than last year's TAC.


SOME STATEMENT ABOUT LIKELIHOOD OF REMAININGS IN THE HEALTHY ZONE?? 

:::

### Other Management Questions [if applicable, otherwise do not include] {-}

::: {custom-style="fsar-bullet"}
[Other bullets if applicable]
:::

## BASIS FOR ASSESSMENT {-}

### Assessment Details {-}
[Mandatory]

#### Year Assessment Approach was Approved {-}
[2022 (Johnson et al., 2024; DFO 2024)]

#### Assessment Type {-}
[Mandatory; select one of the following: Full Assessment, Interim Year Update, or Harvest Decision Rule (HDR) Update. See the Guidance Document for definitions.] 

Interim Year Update

#### Most Recent Assessment Date {-}
1.	Last Full Assessment: [March 1-4, 2022 (Johnson et al., 2024; DFO 2024a)]
2.	Last Interim Year Update: [December 6, 2023 (DFO 2024b)]

#### Assessment Approach {-}
1.	Broad category: [Select one of the following: Data-poor (e.g., catch-MSY, very limited data, etc.), index-based (trends in empirical indices only), single stock assessment model, model ensemble, MSE (full or MSE-lite), Wild Salmon Policy Integrated Assessment, multiple approaches, other (specify)] 

- index-based (trends in empirical indices only), single stock assessment model, MSE-lite

2.	Specific category: [Select one or as many as apply: No data, Index-based (fishery-dependent indices only), Index-based (including fishery-dependent and fishery-independent indices), Delay difference, Biomass dynamics, Virtual population analysis, Statistical catch-at-length, Statistical catch-at-age, State-Space, Sequential Population Analysis, Surplus Production, Stock-Recruitment Relationship, Potential Biological Removal, Other (please explain)]

- Statistical catch-at-length; index-based (including fishery-dependent and fishery-independent indices)

[Optional short summary of stock assessment methodology. Include the method of considering ecosystem variables. Include a reference to a Research Document or other publication, and/or footnote to the working paper. Maximum length 100 words.]

A new assessment framework was adopted in 2022. This framework used a catch-at-length model with sex and age structure to assess the stock status and provide maximum sustainable yield (MSY) based reference points. The performance of a suite of harvest control rules were assessed through closed loop simulation (MSE-lite).


### Stock Structure Assumption {-}
Stock overview information: [reference to a Research Document or other publication, and/or footnote to the working paper]

[Mandatory. Maximum length 150 words.]
In 1987, two management units, 4RST and 3NOPs4VWX5Zc, were defined for Atlantic Halibut in Canadian waters, based primarily on tagging studies (McCracken 1958, Bowering 1986, Stobo et al. 1988) and differences in growth rates between fish caught on the Scotian Shelf and in the Gulf of St. Lawrence (Neilson and Bowering 1989). Evidence from genetics studies (Kess et al. 2021), conventional tagging (McCracken 1958, Bowering 1986, Stobo et al. 1988, Kanwit 2007, den Heyer et al. 2012), and recent satellite tagging (Le Bris et al. 2018, James et al. 2019) continues to support the separation. A recent review of the spatial ecology of Atlantic Halibut in the context of climate change summarizes this work, as well as the evidence for subpopulation structure within the existing management units (Shackell et al. 2022).

do the below:
- text directly from the Data Inputs Res Doc (better to just summarize by saying something like "this is the stock NAFO divisions and it was mostly defined using conventional tagging (ref Data Inputs)"?) - guidence doc looks like it's looking for some reference to ref points (ex: this management area is considered a single stock with a single set of reference points?)



### Reference Points {-}
[Mandatory. A table may be used in lieu of bullets for this section. Refer to the Guidance Document for further instructions.]

In March 2022, the SFGAC adopted an articulated harvest control rule (HCR) based on the index of exploitable biomass from the stratified Random Halibut Survey (Fig HCR XX), Although reference points have been defined based on SSB, reference points and control points for the HCR use the Stratified Random Halibut Survey biomass that correspond to SSB (LRP = 0.4B~MSY~, USR = 0.8 B~MSY~, and 1.2B~MSY~, Table XX) ref SAR? . 

Table 1. MSY-based reference points (limit reference point [LRP] and upper stock reference [USR]) and control points (LRP, USR, and 1.2 BMSY) for Atlantic Halibut spawning stock biomass (female) and Stratified Random Industry-DFO Halibut Longline Survey Biomass. Survey biomass estimates are derived from model equilibrium survey biomass at long-term fishing mortality rates that produce the corresponding female spawning stock biomass estimates shown. Catchability q is used to convert the index of exploitable biomass from the Stratified Random Halibut Survey (kg/1,000 hooks) to the Stratified Random Halibut Survey Biomass (kt).

```{r}
refpoints <- read.csv("C:/Users/harperd/Documents/Halibut/GitDataInputs/DataInputs_2023/DataInputs/FSAR/data/refpoints.csv")

colnames(refpoints) <- c("Model-estimated Biomass", "LRP", "USR", "BMSY", "1.2 BMSY", "Catability q")

knitr::kable(refpoints)
```


#### Other Stock Reference Points {-}
[If applicable (otherwise, do not include). A table may be used in lieu of bullets for this section.] 

#### Hatchery Reference Points {-}
[If applicable (otherwise, do not include). A table may be used in lieu of bullets for this section.] 

#### Habitat or Ecosystem Reference Points {-}
[If applicable (otherwise, do not include). A table may be used in lieu of bullets for this section.] 

### Harvest Decision Rule {-}
[Optional (otherwise, do not include). Maximum length 100 words.]
The HCR has two steps. First, a fishing mortality rate (F) is applied to the exploitable biomass estimated from the three-year mean of the Stratified Random Halibut Survey to calculate TAC. The articulated HCR includes three control points, the LRP, USR, and 1.2B~MSY~. Below the LRP, F is capped at 0.05 and above the USR it is relative to F~MSY~ (i.e., F = 0.8 F~MSY~ when  b = 0.8 B~MSY~) up to the third control point (1.2 BMSY), above which it is capped at 1.2 F~MSY~. Second, the percentage change from last year's TAX is limited by the sliding inter-annual TAC change threshold, where the limit of change ranges from 15% at the USR to 100% at the LRP. 

Two panel figure of the rule (figure 3 from last year's doc)

### Enhancement Plan {-}
[Optional (otherwise, do not include). Maximum length 100 words.]

### Habitat Restoration Plan {-}
[Optional (otherwise, do not include). Maximum length 100 words.]

### Data {-}
[Mandatory. Maximum length 300 words.] 

Data presented:

- Industry-DFO Halibut Longline Survey (Stratified Random Survey) halibut catch per 1000 hooks (2017-2024)

- DFO Maritimes Region Summer Ecosystem Survey (RV Summer Survey; NAFO Divs. 4VWX) number halibut per tow (1970-2024)

- Atlantic Halibut landings - Canada and international from Northwest Atlantic Fisheries Organization (NAFO)

- 2024 Preliminary Maritimes Region Atlantic Halibut Landings from MARFIS Database

Data changes: [Optional (otherwise, do not include).]

- Data from the 2021 RV Summer Survey that was not previously included due lack of conversion factors between new and old vessel have now been included in the time series following completion of comparative fishing and calculation of conversion factors. Imcomplete coverage remains in the following years: 2018, 2021, and 2022.

## ASSESSMENT {-}
[Mandatory panel of images comes first – example shown. Refer to the Guidance Document for options and further instructions.]

Stock status in interim years is assessed based on the three-year mean exploitable biomass index from the Stratified Random Halibut Survey. The next assessment framework review is currently scheduled for 2026. However, if the three-year mean index of exploitable biomass from the Stratified Random Halibut Survey falls above or below the 90% probability envelope from the closed-loop simulation for that index, a new framework will be triggered. <- not sure where to include this information...so just put it here for now 

(ref:caption) (A) Catch and Total Allowable Catch, (B) Spawning Stock Biomass (SSB) in relation to the Limit Reference Point (LRP; X) and (proposed) Upper Stock Reference (USR; X), (C) Fishing Mortality (F; X), Removal Reference (RR) and Natural Mortality (M), (D) Recruitment (numbers).

```{r base1, fig.cap="(ref:caption)"}


Assessment.Year=2024


source(file.path(getwd(), "directories.r"))
source(file.path(wd, "passwords.r")) 
setwd(file.path(wd,'FSAR'))

source("real-data.R")
source("plot-timeseries.R")


in.df <- fsar_data(Assessment.Year=Assessment.Year,datadir=datadir)

## English template
fsar_plot_base(in.df, "English")
```

[Optional. Add an additional optional panel of 4 indicators if relevant to developing advice.] 
[Optional. If the FSAR or FSRR is being used to provide advice for more than one stock, the mandatory panel can be repeated for each stock.]

### Historical and Recent Stock Trajectory and Trends {-}
[Mandatory. Maximum text length one page. Must be first section in Assessment section.]
Guidance calls this section Stock Status and Trends

[Report trends in relevant indicator(s), one per paragraph, in the following format:]

#### [Indicator 1] [specify name, e.g., Abundance]: {-}
[Provide values and trends for the stock (as applicable).]
Biomass - Halibut Longline Survey

The Industry-DFO Halibut Longline Survey provides an index of exploitable biomass of Atlantic Halibut. The survey is conducted by commercial fishermen with onboard observers between May and August, and spans NAFO areas 3NOP4VWX5Zc. The biomass index from the previous Fixed Station Halibut Survey (1998–2022) was used to provide harvest advice until 2021. Following the last framework in 2022, harvest advice in interim years is based now based on the three-year mean exploitable biomass index from the Stratified Random Survey. Inter-annual changes in the exploitable biomass index for the Stratified Random Survey have been minimal (Figure XX). 

With the switch to an exclusively stratified random design in 2023, there was also a change in the allocation of stations and stratification of the stock area. A new stratum in 2023 was introduced in an area in the northeast where there has been little to no catch of Atlantic Halibut in the Stratified Random Halibut Survey since 2017 (Figure XX). Allocation of stations in the new stratum was reduced, while for the rest of the strata allocation was proportional to area. 

Although not all stations were completed (Figure XX) this year, there was adequate coverage in all strata to effectively calculate the index of exploitable biomass from the Stratified Random survey. The 2024 index of exploitable biomass and the three-year mean are the highest in the time series. Inter-annual changes in the three-year mean have remained relatively stable through the time series with a small increase in 2024 from the previous year.

The LRP is 0.4 BMSY = 10.9 kt and the USR is 0.8 B~MSY~ = 21.8 kt (Table 2). The 2024 biomass from the Stratified Random Halibut Survey was estimated using a catchability of 0.002 (Table 2) to be 55.5 kt (95% confidence interval: 42.8, 68.2). Based on the three-year mean (Figure 4), the exploitable biomass index from the Stratified Random Halibut Survey for 2024 is 51.9 kt which is 1.9 B~MSY~, and substantially higher than the USR (Table 2), putting this stock in the healthy zone.

Include the survey station/strata map (same as last year but with no station locations/completion)


#### [Indicator 1] [specify name, e.g., Abundance]: {-}
[Provide values and trends for the stock (as applicable).]
Index of Recruitment - DFO Maritimes Region Summer Ecosystem Survey (RV Survey; NAFO Divs. 4VWX)

The DFO Maritimes Region Summer Ecosystem Survey (RV Survey; NAFO Divs. 4VWX) has been conducted typically between July–August since 1970. The RV survey provides an index of recruitment for Atlantic Halibut. The median size of Atlantic Halibut caught in the trawl survey is between 40–50 cm. The catch of Atlantic Halibut in this survey increased between 2000 and 2011 (Figure 4) but then declined to the lowest value since 2011 in 2020. Since 2020, catch has been increasing.

The 2018 and 2022 DFO Summer Ecosystem RV Survey did not cover all strata due to mechanical issues with the vessel; therefore, the abundance index from those two years cannot be compared with other years.

A conversion factor for Atlantic halibut was estimated (1:1) and data from the 2021 survey that was not included in previous years has now been incorporated into the time series, however converage remains incomplete.



The abundance index for 2024 is the second highest value in the time series. The length composition and the index of abundance are reviewed annually, exclusively for context and do not directly contribute to the HCR outputs.

Maybe include length composition figure?

#### [Indicator 1] [specify name, e.g., Abundance]: {-}
[Provide values and trends for the stock (as applicable).]
???

#### Unknown {-}
[Optional. Indicate what indicators are not available for the stock (as applicable); otherwise, do not include.]

### History of [Management; Landings/Harvest/Effort, TAC \& Catch Advice] {-}
[Mandatory. Title can vary. Maximum text length one page.] 
Landings? and Description of the Fishery?

Until 1988, the Atlantic halibut fishery in 3NOPs4VWX5Zc was unregulated. Within this management unit, Atlantic halibut are fished mainly by longline. A TAC of 3,200 t was first established in 1998 and, in response to an eight-year decline in landgins, was reduced to a low of 850 t in 1995. Since 1995, management plans and license conditions have required the released of Atlantic halibut less than 81 cm. Beginning in 1999, the TAC has increased and peaked at 5,507t in 2020 (Table XX, Figure XX). Every year since 2017, 100 t of the TAC have been set aside to cover catches by USA and France within the stock area (comment on recent change for france??).
NAFO statistics are used to describe removals up to and including 2023 because landings occur in two DFO regions (Maritimes and Newfoundland and Labrador (NL)). Other countries including Portugal, Spain, France, and the United States of America harvest Atlantic halibut from within the stock area as well, but outside Canada's Exclusive Economic Zone (EEZ). Landings from the Maritimes Fisheries Information System (MARFIS) database are reported as preliminary Canadian landings for 2024. These preliminary landgins do not include landings in NL. The majority of landgins occur on the Scotian Shelf (NAFO Divs. 4VWX)

Table of landings

Figure of landings? Actually going in the 4 panel...


#### History of Hatchery and Supplementation {-}
[If applicable (otherwise, do not include).]

#### History of Freshwater Habitat Impacts {-}
[If applicable (otherwise, do not include).]

### Projections [optional titles: Simulations, Future Scenarios] {-}
[Mandatory. Title can vary. Tables or figures can be used as necessary to support decision-making.]

### Ecosystem and Climate Change Considerations {-}
[Mandatory. Maximum 300 words. May be moved earlier in the Assessment section.]

The environmental conditions affecting the stock are currently unknown. Recent work (Czich paper and description!) has identified new information on the relationships between the biology of the fish (growth rates and age at maturation) and the environmental conditions (habitat and temperature). It is DFO’s intent to use this new information in the 2020 Framework Assessment to better inform DFO Science advice and the development of harvest strategies.

### Stock Advice {-}

Stock status in interim years is assessed based on the three-year mean exploitable biomass index from the Stratified Random Halibut Survey. Closed-loop simulation testing conducted at the last framework projected a gradual return to BMSY. So far, stock status has remained well above BMSY out-performing projections. The next assessment framework review is currently scheduled for 2026. However, if the three-year mean index of exploitable biomass from the Stratified Random Halibut Survey falls above or below the 90% probability envelope from the closed-loop simulation for that index, a new framework will be triggered.

### Harvest Decision Rule Outputs {-}
[If applicable (otherwise, do not include).] 
As the 2024 three-year mean exploitable biomass based on the Stratified Random Halibut Survey is higher than 1.2 BMSY (32.8 kt, Table XX), the highest fishing mortality, 1.2 FMSY (0.104), would be applied using the articulated HCR. This results in a TAC advice for 2025-2026 of 5,149 t. As this is within the maximum annual change of 15% from the 2024-2025 TAC (4,927 t), the 2025-2026 TAC advice is 5,149 t. This application of the HCR assumes all removals from Canadian and international fisheries are included. - directly from previous doc

### Evaluation of Exceptional Circumstances/Assessment Triggers {-}
[If applicable (otherwise, do not include).] 

## BYCATCH {-}
[Optional (otherwise, do not include).] 

## PROCEDURE FOR INTERIM YEAR UPDATES {-}
[Optional (otherwise, do not include).] 

## OTHER MANAGEMENT QUESTIONS {-}
[Optional (otherwise, do not include).] 

## SOURCES OF UNCERTAINTY {-}
[Mandatory. Maximum length 500 words.]

The Industry-DFO Halibut Longline Survey indices are the mean catch per tow and do not 
account for variation associated with hook competition, vessels, and temperature effects.

The interpretation of trends assumes no changes in vital rates, such as, growth or fecundity, 
that impact the dynamics of the population. Therefore, it is not known if, or how, vital rates and 
population growth rate will change with changes in stock size and/or variable environmental 
conditions. Further, there is uncertainty in the stock recruitment relationship and natural 
mortality in Halibut which can affect outcomes, especially reference points. <-re work this for something about ecosystem stuff. Potential for shifts in habitat usage/movement-spawning areas, wider distribution/migration <- look at data inputs doc

Data issues: ASO coverage/data quality/reliability issues. Reduced # of stations in some areas (although we had minimum coverage for index).


### Research Recommendations {-}
[Optional (otherwise, do not include).] 

## LIST OF MEETING PARTICIPANTS {-}
[Mandatory. The list is to include the name of each participant and their affiliation.]

## SOURCES OF INFORMATION {-}
[Mandatory.]


[References should only be included to understand the advice, and should not be included to give credit to the work as they are done in the Research Document. References to the unpublished working paper should be integrated as footnotes through this document.]

[References must be formatted according to the Guide for the Production of Fisheries and Oceans Canada Reports.]

[References listed in the References Cited section should be publicly available when the report is posted on the website. The only exception is for books or articles that have been accepted for publication (page proof stage or where there is an acceptance letter). These can be listed followed by the notation “in press”. All others should be included in the text as unpublished data or working papers.]
