---
title: "Update Advice"
author: "Brad"
date: "2022-10-13"
output: html_document
---


```{r setup, include=FALSE}
library(SpatialHub)
library(PBSmapping)


ds_all <- Mar.datawrangling::load_datasources()

devtools::load_all(".")

source(file.path(getwd(), "directories.r"))
source(file.path(wd, "passwords.r")) 
setwd(file.path(wd,'FSAR'))
Assessment.Year <- 2025


# Update data from database 

   # 1. Establish Oracle connection first (NEW REQUIREMENT)
    #cxn <- ROracle::dbConnect(DBI::dbDriver("Oracle"), uid, pwd, "PTRAN")

    #Mar.datawrangling::get_data(db='isdb',cxn=cxn, force.extract=T)
    # open and check setprofile_wide to confirm that you have year, latitude & longitude

    # Mar.datawrangling::get_data(db='rv',cxn=cxn,force.extract=T)


source("real-data.R")
in.df <- fsar_data(Assessment.Year=Assessment.Year,datadir=datadir)

language="English"

```


## Update data from database 

```{r }


# Update data from database 

# Mar.datawrangling::get_data(db='isdb',data.dir=datadir,fn.oracle.username = uid, fn.oracle.password = pwd, fn.oracle.dsn = 'ptran',usepkg='roracle',force.extract=T)
# open and check setprofile_wide to confirm that you have year, latitude & longitude

# RVdata<-get4VWXRV(uid, pwd, use.local=F,datadir=datadir)

```

# Landings
```{r}

  # top-left panel - Catch and TAC
  colrs<-c('#377eb8','#d53e4f','#7fbc41')
  idx <- which(in.df$panel.category == "Catch")
  yl <- c(0, max(in.df[idx, "ts.value"],na.rm=T) * 1.2)
  tl.df <- in.df[idx, ]

  if (language == "English") {
    x.lab <- "Year"
    y.lab <- "Catch (T)"
    legend.text <- c("Canada","Foreign", "TAC")
  }
  if (language == "French") {
    x.lab <- "Année"
    y.lab <- "Capture"
    legend.text <- c("Capture-tonne", "TAC-tonne")
  }


  y1<-min(tl.df$year)
  plm<-pivot_wider(tl.df,names_from = year, values_from = ts.value)
  plm<-as.matrix(plm[c(1,2,4),-(1:2)])
  plm[is.na(plm)]<-0


  barplot(plm,ylim=yl, axes = FALSE, col=c(colrs[1],colrs[2],colrs[1]),border=NA,density=c(NA,NA,40),axisnames=F)
  #axis(1,at=seq(0.7,(Assessment.Year-1959)*1.2,1.2),lab=F,tck=-0.01) ##change to 77 in sequence to increase axis length (ticks)
  axis(1,at=seq(0.7,(Assessment.Year-y1+1)*1.2,12),lab=seq(y1,Assessment.Year,10))##75 remains (no change) since 2023 not a multiple of 10
  axis(2, at = seq(0, yl[2], 1000), las = 1,
       labels = prettyNum(seq(0, yl[2],1000),
                          big.mark = ',',
                          scientific = FALSE))
  lines(seq(0.7,(Assessment.Year-y1+1)*1.2,1.2),in.df$ts.value[in.df$ts.name=="TAC"],col=colrs[3],lwd=3) ##change to 78 to accommodate additional year
  #legend("topright","(A)", bty = "n", cex=1.25)
  legend('topleft',
         legend.text,
         pch=c(15,15,NA),
         col=c(colrs[1],colrs[2],colrs[3]),
         lwd=c(NA,NA,3),
         box.lty=0,
         bg=NA,
         cex=0.8)

  mtext(side = 1, x.lab, line = 2, cex = 1)
  mtext(side = 2, y.lab, line = 3.2, cex = 1)
  box()

```
Figure (A) Northwest Atlantic Fisheries Organization (NAFO) reported Canadian (blue) and foreign (red) landings (tonnes) for 3NOPs4VWX5Zc Atlantic Halibut. Landings for 2024 (hashed bar) are preliminary, and taken from the Maritimes Fisheries Information System (MARFIS) as of October 31, 2024 (this does not include landings in Newfoundland which are not yet available). The solid green line is the Canadian total allowable catch (TAC). The NAFO 21A table of landings by country is reported by calendar year; however, the TAC for the stock is set for the period of April–March.

# Biomass
```{r}


  # top-right panel
  idx <- which(in.df$panel.category == "Biomass")
  yl <- c(0, max(in.df[idx, "ts.value"]) * 1.2)
  tr.df <- in.df[idx, ]
  #xl <- range(tr.df$year)
  xl <- c(1970,Assessment.Year+3)

  if (language == "English") {
    y.lab <- "Biomass (KT)"
    legend.text <- c("Halibut Survey (modelled)","95% confidence","Halibut Survey (projection)","90% projection envelope", "Halibut Survey (index)", "3-year mean of index","USR", "LRP")
  }
  if (language == "French") {
    y.lab <- "Biomasse KT"
    legend.text <- c("BSR-tonne", "confiance à 95%", "NRS-tonne", "NRL-tonne")
  }

  plot(ts.value ~ year, data = tr.df, type = "n", axes = FALSE, xlab = "", ylab = "", ylim = yl,xlim=xl)

  ## Vulnerable Biomass
  ## lower and upper
  yrs<-tr.df$year[tr.df$ts.name == "HSpredlow"]
  cil<-tr.df$ts.value[tr.df$ts.name == "HSpredlow"]
  cih<-tr.df$ts.value[tr.df$ts.name == "HSpredhigh"]

  polygon(  x = c(yrs,rev(yrs)),y = c(cil,rev(cih)), col = 'grey50', border = NA )
  #lines(ts.value ~ year, data = tr.df[which(tr.df$ts.name == "HSpredlow"), ], type = "l", lty = 2)
  #lines(ts.value ~ year, data = tr.df[which(tr.df$ts.name == "HSpredhigh"), ], type = "l", lty = 2)
  lines(ts.value ~ year, data = tr.df[which(tr.df$ts.name == "HSpred"), ], type = "l", lwd = 1,col='grey30',lty=1)

  ## Vulnerable Biomass projection
  ## lower and upper
  yrs<-tr.df$year[tr.df$ts.name == "HSprojlow"]
  cil<-tr.df$ts.value[tr.df$ts.name == "HSprojlow"]
  cih<-tr.df$ts.value[tr.df$ts.name == "HSprojhigh"]

  polygon(  x = c(yrs,rev(yrs)),y = c(cil,rev(cih)), col ='grey80', border = NA )
  #lines(ts.value ~ year, data = tr.df[which(tr.df$ts.name == "HSpredlow"), ], type = "l", lty = 2)
  #lines(ts.value ~ year, data = tr.df[which(tr.df$ts.name == "HSpredhigh"), ], type = "l", lty = 2)
  lines(ts.value ~ year, data = tr.df[which(tr.df$ts.name == "HSproj"), ], type = "l", lwd = 1,col='grey50',lty=2)

  ## SurveyBiomass
  lines(ts.value ~ year, data = tr.df[which(tr.df$ts.name == "HSobs"), ], type = "p",pch=16)
  lines(ts.value ~ year, data = tr.df[which(tr.df$ts.name == "HSobs_3yrm"), ], type = "l", lwd = 3,col='blue')


  ## LRP and USR
  abline(h=10.9, lty = 3, lwd = 2, col = "red")
  abline(h=21.8, lty = 3, lwd = 2, col = "forestgreen")
  #lines(ts.value ~ year, data = tr.df[which(tr.df$ts.name == "SSBlrp-MT"), ], type = "l")
  #lines(ts.value ~ year, data = tr.df[which(tr.df$ts.name == "SSBusr-MT"), ], type = "l"

  legend("topright","(B)", bty = "n", cex=1.25)
  legend("topleft",
         legend.text,
         lty = c(1, NA, 2, NA, NA, 1, 3, 3),
         lwd = c(1, NA, 1, NA, NA, 3, 2, 2),
         pch = c(NA,15,NA,15,16,NA,NA,NA),
         col = c('grey30','grey50','grey50','grey80',"black", "blue", "forestgreen", "red"),
         box.lty=0,
         bg=NA,
         cex=0.8
  )

  axis(side = 1, padj = -0.5)
  axis(side = 2, las = 1, hadj = 0.9)
  mtext(side = 1, x.lab, line = 2, cex = 0.75)
  mtext(side = 2, y.lab, line = 3, cex = 0.75)
  box()


```
Figure  (B) Stratified Random Halibut Survey Biomass (B) in relation to the Limit Reference Point (LRP; 10.9 kt) and Upper Stock Reference (USR; 21.8 kt).

# Mortality
```{r}


  ## bottom-left panel
  idx <- which(in.df$panel.category == "Fishing")
  yl <- c(0, max(in.df[idx, "ts.value"]) * 1.2)
  bl.df <- in.df[idx, ]

  if (language == "English") {
    y.lab <- "Mortality (1/yr)"
    legend.text <- c("Fishing", "95% confidence", "Natural", "95% confidence", "1.2FMSY", "FMSY", "0.8FMSY")
  }
  if (language == "French") {
    y.lab <- "Mortalité"
    legend.text <- c("F-1/yr", "confiance à 95%", "RP-1/yr", "M-1/yr")
  }

  plot(ts.value ~ year, data = bl.df, type = "n", axes = FALSE, xlab = "", ylab = "", ylim = yl,xlim=xl)

  ## Fishing
  yrs<-bl.df$year[bl.df$ts.name == "Utlow"]
  cil<-bl.df$ts.value[bl.df$ts.name == "Utlow"]
  cih<-bl.df$ts.value[bl.df$ts.name == "Uthigh"]

  polygon(  x = c(yrs,rev(yrs)),y = c(cil,rev(cih)), col = 'grey', border = NA )
  lines(ts.value ~ year, data = bl.df[which(bl.df$ts.name == "Ut"), ], type = "l", lwd = 1,col='grey40',lty=1)

  ## natural mortality
  yrs<-bl.df$year[bl.df$ts.name == "Mtlow"]
  cil<-bl.df$ts.value[bl.df$ts.name == "Mtlow"]
  cih<-bl.df$ts.value[bl.df$ts.name == "Mthigh"]

  polygon(  x = c(yrs,rev(yrs)),y = c(cil,rev(cih)), col = rgb(1,0,0,0.2), border = NA )
  lines(ts.value ~ year, data = bl.df[which(bl.df$ts.name == "Mt"), ], type = "l", lwd = 1,col='red',lty=2)

  #removal reference
  abline(h=0.0877*c(1.2,1,0.8), lty = 3, lwd = 2,col=c("blue", "forestgreen", "orange"))

  #lines(ts.value ~ year, data = bl.df[which(bl.df$ts.name == "M-1/yr"), ], lty = 1, lwd = 2, col = grey(0.5))
  
  B<-subset(in.df,ts.name=="HSobs_3yrm")
  L<-subset(in.df,panel.category=="Catch"&year%in%B$year)
  U<-(L$ts.value[L$ts.name=="CanadaTotal"]+L$ts.value[L$ts.name=="ForeignTotal"])/(B$ts.value*1000)
  f<- log(1-U)*-1
  points(B$year,f,pch=16)


  legend("topright","(C)", bty = "n", cex=1.25)
  legend("topleft",
         legend.text,
         lty = c(1, NA, 2, NA, 3, 3, 3),
         lwd = c(1, NA, 1, NA, 2, 2, 2),
         pch = c(NA, 15, NA, 15, NA, NA, NA),
         col = c('grey40','grey','red',rgb(1,0,0,0.2), "blue", "forestgreen", "orange"),
         box.lty = 0,
         bg=NA,
         cex=0.8
  )

  axis(side = 1, padj = -0.5)
  axis(side = 2, las = 1, hadj = 0.9)
  # axis(side = 4, las = 1, hadj = 0.9)
  mtext(side = 1, x.lab, line = 2, cex = 0.75)
  mtext(side = 2, y.lab, line = 3, cex = 0.75)

  box()


```
Figure (C) Fishing and Natural Mortality in relation to the F~MSY~ Removal References. 

# Recruitment

```{r}
  ## bottom-right panel
  idx <- which(in.df$panel.category == "Recruitment")
  yl <- c(0, max(in.df[idx, "ts.value"]) * 1.2)
  br.df <- in.df[idx, ]

  if (language == "English") {
    y.lab <- "Abundance (millions)"
    legend.text <- c("RV Survey (modelled)", "95% confidence", "RV Survey (Index)")
  }
  if (language == "French") {
    y.lab <- "Recrutement"
    legend.text <- c("R-E06", "confiance à 95%")
  }
  plot(ts.value ~ year, data = br.df, type = "n", axes = FALSE, xlab = "", ylab = "", ylim = yl)
  ## Vulnerable Abundance
  ## lower and upper
  yrs<-br.df$year[br.df$ts.name == "RVpredlow"]
  cil<-br.df$ts.value[br.df$ts.name == "RVpredlow"]
  cih<-br.df$ts.value[br.df$ts.name == "RVpredhigh"]

  polygon(  x = c(yrs,rev(yrs)),y = c(cil,rev(cih)), col = 'grey', border = NA )
  lines(ts.value ~ year, data = br.df[which(br.df$ts.name == "RVpred"), ], type = "l", lwd = 1,col='grey40',lty=1)

  ## Survey Abundance
  lines(ts.value ~ year, data = br.df[which(br.df$ts.name == "RVobs"), ], type = "p",pch=16)

  legend("topright","(D)", bty = "n", cex=1.25)
  legend("topleft",
         legend.text,
         lty = c(1, NA, NA),
         lwd = c(1, NA, NA),
         pch = c(NA,15,16),
         col = c('grey40','grey',"black"),
         box.lty = 0,
         bg=NA,
         cex=0.8
  )


  axis(side = 1, padj = -0.5)
  mtext(side = 1, x.lab, line = 2, cex = 0.75)
  axis(side = 2, las = 1, hadj = 0.9)
  mtext(side = 2, y.lab, line = 3, cex = 0.75)
  box()



```
Figure (D) Recruitment index from the RV Summer Survey in numbers.

```{r, message=FALSE, echo = FALSE, warning=FALSE, results="hide", fig.cap="(ref:caption)"}
 
library(SpatialHub)
library(PBSmapping)
load(file.path(datadir,"SurveyStrata2022.RData"))
BO1<-data.frame(PID=1,POS=1:6,X=c(-47,-55,-55,-56.15283,-56.15283,-47),Y=c(45.5,45.5,46.25,46.25,48,48))

BO1<-joinPolys(BO1,surveyPolyLL,operation="INT")


BO2<-data.frame(PID=1,POS=1:4,X=c(-68,-68,-67,-67),Y=c(41,46,46,41))

BO2<-joinPolys(BO2,surveyPolyLL,operation="INT")

HalSurvey2024 <- read.csv("fsardata/2024 Halibut Longline Survey Stations Assigned.csv")
completed2024 <- read.csv("fsardata/2024 Halibut Longline Survey Stations Completed.csv")

xl=c(-68,-47)
yl=c(40,48)
png(file.path(wd,'figures',"SurveyMap.png"), width =8, height = 6,units='in',pointsize=12, res=300,type='cairo')

bioMap(xlim = xl, ylim = yl,mapRes="MR",poly.lst = list(surveyStrataPolyLL,props), plot.rivers=F,isobaths = NULL)
addPolys(BO1,density=20,col='black')
sp::plot(NAFO,add=T)
 points(lat.DecDeg~lon.DecDeg,HalSurvey2024,pch=21,bg = "yellow",cex=0.5)
 points(lat.DecDeg_Assigned~lon.DecDeg_Assigned,completed2024,pch=21, bg = "black",cex=0.5)
text(x = -56.5, y = 43, labels = "4Vs")
text(x = -59.55, y = 46.37, labels = "4Vn")
text(x = -61, y = 42, labels = "4W")
text(x = -64.5, y = 41.5, labels = "4X")
text(x = -66.25, y = 41.9, labels = "5Zc")
text(x = -48.75, y = 42.5, labels = "3N")
text(x = -53.2, y = 43.7, labels = "3O")
text(x = -55.5, y = 45.9, labels = "3Ps")
text(x = -58.7, y = 47.35, labels = "3Pn")
dev.off()
```
Figure 3. The Stratified Random Industry-DFO Halibut Longline survey area is separated into five zones across the Scotian Shelf (SS) and southern Grand Banks (GB) represented by different colours. Each zone is further separated into depth strata, shown in different shades of each colour. The hatched area is a new strata with lower station allocation introduced in 2023. The survey area includes division 3Pn, which is not part of the management unit. Two hundred stations were assigned in the 2024 survey, black dots indicate completed stations and yellow dots indicate incomplete stations. Solid black lines represent Northwest Atlantic Fisheries Organization Division boundaries, while international boundaries are marked with dashed lines.


