---
title: "DataChecking"
author: "Brad"
date: "April 21, 2021"
output: html_document
---

```{r setup, include=FALSE}
library(devtools)
#install_github("Maritimes/Mar.datawrangling")
#install_github("Maritimes/Mar.bycatch")
 library(Mar.datawrangling)
 library(Mar.bycatch)
 library(tidyverse)

source(file.path(getwd(), "directories.r"))
#datadir="C:/Users/hubleyb/Documents/Halibut/data"
#wd="C:/Users/hubleyb/Documents/Halibut/Git/DataInputs/"
#datadir="C:/Users/harperd/Documents/Halibut/RDataVault"
#wd="C:/Users/harperd/Documents/Halibut/GitDataInputs/DataInputs/"
source(file.path(wd, "passwords.r"))


```




## Checking difference between ISFISHLENGTHS and ISFISH
```{r}
#sum(abs(FixedSurvey2[,paste0("L",seq(10,260,5))]-FixedSurvey[,paste0("L",seq(10,260,5))]),na.rm=T)
#sum(FixedSurvey2[,paste0("L",seq(10,260,5))],na.rm=T)
#sum(FixedSurvey[,paste0("L",seq(10,260,5))],na.rm=T)

#FixedSurvey2020_2 <- subset(FixedSurvey2,YEAR==2020)
#FixedSurvey2020 <- subset(FixedSurvey,YEAR==2020)

#FixedSurvey2020[rowSums(abs(FixedSurvey2020_2[,paste0("L",seq(10,260,5))]-FixedSurvey2020[,paste0("L",seq(10,260,5))]),na.rm=T)>0,]


#subset(FixedSurvey,TRIP=="J20-0019"&SET_NO==2)
#subset(FixedSurvey2,TRIP=="J20-0019"&SET_NO==2)
#subset(isdb$ISFISH,CATCH_ID==101911351&SEXCD_ID==0)


LFcheck<-checkLF(datadir=datadir)

subset(LFcheck,!MATCH&YEAR==2020&SETCD_ID==5)
subset(HALIBUTSURVEY,!MATCH&YEAR==2020&TRIP=="J20-0020A"&SETCD_ID%in%4:5)
subset(LFcheck,!MATCH&YEAR==2020&SETCD_ID==5&!is.na(SMPL_ID))


 sum(subset(LFcheck,YEAR==2020&SETCD_ID==5)$ISFISH_COUNT)

 sum(subset(LFcheck,YEAR==2020&SETCD_ID==5)$NUM_AT_LENGTH)
 
 sum(subset(LFcheck,SETCD_ID==5)$ISFISH_COUNT)
 sum(subset(LFcheck,SETCD_ID==5)$NUM_AT_LENGTH)
 sum(subset(LFcheck,SETCD_ID==4)$ISFISH_COUNT)
 sum(subset(LFcheck,SETCD_ID==4)$NUM_AT_LENGTH)
 sum(subset(LFcheck,SETCD_ID==10)$ISFISH_COUNT)
 sum(subset(LFcheck,SETCD_ID==10)$NUM_AT_LENGTH)
 

```
## Checked a few differences in the data sheets and they persisted on the page, i.e. length frequency form was inconsistant with halibut survey tagging observations. There are more fish observed in ISFISHLENGTHS so go with that for size composition. 

# checking for duplicate stations
```{r}
FixedSurvey2<-FixedSurveyData(datadir=datadir,by.sex=F)
FixedSurvey2$SY <- paste0(FixedSurvey2$YEAR,FixedSurvey2$STATION)
dr<-subset(FixedSurvey2,duplicated(SY))
fsdups<-subset(FixedSurvey2,SY%in%dr$SY)
fsdups
RandomSurvey2<-RandomSurveyData(datadir=datadir,by.sex=F)
RandomSurvey2$SY <- paste0(RandomSurvey2$YEAR,RandomSurvey2$STATION)
dr<-subset(RandomSurvey2,duplicated(SY))
fsdups<-subset(RandomSurvey2,SY%in%dr$SY)
fsdups
```


```{r}

FixedSurvey<-FixedSurveyData(datadir=datadir,by.sex=T, bins=bins, LF.from = "ISFISH",correct.splitsets = F)
FixedSurvey2<-FixedSurveyData(datadir=datadir,by.sex=T, bins=bins, LF.from = "ISFISHLENGTHS",correct.splitsets = F)

  Yrs = 2009:2020

  fixedLF=constructLF(LFdata=FixedSurvey,bins=bins, Yrs = Yrs)
  fixedLF2=constructLF(LFdata=FixedSurvey2,bins=bins, Yrs = Yrs)

      BarPlotLF(fixedLF,yrs=Yrs,rel=T,filen=file.path(wd,"figures","fixedLF3"))
    BarPlotLF(fixedLF2,yrs=Yrs,rel=T,filen=file.path(wd,"figures","fixedLF4"))
    
    
  HALIBUTSURVEY<- FixedSurvey 
    
  HALIBUTSURVEY$SY <- paste0(HALIBUTSURVEY$YEAR,HALIBUTSURVEY$STATION,HALIBUTSURVEY$SEXCD_ID)
  dr<-subset(HALIBUTSURVEY,duplicated(SY))

  fsdups<-subset(HALIBUTSURVEY,SY%in%dr$SY)

  newd<-fsdups %>%
    group_by(YEAR,STATION,VESS_ID,SEXCD_ID ) %>%
    summarise(SOAKMINP3P1=sum(SOAKMINP3P1),NUM_HOOK_HAUL=sum(NUM_HOOK_HAUL),across(EST_NUM_CAUGHT:NUM_MEASURED, sum, na.rm=T)) %>%
    right_join(.,dplyr::select(dr,YEAR,STATION,which(!names(dr)%in%names(.)))) %>%
    filter(!duplicated(SY)) %>%
    data.frame()

  HALIBUTSURVEY <- left_join(subset(HALIBUTSURVEY,!duplicated(SY)),newd)
  fixedLF=constructLF(LFdata=HALIBUTSURVEY,bins=bins, Yrs = Yrs)
      BarPlotLF(fixedLF,yrs=Yrs,rel=T,filen=file.path(wd,"figures","fixedLF5"))

```

Checking length against weight for random survey halibut


```{r}
halibut <- ISSETPROFILE_WIDE %>%
  filter(YEAR == 2020) %>%
  left_join(ISFISHSETS, by = "FISHSET_ID")%>%
  filter(SETCD_ID == 5) %>%
  left_join(ISCATCHES, by = "FISHSET_ID") %>%
  filter(SPECCD_ID == 30) %>%
  left_join(ISFISH, by = "CATCH_ID")

halibut$SEXCD_ID <- as.character(halibut$SEXCD_ID)

ggplot(halibut, aes(x=FISH_LENGTH, fill=SEXCD_ID, color=SEXCD_ID)) +
  geom_histogram(binwidth = 1, position="identity", alpha=0.5) +
  theme_classic()
```

