---
title: "FisheryData"
author: "Brad"
date: "March 1, 2021"
output: html_document
---

```{r setup, include=FALSE}
library(devtools)
#install_github("Maritimes/Mar.datawrangling")
#1
 library(Mar.datawrangling)
 library(Mar.fleets)
 library(tidyverse)
library(CircStats)
library(lubridate)
library(rgdal)
#install_github("BradHubley/SpatialHub")
library(SpatialHub)


source(file.path(getwd(), "directories.r"))
#datadir="C:/Users/hubleyb/Documents/Halibut/data"
#wd="C:/Users/hubleyb/Documents/Halibut/Git/DataInputs/"
#datadir="C:/Users/harperd/Documents/Halibut/RDataVault"
#wd="C:/Users/harperd/Documents/Halibut/GitDataInputs/DataInputs/"
source(file.path(wd, "passwords.r"))
ds_all <- Mar.datawrangling::load_datasources()
devtools::load_all(".")

lwA=0.006803616
lwB=3.119924

```

## Update data from database 
## Marfis from data,wrangling
```{r }

#get_data(db='marfis',data.dir=datadir,fn.oracle.username = uid, fn.oracle.password = pwd, fn.oracle.dsn = 'ptran',usepkg='roracle',force.extract=T)

MARFISdataLL<-getMARFISdata(datadir,sp=130,gear=1)
MARFISdataLL$CLASS <-NA
MARFISdataLL$CLASS[MARFISdataLL$LOA<35]<-"LL35"
MARFISdataLL$CLASS[MARFISdataLL$LOA>=35&MARFISdataLL$LOA<45]<-"LL35-45"
MARFISdataLL$CLASS[MARFISdataLL$LOA>=45&MARFISdataLL$LOA<65]<-"LL45-65"
MARFISdataLL$CLASS[MARFISdataLL$LOA>65]<-"LL65"
MARFISdataOT<-getMARFISdata(datadir,sp=130,gear=2)
MARFISdataOT$CLASS <-"OT"

MARFISdata<-rbind(MARFISdataLL,MARFISdataOT)

```

# Make Tables by class

```{r }
yrs1<-2002:2011
yrs2<-2012:2021

table1<- MARFISdata %>% filter(!is.na(CLASS)&YEAR%in%yrs1&!is.na(NAFO)) %>% group_by(NAFO,CLASS) %>% summarise(Landings = round(sum(RND_WEIGHT_KGS,na.rm=T)/1000)) %>% pivot_wider(names_from = CLASS,values_from = Landings) %>% tibble()
table1<-table1[c("NAFO","LL35", "LL35-45", "LL45-65", "LL65", "OT")]


table2<- MARFISdata %>% filter(!is.na(CLASS)&YEAR%in%yrs2&!is.na(NAFO)) %>% group_by(NAFO,CLASS) %>% summarise(Landings = round(sum(RND_WEIGHT_KGS,na.rm=T)/1000)) %>% pivot_wider(names_from = CLASS,values_from = Landings) %>% tibble()
table2<-table2[c("NAFO","LL35", "LL35-45", "LL45-65", "LL65", "OT")]

write.csv(table1,file.path(datadir,"MARFIStable1.csv"),row.names=F)
write.csv(table2,file.path(datadir,"MARFIStable2.csv"),row.names=F)


```



# Make Tables by month

```{r }
yrs1<-2016:2018
yrs2<-2019:2021

MARFISdata$MONTH<-month(MARFISdata$DATE_FISHED)

table1<- MARFISdata %>% filter(YEAR%in%yrs1&!is.na(NAFO)) %>% group_by(NAFO,MONTH) %>% summarise(Landings = round(sum(RND_WEIGHT_KGS,na.rm=T)/1000)) %>% pivot_wider(names_from = MONTH,values_from = Landings) %>% tibble()


table2<- MARFISdata %>% filter(!is.na(CLASS)&YEAR%in%yrs2&!is.na(NAFO)) %>% group_by(NAFO,CLASS) %>% summarise(Landings = round(sum(RND_WEIGHT_KGS,na.rm=T)/1000)) %>% pivot_wider(names_from = CLASS,values_from = Landings) %>% tibble()

write.csv(table1,file.path(datadir,"MARFIS_nafo_month_2016-2018.csv"),row.names=F)
write.csv(table2,file.path(datadir,"MARFIS_nafo_month_2019-2021.csv"),row.names=F)


```




# plot annual
```{r}
yrs<-2001:2005
lvls=seq(0,7000,1000)
xl=c(-69,-47)
yl=c(40,48)

  tmp<-subset(MARFISdata,YEAR%in%yrs)
  catch.data <- na.omit(data.frame(EID=1:nrow(tmp),X=tmp$LONGITUDE, Y=tmp$LATITUDE, Z=tmp$RND_WEIGHT_KGS  ))


catch.grids <- SpatialHub::gridData(catch.data,FUN=sum,lvls=lvls,grid.size=10,border=NA)
png(file.path(wd,'figures',paste0('marfisCatch',min(yrs),'-',max(yrs),'.png')),width=9,height=5,units='in',pointsize=12, res=300,type='cairo')
SpatialHub::bioMap(xlim = xl, ylim = yl, mapRes="MR", poly.lst=catch.grids)
SpatialHub::contLegend('bottomright', lvls=lvls/1000, Cont.data=catch.grids, title="t",cex=.7)
dev.off()

```

### Loop for annual

```{r}
yrs=2001:2021
lvls=seq(0,7000,1000)
xl=c(-69,-47)
yl=c(40,48)

for(y in yrs){
  #catch.data <- na.omit(with(subset(marf,year(DATE_FISHED)==y),data.frame(LOG_EFRT_STD_INFO_ID,LONGITUDE,LATITUDE,RND_WEIGHT_KGS)))
  #catch.data$LOG_EFRT_STD_INFO_ID <- 1:nrow(catch.data)
  #summary(catch.data)
  tmp<-subset(MARFISdata,YEAR==y)
  catch.data <- na.omit(data.frame(EID=1:nrow(tmp),X=tmp$LONGITUDE, Y=tmp$LATITUDE, Z=tmp$RND_WEIGHT_KGS  ))
  catch.grids <- SpatialHub::gridData(catch.data,FUN=sum,lvls=lvls,grid.size=20)
  
  #pdf(file.path(wd,'figures',paste0('marfisCatch',y,'.pdf')),width=10,height=7)
    png(file.path(wd,'figures',paste0('marfisCatch',y,'.png')),width=9,height=5,units='in',pointsize=12, res=300,type='cairo')
    
  
  SpatialHub::bioMap(xlim = xl, ylim = yl, mapRes="MR", poly.lst=catch.grids)
  SpatialHub::contLegend('bottomright', lvls=lvls/1000, Cont.data=catch.grids, title="MT",cex=0.8)
  dev.off()

}



```


### Loop for quarterly year blocks

```{r}
q=1:4
yrblocks <- list(2001:2005,2006:2010,2011:2015,2016:2020)

lvls=seq(0,7000,1000)
xl=c(-69,-47)
yl=c(40,48)
  
for(i in q){
  for(j in 1:length(yrblocks)){
    #catch.data <- na.omit(with(subset(marf,quarter(DATE_FISHED)==i&year(DATE_FISHED)%in%yrblocks[[j]]),data.frame(LOG_EFRT_STD_INFO_ID,LONGITUDE,LATITUDE,RND_WEIGHT_KGS)))
    #catch.data$LOG_EFRT_STD_INFO_ID <- 1:nrow(catch.data)
    #print(summary(catch.grids))
  tmp<-subset(MARFISdata,quarter(DATE_FISHED)==i&YEAR%in%yrblocks[[j]])
  catch.data <- na.omit(data.frame(EID=1:nrow(tmp),X=tmp$LONGITUDE, Y=tmp$LATITUDE, Z=tmp$RND_WEIGHT_KGS  ))
  catch.grids <- SpatialHub::gridData(catch.data,FUN=sum,lvls=lvls,grid.size=20)

    #pdf(file.path(wd,'figures',paste0('marfisCatch','Q',i,min(yrblocks[[j]]),'-',max(yrblocks[[j]]),'.pdf')),width=10,height=7)
    png(file.path(wd,'figures',paste0('marfisCatch','Q',i,min(yrblocks[[j]]),'-',max(yrblocks[[j]]),'.png')),width=9,height=5,units='in',pointsize=12, res=300,type='cairo')
    
    SpatialHub::bioMap(xlim = xl, ylim = yl, mapRes="MR", poly.lst=catch.grids,title = paste('Q',i,min(yrblocks[[j]]),'-',max(yrblocks[[j]])),nafo=T)
    SpatialHub::contLegend('bottomright', lvls=lvls/1000, Cont.data=catch.grids, title="t",cex=0.8)
  
    dev.off()
  }
}


	



```


### Loop for monthly

```{r}
	require(animation)# for creating gif
  yrs = 2002:2020
	### GIF animations ###
	## set some options first
	oopt = ani.options(interval = 0.4, nmax = length(yrs)* 12, outdir=file.path(wd,'figures'))
	saveGIF({
  for(y in yrs){
    for(i in 1:12){
    tmp<-subset(MARFISdata,month(DATE_FISHED)==i&YEAR%in%yrblocks[[j]])
    catch.data <- data.frame(EID=1:nrow(tmp),X=tmp$LONGITUDE, Y=tmp$LATITUDE, Z=tmp$RND_WEIGHT_KGS  )
    catch.grids <- SpatialHub::gridData(catch.data,FUN=sum,lvls=lvls,grid.size=20)
      #print(summary(catch.grids))
      
      #pdf(file.path(wd,'figures',paste0('marfisCatch',y,i,'.pdf')),width=10,height=7)
      
      SpatialHub::bioMap(xlim = xl, ylim = yl, mapRes="MR", poly.lst=catch.grids,title = paste(month(i,label=T),y))
      SpatialHub::contLegend('bottomright', lvls=lvls/1000, Cont.data=catch.grids, title="MT",cex=0.8)
    #dev.off()
    
  }

	  ani.pause() ## pause for a while (’interval’)
	}
	}, interval = 0.1, movie.name = "marfis.gif", ani.width = 900, ani.height = 600)



```


```{r}
marf$YEAR<-year(marf$DATE_FISHED)
marf$NAFO <-substr(marf$NAFO_AREA,1,2)
marf.tab<-marf %>% group_by(YEAR,NAFO) %>% summarise(catch=round(sum(RND_WEIGHT_KGS/1000,na.rm=T),2))%>% pivot_wider(.,names_from = NAFO,values_from = catch)
marf.tab
write.csv(marf.tab,"MarfisHalibutLandings_t.csv",row.names=F)
```

```{r}
 SpatialHub::bioMap("WSS", mapRes="MR",nafo='all',main=2020)
 
 points(LATITUDE~LONGITUDE,subset(marf,YEAR==2020&NAFO=="4X"))
 points(LATITUDE~LONGITUDE,subset(marf,YEAR==2020&NAFO=="5Z"),col='blue')
 
 SpatialHub::bioMap("WSS", mapRes="MR",nafo='all',main=2019)
 
 points(LATITUDE~LONGITUDE,subset(marf,YEAR==2019&NAFO=="4X"))
 points(LATITUDE~LONGITUDE,subset(marf,YEAR==2019&NAFO=="5Z"),col='blue')
 
```

## Other flatfish landings
SPECIES_CODE SPECIES_NAME
140          AMERICAN PLAICE
141          YELLOWTAIL
142          GREYSOLE/WITCH
143          WINTER FLOUNDER
144          GREENLAND HALIBUT/TURBOT
149          FLOUNDER, UNSPECIFIED

# get data
```{r}

MARFISdataFF<-getMARFISdata(datadir,sp=c(140,141,142,143,144,145,149))



 tlff<-with(MARFISdataFF,tapply(RND_WEIGHT_KGS,YEAR,sum,na.rm=T))
 tl140<-with(subset(MARFISdataFF,SPECIES_CODE==140),tapply(RND_WEIGHT_KGS,YEAR,sum,na.rm=T))
 tl141<-with(subset(MARFISdataFF,SPECIES_CODE==141),tapply(RND_WEIGHT_KGS,YEAR,sum,na.rm=T))
 tl142<-with(subset(MARFISdataFF,SPECIES_CODE==142),tapply(RND_WEIGHT_KGS,YEAR,sum,na.rm=T))
 tl143<-with(subset(MARFISdataFF,SPECIES_CODE==143),tapply(RND_WEIGHT_KGS,YEAR,sum,na.rm=T))
 tl144<-with(subset(MARFISdataFF,SPECIES_CODE==144),tapply(RND_WEIGHT_KGS,YEAR,sum,na.rm=T))
 tl145<-with(subset(MARFISdataFF,SPECIES_CODE==145),tapply(RND_WEIGHT_KGS,YEAR,sum,na.rm=T))
 tl149<-with(subset(MARFISdataFF,SPECIES_CODE==149),tapply(RND_WEIGHT_KGS,YEAR,sum,na.rm=T))
  

```


# plot
```{r}
plot(as.numeric(names(tlff)),tlff/1000,type='l',xlab='',ylab="Total Flatfish Landings (t)",ylim=c(0,5000))
 lines(as.numeric(names(tl140)),tl140/1000,lty=2,col=2)
 lines(as.numeric(names(tl141)),tl141/1000,lty=3,col=3)
 lines(as.numeric(names(tl142)),tl142/1000,lty=4,col=4)
 lines(as.numeric(names(tl143)),tl143/1000,lty=5,col=5)
 lines(as.numeric(names(tl144)),tl144/1000,lty=6,col=6)
 lines(as.numeric(names(tl145)),tl145/1000,lty=7,col=7)
 lines(as.numeric(names(tl149)),tl149/1000,lty=8,col=8)
 legend('topright',legend=c(140,141,142,143,144,145,149),lty=2:8,col=2:8)

```



## Port sample data
```{r}
bins=seq(0,260,5)
PortData<-PortSampleData(datadir=datadir,by.sex=T, bins=bins)


 Yrs1 = 1985:2002 
 Yrs2 = 2003:2020 
 #dput(sort(unique(ObserverData$NAFAREA_ID)))
 nafo3 = c( "3N", "3NA", "3NB", "3NC", "3ND", "3NE", "3NF", "3O", "3OA", "3OC", "3OD", "3OE", "3PN", "3PS", "3U")
 nafo4 = c("4V","4VB", "4VC", "4VN", "4VNW", "4VS", "4VSB", "4VSC","4VSW","4VW","4VX", "4W", "4WD", "4WE", "4WF", "4WG", "4WH", "4WJ", "4WK", "4WL", "4WM","4WX","4XMNOPQ", "4X", "4XL", "4XM", "4XN", "4XO", "4XP", "4XQ", "4XR", "4XU", "4XS", "4XRS", "5YB", "5ZJ", "5ZM", "5ZN","5ZEJ")
 

  portNAFO3OT1=constructLF(LFdata=subset(PortData,NAFO%in%nafo3&GEAR=="OT"),bins=bins, Yrs = Yrs1)
  BarPlotLF(portNAFO3OT1,yrs=Yrs1,rel=F,filen=file.path(wd,"figures","portNAFO3OT1"))

  portNAFO3OT2=constructLF(LFdata=subset(PortData,NAFO%in%nafo3&GEAR=="OT"),bins=bins, Yrs = Yrs2)
  BarPlotLF(portNAFO3OT2,yrs=Yrs2,rel=F,filen=file.path(wd,"figures","portNAFO3OT2"))

  portNAFO3LL1=constructLF(LFdata=subset(PortData,NAFO%in%nafo3&GEAR=="LL"),bins=bins, Yrs = Yrs1)
  BarPlotLF(portNAFO3LL1,yrs=Yrs1,rel=F,filen=file.path(wd,"figures","portNAFO3LL1"))

  portNAFO3LL2=constructLF(LFdata=subset(PortData,NAFO%in%nafo3&GEAR=="LL"),bins=bins, Yrs = Yrs2)
  BarPlotLF(portNAFO3LL2,yrs=Yrs2,rel=F,filen=file.path(wd,"figures","portNAFO3LL2"))

  portNAFO4OT1=constructLF(LFdata=subset(PortData,NAFO%in%nafo4&GEAR=="OT"),bins=bins, Yrs = Yrs1)
  BarPlotLF(portNAFO4OT1,yrs=Yrs1,rel=F,filen=file.path(wd,"figures","portNAFO4OT1"))

  portNAFO4OT2=constructLF(LFdata=subset(PortData,NAFO%in%nafo4&GEAR=="OT"),bins=bins, Yrs = Yrs2)
  BarPlotLF(portNAFO4OT2,yrs=Yrs2,rel=F,filen=file.path(wd,"figures","portNAFO4OT2"))

  portNAFO4LL1=constructLF(LFdata=subset(PortData,NAFO%in%nafo4&GEAR=="LL"),bins=bins, Yrs = Yrs1)
  BarPlotLF(portNAFO4LL1,yrs=Yrs1,rel=F,filen=file.path(wd,"figures","portNAFO4LL1"))

  portNAFO4LL2=constructLF(LFdata=subset(PortData,NAFO%in%nafo4&GEAR=="LL"),bins=bins, Yrs = Yrs2)
  BarPlotLF(portNAFO4LL2,yrs=Yrs2,rel=F,filen=file.path(wd,"figures","portNAFO4LL2"))
  
 

```


## Port sample data plots for document
```{r}
bins=seq(0,260,5)
PortData<-PortSampleData(datadir=datadir,by.sex=T, bins=bins)
PortData$FISHERY<-PortData$GEAR
PortData$FISHERY[!is.na(PortData$TRIP)]<-"CI"
PortSampleTable<-PortData %>% 
  group_by(FISHERY,YEAR)%>% 
  distinct(SAMPLE)%>%
  count()%>%
  pivot_wider(names_from = FISHERY,values_from = n)%>%
  arrange(YEAR)

PortSampleTable[is.na(PortSampleTable)]<-0

write.csv(PortSampleTable,"PortSampleTable.csv",row.names=F)


YrsPort=1989:2020

 #dput(sort(unique(ObserverData$NAFAREA_ID)))
 nafo3 = c( "3N", "3NA", "3NB", "3NC", "3ND", "3NE", "3NF", "3O", "3OA", "3OC", "3OD", "3OE", "3PN", "3PS", "3U")
 nafo4 = c("4V","4VB", "4VC", "4VN", "4VNW", "4VS", "4VSB", "4VSC","4VSW","4VW","4VX", "4W", "4WD", "4WE", "4WF", "4WG", "4WH", "4WJ", "4WK", "4WL", "4WM","4WX","4XMNOPQ", "4X", "4XL", "4XM", "4XN", "4XO", "4XP", "4XQ", "4XR", "4XU", "4XS", "4XRS", "5YB", "5ZJ", "5ZM", "5ZN","5ZEJ")
 


  portNAFO4OT1=constructLF(LFdata=subset(PortData,NAFO%in%nafo4&GEAR=="OT"),bins=bins, Yrs = YrsPort)
  BarPlotLF(portNAFO4OT1,yrs=YrsPort,rel=F,filen=file.path(wd,"figures","portNAFO4OT"),rows=16,graphic='png',ax=c(rep(2,16),rep(4,16)))

  portNAFO4LL1=constructLF(LFdata=subset(PortData,NAFO%in%nafo4&GEAR=="LL"),bins=bins, Yrs = YrsPort)
  BarPlotLF(portNAFO4LL1,yrs=YrsPort,rel=F,filen=file.path(wd,"figures","portNAFO4LL"),rows=16,graphic='png',ax=c(rep(2,16),rep(4,16)))


 

```

## Observer Data
```{r}
bins=seq(0,260,5)

ObserverData<-ObsData(datadir=datadir,by.sex=T, bins=bins)

 Yrs1 = 1991:2020 
 Yrs2 = 2006:2020 
 #dput(sort(unique(ObserverData$NAFO)))
 nafo3 = c( "3N", "3NA", "3NB", "3NC", "3ND", "3NE", "3NF", "3O", "3OA", "3OC", "3OD", "3OE", "3PN", "3PS", "3U")
 nafo4 = c("4V","4VB", "4VC", "4VN", "4VNW", "4VS", "4VSB", "4VSC","4VSW","4VW","4VX", "4W", "4WD", "4WE", "4WF", "4WG", "4WH", "4WJ", "4WK", "4WL", "4WM","4WX","4XMNOPQ", "4X", "4XL", "4XM", "4XN", "4XO", "4XP", "4XQ", "4XR", "4XU", "4XS", "4XRS", "5YB", "5ZJ", "5ZM", "5ZN","5ZEJ")
 

  obsNAFO3OT1=constructLF(LFdata=subset(ObserverData,NAFO%in%nafo3&GEAR=="OT"),bins=bins, Yrs = Yrs1)
  BarPlotLF(obsNAFO3OT1,yrs=Yrs1,rel=F,filen=file.path(wd,"figures","obsNAFO3OT1"))

  obsNAFO3OT2=constructLF(LFdata=subset(ObserverData,NAFO%in%nafo3&GEAR=="OT"),bins=bins, Yrs = Yrs2)
  BarPlotLF(obsNAFO3OT2,yrs=Yrs2,rel=F,filen=file.path(wd,"figures","obsNAFO3OT2"))

  obsNAFO3LL1=constructLF(LFdata=subset(ObserverData,NAFO%in%nafo3&GEAR=="LL"),bins=bins, Yrs = Yrs1)
  BarPlotLF(obsNAFO3LL1,yrs=Yrs1,rel=F,filen=file.path(wd,"figures","obsNAFO3LL1"))

  obsNAFO3LL2=constructLF(LFdata=subset(ObserverData,NAFO%in%nafo3&GEAR=="LL"),bins=bins, Yrs = Yrs2)
  BarPlotLF(obsNAFO3LL2,yrs=Yrs2,rel=F,filen=file.path(wd,"figures","obsNAFO3LL2"))

  obsNAFO4OT1=constructLF(LFdata=subset(ObserverData,NAFO%in%nafo4&GEAR=="OT"),bins=bins, Yrs = Yrs1)
  BarPlotLF(obsNAFO4OT1,yrs=Yrs1,rel=F,filen=file.path(wd,"figures","obsNAFO4OT1"))

  obsNAFO4OT2=constructLF(LFdata=subset(ObserverData,NAFO%in%nafo4&GEAR=="OT"),bins=bins, Yrs = Yrs2)
  BarPlotLF(obsNAFO4OT2,yrs=Yrs2,rel=F,filen=file.path(wd,"figures","obsNAFO4OT2"))

  obsNAFO4LL1=constructLF(LFdata=subset(ObserverData,NAFO%in%nafo4&GEAR=="LL"),bins=bins, Yrs = Yrs1)
  BarPlotLF(obsNAFO4LL1,yrs=Yrs1,rel=F,filen=file.path(wd,"figures","obsNAFO4LL1"))

  obsNAFO4LL2=constructLF(LFdata=subset(ObserverData,NAFO%in%nafo4&GEAR=="LL"),bins=bins, Yrs = Yrs2)
  BarPlotLF(obsNAFO4LL2,yrs=Yrs2,rel=F,filen=file.path(wd,"figures","obsNAFO4LL2"))
  
  
  
   
 nafo4X = c( "4X", "4XL", "4XM", "4XN", "4XO", "4XP", "4XQ", "4XR", "4XS", "5YB", "5ZJ", "5ZM", "5ZN")

  
  
 nafo4VW = c("4VB", "4VC", "4VN", "4VS", "4W", "4WD", "4WE", "4WF", "4WG", "4WH", "4WJ", "4WK", "4WL", "4WM")

  
  
  obsNAFO4XLL2=constructLF(LFdata=subset(ObserverData,NAFO%in%nafo4X&GEAR=="LL"),bins=bins, Yrs = Yrs2)
  BarPlotLF(obsNAFO4XLL2,yrs=Yrs2,rel=T,filen=file.path(wd,"figures","obsNAFO4XLL2"))

  obsNAFO4VWLL2=constructLF(LFdata=subset(ObserverData,NAFO%in%nafo4VW&GEAR=="LL"),bins=bins, Yrs = Yrs2)
  BarPlotLF(obsNAFO4VWLL2,yrs=Yrs2,rel=T,filen=file.path(wd,"figures","obsNAFO4VWLL2"))

  ### WTF lots of baby halibut????3
  look<-subset(ObserverData,NAFO%in%nafo4&GEAR=="OT"&YEAR==2020&L30>0) # estimated # caught is 2
  Mar.datawrangling::get_data_custom(schema="observer", data.dir = datadir, tables = c("ISFISHLENGTHS","ISSAMPLES"))
  subset(ISSAMPLES,CATCH_ID==101925022)
  subset(ISFISHLENGTHS,SMPL_ID==100452896) # there's the length frequency
  Mar.datawrangling::get_data(db='isdb',data.dir=datadir)
  subset(ISCATCHES,CATCH_ID==101925022) # nothing
  subset(ISFISHSETS,FISHSET_ID==100336369)
  subset(ISTRIPS,TRIP_ID==100057153)

```

# observer data plots for document
```{r}
bins=seq(0,260,5)
ObserverData<-ObsData(datadir=datadir,by.sex=T, bins=bins)
 
ObserverDataTable<-filter(ObserverData,!is.na(NUM_MEASURED)) %>% 
  group_by(TRIPCD_ID,YEAR)%>% 
  distinct(FISHSET_ID)%>%
  count()%>%
  pivot_wider(names_from = TRIPCD_ID,values_from = n)%>%
  arrange(YEAR)

ObserverDataTable[is.na(ObserverDataTable)]<-0

write.csv(ObserverDataTable,"ObserverDataTable.csv",row.names=F)




 YrsObs = 1991:2020 
 #dput(sort(unique(ObserverData$NAFO)))
 nafo3 = c( "3N", "3NA", "3NB", "3NC", "3ND", "3NE", "3NF", "3O", "3OA", "3OC", "3OD", "3OE", "3PN", "3PS", "3U")
 nafo4 = c("4V","4VB", "4VC", "4VN", "4VNW", "4VS", "4VSB", "4VSC","4VSW","4VW","4VX", "4W", "4WD", "4WE", "4WF", "4WG", "4WH", "4WJ", "4WK", "4WL", "4WM","4WX","4XMNOPQ", "4X", "4XL", "4XM", "4XN", "4XO", "4XP", "4XQ", "4XR", "4XU", "4XS", "4XRS", "5YB", "5ZJ", "5ZM", "5ZN","5ZEJ")
 

  obsNAFO3OT1=constructLF(LFdata=subset(ObserverData,NAFO%in%nafo3&GEAR=="OT"),bins=bins, Yrs = YrsObs)
  BarPlotLF(obsNAFO3OT1,yrs=YrsObs,rel=F,filen=file.path(wd,"figures","obsNAFO3OT"),rows=15,graphic='png',ax=c(rep(2,15),rep(4,15)))

  obsNAFO3LL1=constructLF(LFdata=subset(ObserverData,NAFO%in%nafo3&GEAR=="LL"),bins=bins, Yrs = YrsObs)
  BarPlotLF(obsNAFO3LL1,yrs=YrsObs,rel=F,filen=file.path(wd,"figures","obsNAFO3LL"),rows=15,graphic='png',ax=c(rep(2,15),rep(4,15)))

  obsNAFO4OT1=constructLF(LFdata=subset(ObserverData,NAFO%in%nafo4&GEAR=="OT"),bins=bins, Yrs = YrsObs)
  BarPlotLF(obsNAFO4OT1,yrs=YrsObs,rel=F,filen=file.path(wd,"figures","obsNAFO4OT"),rows=15,graphic='png',ax=c(rep(2,15),rep(4,15)))

  obsNAFO4LL1=constructLF(LFdata=subset(ObserverData,NAFO%in%nafo4&GEAR=="LL"),bins=bins, Yrs = YrsObs)
  BarPlotLF(obsNAFO4LL1,yrs=YrsObs,rel=F,filen=file.path(wd,"figures","obsNAFO4LL"),rows=15,graphic='png',ax=c(rep(2,15),rep(4,15)))

```


# map location of data
```{r}

xl=c(-69,-58)
yl=c(41,46)

obs4x20<-subset(ObserverData,NAFAREA_ID%in%nafo4X&GEAR=="LL"&YEAR==2020)
obs4x19<-subset(ObserverData,NAFAREA_ID%in%nafo4X&GEAR=="LL"&YEAR==2019)

obs4x19<-subset(ObserverData,NAFAREA_ID=='5ZJ'&GEAR=="LL"&YEAR==2019)
h<-constructLF(obs4x19,bins=bins,Yrs=2019)

obs20fundianchannel<-subset(ObserverData,TRIP_ID%in%c(100057412,100057428  ))
f<-constructLF(obs20fundianchannel,bins=bins,Yrs=2020)

obs20northgeorges<-subset(ObserverData,TRIP_ID%in%c(100057154, 10005735, 100057408, 100057487, 100057489))
g<-constructLF(obs20northgeorges,bins=bins,Yrs=2020)


g
SpatialHub::bioMap(xlim = xl, ylim = yl, mapRes="MR")
with(obs4x20,points(LONGITUDE,LATITUDE,pch=16,cex=1,col=rgb(1,0,0,0.2)))
with(obs4x19,points(LONGITUDE,LATITUDE,pch=16,cex=1,col=rgb(0,0,1,0.2)))


SpatialHub::bioMap(xlim = xl, ylim = yl, mapRes="MR")
with(subset(ObserverData,NAFAREA_ID=='5ZJ'&GEAR=="LL"&YEAR==2020),points(LONGITUDE,LATITUDE,pch=16,cex=1,col=rgb(1,0,0,0.2)))

```











## from Mar.fleets
#```{r}
yrs = 2002:2020
fleets <- list()
        for (i in 1:length(yrs)){
          cat(yrs[i])
          thisyear = fleet_halibut(useLocal=F, data.dir=datadir, dateStart=paste0(yrs[i],'-01-01'), dateEnd=paste0(yrs[i],'-12-31'), keepSurveyTrips=T, oracle.username=uid, oracle.password=pwd, oracle.dsn='ptran',usepkg="roracle")
          fleets[[i]]<-thisyear$marf$MARF_SETS
          rm(thisyear)
        }
fleets[[1]]$CALC_AREA<-NA
marf<-do.call("rbind",fleets)


catch.data <- na.omit(with(subset(marf,year(DATE_FISHED)==2020),data.frame(LOG_EFRT_STD_INFO_ID,LONGITUDE,LATITUDE,RND_WEIGHT_KGS)))
catch.data$LOG_EFRT_STD_INFO_ID <- 1:nrow(catch.data)
summary(catch.data)

#```

