---
title: "Update Advice - Harper Test"
author: "Danni"
date: "2022-10-13"
output: html_document
---


```{r setup, include=FALSE}
library(ROracle)
library(devtools)
library(tidyverse)
require(PBSmapping)
#library(pipeR)
#install.packages("remotes")
#remotes::install_github("mlesnoff/rchemo")
#library(rchemo)
source(file.path(getwd(), "directories.r"))
source(file.path(wd, "passwords.r")) 
#source(file.path("C:/Users/harperd/Documents/Halibut/GitDataInputs/DataInputs_2023/DataInputs/passwords.r")) 

#datadir="C:/Users/harperd/Documents/Halibut/RDataVault"
#wd="C:/Users/harperd/Documents/Halibut/GitDataInputs/DataInputs_2023/DataInputs"

ds_all <- Mar.datawrangling::load_datasources()

devtools::load_all(".")

bins=seq(0,260,5)
# labelled by upper bound


Current.Year <- 2024


```


## Update data from database 

```{r }


# Update data from database 

Mar.datawrangling::get_data(db='isdb',data.dir=datadir,fn.oracle.username = uid, fn.oracle.password = pwd, fn.oracle.dsn = 'ptran',usepkg='roracle',force.extract=T)
# open and check setprofile_wide to confirm that you have year, latitude & longitude

RVdata<-get4VWXRV(uid, pwd, use.local=F,datadir=datadir)

```



#NS RV survey

```{r}

#RVdata<-get4VWXRV(uid, pwd, use.local=F,datadir=datadir)
RVdata<-get4VWXRV(uid, pwd, use.local=T,datadir=datadir) #pulls RV data from data vault on computer
RVinputs<-prepRVdata(RVdata,bins=bins,years=1970:Current.Year, raw=T) #preps data for each year 

RVinputs$NSRV_unsexed<-RVinputs$NSRV_combined-(RVinputs$NSRV_males+RVinputs$NSRV_females) # unsexed
NSRV_Index<-RVinputs$Index
NSRV_Lengths<-list(NSRV_males=RVinputs$NSRV_males,
                   NSRV_females=RVinputs$NSRV_females,
                   NSRV_combined=RVinputs$NSRV_combined,
                   NSRV_unsexed=RVinputs$NSRV_unsexed)
# 
#NSRV_Index$NPT[NSRV_Index$year==2021]<-NA

BubblePlotLF(NSRV_Lengths,bins=bins,yrs=1970:2023,path=file.path(wd,'figures'),filen='NLRVSurvey',inches=0.3)

nsrvl<-revNAL(NSRV_Lengths[-3])
YrsObs = 1994:2023 
BarPlotLF(nsrvl[which(names(nsrvl)%in%YrsObs)],yrs=YrsObs,rel=F,filen=file.path(wd,"figures","NSRVbarplot"),rows=15,graphic='png',ax=c(rep(2,15),rep(4,15)))

## data for comparing 4X and 4VSW

strata4x<-c(470:495)
RVinputs4X<-prepRVdata(RVdata,strat=strata4x,bins=bins,years=1970:Current.Year, raw=T)

strata4VSW<-c(440:469)
RVinputs4VSW<-prepRVdata(RVdata,strat=strata4VSW,bins=bins,years=c(1970:2017, 2019:2020,2023), raw=T)


NSRV_Lengths4X<-list(NSRV_males=RVinputs4X$NSRV_males,
                   NSRV_females=RVinputs4X$NSRV_females,
                   NSRV_combined=RVinputs4X$NSRV_combined)

NSRV_Lengths4VSW<-list(NSRV_males=RVinputs4VSW$NSRV_males,
                   NSRV_females=RVinputs4VSW$NSRV_females,
                   NSRV_combined=RVinputs4VSW$NSRV_combined)

```
# compare 4X and 4VSW in RV survey
```{r}
png(file.path(wd,'figures',"RVsurvey4Xvs4VSW.png"), width =8, height = 6,units='in',pointsize=12, res=300,type='cairo')
par(mfcol=c(2,1), mar = c(0,4,0,0.5), omi = c(0.5, 0.25, 0.5, 0.25),las=1)

plot(B/tunits~year,RVinputs4X$Index, ylab="Kg / tow", xlab = '', ylim=c(0,5), type='o',pch=16,xlim=c(1970,Current.Year),xaxt='n')
lines(B/tunits~year,RVinputs4VSW$Index,type='o',pch=15,col=2)


plot(NPT~year,RVinputs4X$Index, ylab="Number / tow", xlab = '', ylim=c(0,max(NPT,na.rm=T)*1.2), type='o',pch=16,xlim=c(1970,Current.Year))
lines(NPT~year,RVinputs4VSW$Index,type='o',pch=15,col=2)

dev.off()


```

#Halibut Longline Survey

```{r}

  HSFixed_Index<-FixedSurveyIndex(datadir=datadir,yrs=1998:2022)
  HSRandom_Index_WZS<-StratifiedRandomSurveyIndex(datadir=datadir,yrs=2017:Current.Year, use.calc.weight=F)
  HSRandom_Index_WOZS<-StratifiedRandomSurveyIndex(datadir=datadir,yrs=2017:Current.Year, ZeroStrata=F) 
  ##if looking for findPolys try require(PBSmapping)
  
  with(subset(HSRandom_Index_WZS,Year==Current.Year),c(mean3_biomass,mean3_biomass-KgPKHse*1.96,mean3_biomass+KgPKHse*1.96))/2
  with(subset(HSRandom_Index_WZS,Year==Current.Year),c(KgPKH,KgPKH-KgPKHse*1.96,KgPKH+KgPKHse*1.96))/2
  
  
```

# Harvest Control Rule
```{r}
value <- HSRandom_Index_WZS$mean3_biomass[HSRandom_Index_WZS$Year==Current.Year] 
HCR(value)


#HCR(value, q=0.0019) # when using calculated weight...but HCR function uses q = 0.002

```


## plot survey indices SSR version
```{r}
#RVinputs$Index


png(file.path(wd,'figures',"ModelindicesSSR_WZS_wo2021.png"), width =8, height = 6,units='in',pointsize=12, res=300,type='cairo')
par(mfcol=c(2,1), mar = c(0,4,0,0.5), omi = c(0.5, 0.25, 0.5, 0.25),las=1)

plot(NPT~year,NSRV_Index_pre2021, ylab="Number / tow", xlab = '', ylim=c(0,max(NPT,na.rm=T)*1.2), type='o',pch=16,xlim=c(1970,Current.Year),xaxt='n')
axis(1,lab=F)
lines(NPT~year,NSRV_Index_post2021, ylab="Number / tow", xlab = '', ylim=c(0,max(NPT,na.rm=T)*1.2), type='o',pch=16,xlim=c(1970,Current.Year),xaxt='n')
abline(h = 0.3, lty = "dashed")
points(c(2018,2022),NSRV_Index$NPT[NSRV_Index$year%in%c(2018,2022)],col='red',cex=2)
legend(x="topleft",   # Coordinates (x also accepts keywords)
       legend = c("RV Survey"),
       text.col = "black",
       bty = "n",

)

plot(KgPKH~Year,HSFixed_Index,type='o',pch=16,col='grey40',ylim=c(0,255),xlim=c(1970,Current.Year), ylab="Kg / 1,000 hooks")
lines(mean3_biomass~Year,HSFixed_Index, lwd = 3, col = 'grey40')
lines(KgPKH~Year,HSRandom_Index_WZS,type='o',pch=16,col='blue')
#lines(KgPKH+KgPKHse~Year,HSRandom_Index,lty=2,col='blue')
#lines(KgPKH-KgPKHse~Year,HSRandom_Index,lty=2,col='blue')
lines(mean3_biomass~Year,HSRandom_Index_WZS,lwd=3, col = "blue")
legend(x="topleft", inset = 0.05,   # Coordinates (x also accepts keywords)
       title = "Halibut Longline Survey",
       title.col = "black",
       legend = c("Fixed Station", "Stratified Random"), # Vector with the name of each group
       text.col = c("grey40", "blue"),
       bty = "n",

)

dev.off()


##checking out #/tow and biomass/tow
RVinputs <- data.frame(RVinputs)
RVinputs <- RVinputs %>% mutate(Index.BPT = (Index.B/Index.tunits))

png(file.path(wd,'figures',"RVsurvey_AbundanceVSBiomass.png"), width =8, height = 6,units='in',pointsize=12, res=300,type='cairo')

par(mfcol=c(2,1), mar = c(0,4,0,0.5), omi = c(0.5, 0.25, 0.5, 0.25),las=1)

plot(NPT~year,NSRV_Index, ylab="Number / tow", xlab = '', ylim=c(0,max(NPT,na.rm=T)*1.2), type='o',pch=16,xlim=c(1970,Current.Year),xaxt='n')
axis(1,lab=F)
points(c(2018,2021,2022),NSRV_Index$NPT[NSRV_Index$year%in%c(2018,2021,2022)],col='red',cex=2)

plot(Index.BPT~Index.year,RVinputs, ylab="Biomass / tow", xlab = '', ylim=c(0,max(Index.BPT,na.rm=T)*1.2), type='o',pch=16,xlim=c(1970,Current.Year))
axis(1,lab=F)
points(c(2018,2021,2022),RVinputs$Index.BPT[RVinputs$Index.year%in%c(2018,2021,2022)],col='red',cex=2)

dev.off()
```

# Comparing calculated weight vs estimated weight
```{r}
 HSRandom_Index<-StratifiedRandomSurveyIndex(datadir=datadir,yrs=2017:Current.Year,use.calc.weight = F)
 HSRandom_Index2<-StratifiedRandomSurveyIndex(datadir=datadir,yrs=2017:Current.Year,use.calc.weight = T)

```

```{r}
  HSRandom_Index<-StratifiedRandomSurveyIndex(datadir=datadir,yrs=2017:Current.Year,use.calc.weight = F)
  HSRandom_Index1<-StratifiedRandomSurveyIndex(datadir=datadir,yrs=2017:Current.Year,use.calc.weight = F)

```

##some playing around...trying to figure some stuff out
```{r}
 #index using calculated weight vs estimated weight (I think we use estimated weight)
 plot(KgPKH~Year,HSRandom_Index,ylim=c(0,150),type='l',lwd=2)
 with(HSRandom_Index,lines(Year,mean3_biomass,lty=2))
 with(HSRandom_Index2,lines(Year,KgPKH,lwd=2,col=2))
 with(HSRandom_Index2,lines(Year,mean3_biomass,lty=2,col=2))
 
 # Q-adjusted (using estimated weight) - plot of biomass rather than the index
 plot(KgPKH/0.002~Year,HSRandom_Index,ylim=c(0,60000),type='l',lwd=2,ylab="Biomass")
 with(HSRandom_Index,lines(Year,mean3_biomass/0.002,lty=2))
 with(HSRandom_Index2,lines(Year,KgPKH/0.0019,lwd=2,col=2))
 with(HSRandom_Index2,lines(Year,mean3_biomass/0.0019,lty=2,col=2))
 
 # Q-adjusted (adjusting Fixed with SR)
 plot(KgPKH/4.6~Year,HSFixed_Index,ylim=c(0,60),type='l',lwd=2,ylab="Biomas (kt)",las=1)
 #with(HSFixed_Index,lines(Year,mean3_biomass/0.0046,lty=2))
 with(HSRandom_Index,lines(Year,KgPKH/2,lwd=2,col=2))
# with(HSRandom_Index,lines(Year,mean3_biomass/0.002,lty=2,col=2))
 
 
 # comparing with and without zero strata
 plot(KgPKH~Year,HSRandom_Index_WZS,ylim=c(0,150),type='l',lwd=2)
 with(HSRandom_Index_WZS,lines(Year,mean3_biomass,lty=2)) #black line
 with(HSRandom_Index_WOZS,lines(Year,KgPKH,lwd=2,col=2)) #red line
 with(HSRandom_Index_WOZS,lines(Year,mean3_biomass,lty=2,col=2)) 
 
 # comparing with and without 5.2 and 5.3 combined...not sure this really works how i think it does...but lines ALMOST overlap...red line slightly more conservative
  HSRandom_Index_WOZS_5blend<-StratifiedRandomSurveyIndex(datadir=datadir,yrs=2017:Current.Year, ZeroStrata=F, restratify = T)
  HSRandom_Index_WOZS_test<-StratifiedRandomSurveyIndex(datadir=datadir,yrs=2017:Current.Year, ZeroStrata=F, restratify = F)
  HSRandom_Index5<-StratifiedRandomSurveyIndex(datadir=datadir,yrs=2017:Current.Year, ZeroStrata=T, restratify=F, combine.strata=52:53)
  HSRandom_Index_WZS_5_restrat<-StratifiedRandomSurveyIndex(datadir=datadir,yrs=2017:Current.Year, ZeroStrata=T, restratify=T, combine.strata=52:53)
 
png(file.path(wd,'figures',"LLindex_WZS_5combinedAll.png"), width =8, height = 6,units='in',pointsize=12, res=300,type='cairo')
 plot(KgPKH~Year,HSRandom_Index_WZS,ylim=c(0,150),type='l',lwd=2)
 with(HSRandom_Index_WZS,lines(Year,mean3_biomass,lty=2)) #black line
 with(HSRandom_Index5,lines(Year,KgPKH,lwd=2,col=2)) #red line
 with(HSRandom_Index5,lines(Year,mean3_biomass,lty=2,col=2))
 dev.off()
 
 ##for Bruce on zero strata
 png(file.path(wd,'figures',"LLindex_overtime_WandWOZS.png"), width =8, height = 6,units='in',pointsize=12, res=300,type='cairo')
 plot(KgPKH~Year,HSRandom_Index_WOZS,ylim=c(0,150),type='l',lwd=2)
 with(HSRandom_Index_WZS_5_restrat,lines(Year,KgPKH,lwd=2,col=2)) #red line
 dev.off()
```

##looking at #sets/area for each strata under different strata schemes and over time (to investigate the impact of zero strata)
```{r}
RSdata <- RandomSurveyData(sp=30, datadir, add.gear=F, add.LF=F, hook.data=F)

load(file.path(datadir, "Survey","SurveyStrata.RData"))
StrataAreas1 <-data.frame(StrataAreas)
surveyStrataPolyLL1 <- data.frame(surveyStrataPolyLL)

load(file.path(datadir, "Survey","SurveyStrata2022.RData"))
StrataAreas2022 <-data.frame(StrataAreas)
surveyStrataPolyLL2022 <- data.frame(surveyStrataPolyLL)

load(file.path(datadir, "Survey","SurveyStrata2023.RData"))
StrataAreas2023 <-data.frame(StrataAreas)
surveyStrataPolyLL2023 <- data.frame(surveyStrataPolyLL)

load(file.path(datadir, "Survey","SurveyStrata2023NoBO.RData"))
StrataAreas2023WOZS <-data.frame(StrataAreas)
surveyStrataPolyLL2023WOZS <- data.frame(surveyStrataPolyLL)

RSdata_original <- reStratify(sets = RSdata, strata = surveyStrataPolyLL1)
RSdata_2022 <- reStratify(sets = RSdata, strata = surveyStrataPolyLL2022)
RSdata_2023 <- reStratify(sets = RSdata, strata = surveyStrataPolyLL2023)
RSdata_2023WOZS <- reStratify(sets = RSdata, strata = surveyStrataPolyLL2023WOZS)

#looking using original strata scheme
RSdata_original_summary <- RSdata_original %>%
  group_by(YEAR, PID) %>%
  summarize(n = n())

RSdata_originalWarea <- RSdata_original_summary %>% left_join(StrataAreas1, by = "PID") %>% mutate(prop = signif(n/area)*10000, digits = 3)
RSdata_originalWarea_wide <- RSdata_originalWarea %>% select(PID, YEAR, prop) %>% spread(YEAR, prop)

#looking using 2022 strata scheme
RSdata_2022_summary <- RSdata_2022 %>%
  group_by(YEAR, PID) %>%
  summarize(n = n())

RSdata_originalWarea <- RSdata_original_summary %>% left_join(StrataAreas1, by = "PID") %>% mutate(prop = signif(n/area)*10000, digits = 3)
RSdata_originalWarea_wide <- RSdata_originalWarea %>% select(PID, YEAR, prop) %>% spread(YEAR, prop)


##density for new zero strata allocation design
load(file.path(datadir, "Survey","SurveyStrataOneZeroStrata.RData"))
StrataAreas_ZS2024 <-data.frame(StrataAreas)
StrataAreas_ZS2024$PID <- as.character(StrataAreas_ZS2024$PID)
surveyStrataPolyLL2024 <- data.frame(surveyStrataPolyLL)

allocation2024 <- read.csv("C:/Users/harperd/Documents/Halibut/RDataVault/Survey/allocation2024.csv")

allocation2024$SID <- as.character(gsub("\\.", "", (allocation2024$SID)))
allocation2024[18, "SID"] <- 60
allocation2024 <- allocation2024 %>% rename(PID = SID) %>% select(-Area)

allocation2024 <- allocation2024 %>% left_join(StrataAreas_ZS2024, by = "PID")

allocationdensity2024 <- allocation2024 %>% mutate(density = signif((Sets/area)*100, digits = 3))

allocationdensity2024_WOZS <- allocationdensity2024 %>% filter(PID != 60)
mean(allocationdensity2024_WOZS$density)
```
Survey update document
```{r}
ComIndex<-CommercialIndexData(datadir=datadir,bins=bins,by.sex=F,add.portsampling=F,add.LF = F)
FixedSurvey<-FixedSurveyData(datadir=datadir,by.sex=F, bins=bins, add.LF = F)
RandomSurvey<-RandomSurveyData(sp=30,datadir=datadir,add.gear=F,add.LF=F,bins=bins,by.sex=F,hook.data=F)

png(file.path(wd,'figures',"CI.png"), width =6, height = 4,units='in',pointsize=9, res=300,type='cairo')
   # par(mfcol=c(rows,ceiling(length(yrs)/rows)), mar = c(0,2,0,0.5), omi = c(0.85, 0.75, 0.75, 0.5))

ComIndex$KgPKH<-ComIndex$EST_COMBINED_WT/ComIndex$NUM_HOOK_HAUL*1000
dot <- ComIndex %>% group_by(YEAR) %>% summarize(mean = mean(KgPKH, na.rm = TRUE))
CI<-with(ComIndex,tapply(KgPKH,YEAR, mean, na.rm=T))
plot(as.numeric(names(CI)),CI,xlab = '',ylab = 'Kg / 1,000 Hooks',ylim=c(0,500),type='b',pch=16)
lines(KgPKH~Year,HSRandom_Index_WZS,type='o',pch=16,col='blue')
points(2023,dot$mean[dot$YEAR==2023],col='red',cex=2)



FixedSurvey$KgPKH<-FixedSurvey$EST_COMBINED_WT/FixedSurvey$NUM_HOOK_HAUL*1000
FS<-with(FixedSurvey,tapply(KgPKH,YEAR, mean, na.rm=T))
lines(as.numeric(names(FS)),FS,col='grey',lty=2)
legend(x="topleft", inset = 0.05,   # Coordinates (x also accepts keywords)
       title.col = "black",
       legend = c("Biological Sampling Program","Fixed Station", "Stratified Random"), # Vector with the name of each group
       text.col = c("black","grey40", "blue"),
       bty = "n",

)




dev.off()
```

---
title: "Update Advice - Harper Test"
author: "Danni"
date: "2022-10-13"
output: html_document
---


```{r setup, include=FALSE}
library(ROracle)
library(devtools)
library(tidyverse)
require(PBSmapping)
library(pipeR)
#source(file.path(getwd(), "directories.r"))
#source(file.path(wd, "passwords.r")) 
source(file.path("C:/Users/harperd/Documents/Halibut/GitDataInputs/DataInputs_2023/DataInputs/passwords.r")) 

datadir="C:/Users/harperd/Documents/Halibut/RDataVault"
wd="C:/Users/harperd/Documents/Halibut/GitDataInputs/DataInputs_2023/DataInputs"

ds_all <- Mar.datawrangling::load_datasources()

devtools::load_all(".")

bins=seq(0,260,5)
# labelled by upper bound


Current.Year <- 2023


```

Figures for the 2023 Stock Status Update presentation
```{r}
##map of 2023 strata

load(file.path(datadir,"SurveyStrata2022.RData"))
BO1<-data.frame(PID=1,POS=1:6,X=c(-47,-55,-55,-56.15283,-56.15283,-47),Y=c(45.5,45.5,46.25,46.25,48,48))

BO1<-joinPolys(BO1,surveyPolyLL,operation="INT")


BO2<-data.frame(PID=1,POS=1:4,X=c(-68,-68,-67,-67),Y=c(41,46,46,41))

BO2<-joinPolys(BO2,surveyPolyLL,operation="INT")
 
png(file.path("C:/Users/harperd/Documents/Halibut/GitDataInputs/DataInputs_2023/DataInputs/figures/start2023strata_forDoc.png"), width =8, height = 6,units='in',pointsize=12, res=300,type='cairo')                            
xl=c(-68,-47)
yl=c(40,48)
bioMap(xlim = xl, ylim = yl,mapRes="MR",poly.lst = list(surveyStrataPolyLL,props), plot.rivers=F,isobaths = NULL)
addPolys(BO1,density=20,col='black')
# addPolys(BO2,density=20,col='black')
 #points(lat.DecDeg~lon.DecDeg,assignedstations2023,pch=21,bg = "yellow",cex=0.5)
 #points(lat.DecDeg_Assigned~lon.DecDeg_Assigned,completed2023,pch=21, bg = "black",cex=0.5)
plot(NAFO_test,add=T)
#text(coordinates(NAFO_test)[,1], coordinates(NAFO_test)[,2], NAFO_test$ZONE)
text(x = -56.5, y = 43, labels = "4Vs")
text(x = -59.55, y = 46.37, labels = "4Vn")
text(x = -61, y = 42, labels = "4W")
text(x = -64.5, y = 41.5, labels = "4X")
text(x = -66.25, y = 41.9, labels = "5Zc")
text(x = -48.75, y = 42.5, labels = "3N")
text(x = -53.2, y = 43.7, labels = "3O")
text(x = -55.5, y = 45.9, labels = "3Ps")
text(x = -58.7, y = 47.35, labels = "3Pn")

dev.off()

NAFO_test <- subset(NAFO, ZONE %in% c("3N", "3O", "3Ps", "3Pn","4Vn", "4Vs", "4W", "4X", "5Ze"))

  ##without zero strata

png(file.path("C:/Users/harperd/Documents/Halibut/GitDataInputs/DataInputs_2023/DataInputs/figures/2022strata.png"), width =8, height = 6,units='in',pointsize=12, res=300,type='cairo')                            
xl=c(-69,-47)
yl=c(40,48)
bioMap(xlim = xl, ylim = yl,mapRes="MR",poly.lst = list(surveyStrataPolyLL,props), plot.rivers=F,isobaths = NULL)

dev.off()

  #insets to highlight channel strata (from 2022) and zero strata for 2023
ChannelPoly <- joinPolys(subset(surveyStrataPolyLL,PID==33),subset(surveyStrataPolyLL,PID==43),operation = "UNION")
ChannelPoly <- subset(ChannelPoly,SID==16)

png(file.path("C:/Users/harperd/Documents/Halibut/GitDataInputs/DataInputs_2023/DataInputs/figures/zoomchannelstrata2022.png"), width =8, height = 6,units='in',pointsize=12, res=300,type='cairo') 
 bioMap(xlim=c(-61,-54),ylim=c(44,48),poly.lst = list(surveyStrataPolyLL,props),plot.rivers=F,isobaths = NULL,mapRes="MR",LT=T)
    plot(NAFO,add=T)
    text(coordinates(NAFO)[,1], coordinates(NAFO)[,2],NAFO$ZONE)
 addPolys(ChannelPoly,col=rgb(0,0,1,0.1))
dev.off()

#add on 200 stations assigned
assignedstations2023 <- read.csv("C:/Users/harperd/Documents/Halibut/Survey/HalSurvey2023/Survey Audit 2023/2023 Halibut Longline Survey Stations - w assignmentsJuly18.csv")

png(file.path("C:/Users/harperd/Documents/Halibut/GitDataInputs/DataInputs_2023/DataInputs/figures/start2023stations.png"), width =8, height = 6,units='in',pointsize=12, res=300,type='cairo') 
xl=c(-69,-47)
yl=c(40,48)
bioMap(xlim = xl, ylim = yl,mapRes="MR",poly.lst = list(surveyStrataPolyLL,props), plot.rivers=F,isobaths = NULL)
addPolys(BO1,density=20,col='black')
addPolys(BO2,density=20,col='black')
 points(lat.DecDeg~lon.DecDeg,assignedstations2023,pch=21,bg="yellow",cex=0.5)
dev.off()

#overlay completed 175 stations
completed2023 <- read.csv("C:/Users/harperd/Documents/Halibut/Survey/HalSurvey2023/Survey Audit 2023/survey_dat_edited.csv")

png(file.path("C:/Users/harperd/Documents/Halibut/GitDataInputs/DataInputs_2023/DataInputs/figures/completed2023stations.png"), width =8, height = 6,units='in',pointsize=12, res=300,type='cairo') 
xl=c(-69,-47)
yl=c(40,48)
bioMap(xlim = xl, ylim = yl,mapRes="MR",poly.lst = list(surveyStrataPolyLL,props), plot.rivers=F,isobaths = NULL)

addPolys(BO1,density=20,col='black')
addPolys(BO2,density=20,col='black')
 points(lat.DecDeg~lon.DecDeg,assignedstations2023,pch=21,bg = "yellow",cex=0.5)
 points(lat.DecDeg_Assigned~lon.DecDeg_Assigned,completed2023,pch=21, bg = "black",cex=0.5)
dev.off()

#zoom in on 5.2 and 5.3 area
png(file.path("C:/Users/harperd/Documents/Halibut/GitDataInputs/DataInputs_2023/DataInputs/figures/zoomstrata5.png"), width =8, height = 6,units='in',pointsize=12, res=300,type='cairo') 
xl=c(-55,-47)
yl=c(42.5,45)
bioMap(xlim = xl, ylim = yl,mapRes="MR",poly.lst = list(surveyStrataPolyLL,props), plot.rivers=F,isobaths = NULL)
 points(lat.DecDeg_Assigned~lon.DecDeg_Assigned,completed2023,pch=21, bg = "black")
addPolys(BO1,density=20,col='black') 
dev.off()

##new map with strata design for index calculation this year
load("C:/Users/harperd/Documents/Halibut/RDataVault/Survey/SurveyStrata2023.rdata")

png(file.path("C:/Users/harperd/Documents/Halibut/GitDataInputs/DataInputs_2023/DataInputs/figures/strata2023update.png"), width =8, height = 6,units='in',pointsize=12, res=300,type='cairo') 
xl=c(-69,-47)
yl=c(40,48)
bioMap(xlim = xl, ylim = yl,mapRes="MR",poly.lst = list(surveyStrataPolyLL,props), plot.rivers=F,isobaths = NULL)
 points(lat.DecDeg_Assigned~lon.DecDeg_Assigned,completed2023,pch=21,bg = "black",cex=0.5)
addPolys(BO1,density=20,col='black')
dev.off()
  
#inset zooming into single zero strata
xl=c(-57,-47)
yl=c(45,48)

png(file.path("C:/Users/harperd/Documents/Halibut/GitDataInputs/DataInputs_2023/DataInputs/figures/zoomnewZS.png"), width =8, height = 6,units='in',pointsize=12, res=300,type='cairo') 
bioMap(xlim = xl, ylim = yl,mapRes="MR",poly.lst = list(surveyStrataPolyLL,props), plot.rivers=F,isobaths = NULL)
 points(lat.DecDeg_Assigned~lon.DecDeg_Assigned,completed2023,pch=21, bg = "black")
addPolys(BO1,density=20,col='black')
dev.off()
  
#inset to show combined 5.2 and 5.3
png(file.path("C:/Users/harperd/Documents/Halibut/GitDataInputs/DataInputs_2023/DataInputs/figures/zoomstrata5combined.png"), width =8, height = 6,units='in',pointsize=12, res=300,type='cairo') 
xl=c(-55,-47)
yl=c(42.5,45)
bioMap(xlim = xl, ylim = yl,mapRes="MR",poly.lst = list(surveyStrataPolyLL,props), plot.rivers=F,isobaths = NULL)
 points(lat.DecDeg_Assigned~lon.DecDeg_Assigned,completed2023,pch=21, bg = "black")
addPolys(BO1,density=20,col='black')
dev.off()


```
Making 2024 map (for Nell's NAFO presentation)
```{r}
##map of 2023 strata

load(file.path(datadir,"SurveyStrata2022.RData"))
BO1<-data.frame(PID=1,POS=1:6,X=c(-47,-55,-55,-56.15283,-56.15283,-47),Y=c(45.5,45.5,46.25,46.25,48,48))

BO1<-joinPolys(BO1,surveyPolyLL,operation="INT")


BO2<-data.frame(PID=1,POS=1:4,X=c(-68,-68,-67,-67),Y=c(41,46,46,41))

BO2<-joinPolys(BO2,surveyPolyLL,operation="INT")
 
library(SpatialHub)
HalSurvey2024 <- read.csv("C:/Users/harperd/Documents/Halibut/Survey/HalSurvey2024/LongLineSurvey/data/StratifiedRandom_2024final_St38_May27.csv")

png(file.path("C:/Users/harperd/Documents/Halibut/GitDataInputs/DataInputs_2023/DataInputs/figures/2024survey_NellNAFOpresentation.png"), width =8, height = 6,units='in',pointsize=12, res=300,type='cairo')                            
xl=c(-68,-47)
yl=c(40,48)
bioMap(xlim = xl, ylim = yl,mapRes="MR",poly.lst = list(surveyStrataPolyLL,props), plot.rivers=F,isobaths = NULL)
addPolys(BO1,density=20,col='black')
sp::plot(NAFO,add=T)
points(lat.DecDeg~lon.DecDeg, HalSurvey2024,pch=21, bg="black",cex=0.5)
text(x = -56.5, y = 43, labels = "4Vs")
text(x = -59.55, y = 46.37, labels = "4Vn")
text(x = -61, y = 42, labels = "4W")
text(x = -64.5, y = 41.5, labels = "4X")
text(x = -66.25, y = 41.9, labels = "5Zc")
text(x = -48.75, y = 42.5, labels = "3N")
text(x = -53.2, y = 43.7, labels = "3O")
text(x = -55.5, y = 45.9, labels = "3Ps")
text(x = -58.7, y = 47.35, labels = "3Pn")
dev.off()
```


## Update data from database 

```{r }


# Update data from database 

Mar.datawrangling::get_data(db='isdb',data.dir=datadir,fn.oracle.username = uid, fn.oracle.password = pwd, fn.oracle.dsn = 'ptran',usepkg='roracle',force.extract=T)
# open and check setprofile_wide to confirm that you have year, latitude & longitude

RVdata<-get4VWXRV(uid, pwd, use.local=F,datadir=datadir)

```



#NS RV survey

```{r}

#RVdata<-get4VWXRV(uid, pwd, use.local=F,datadir=datadir)
RVdata<-get4VWXRV(uid, pwd, use.local=T,datadir=datadir) #pulls RV data from data vault on computer
RVinputs<-prepRVdata(RVdata,bins=bins,years=1970:Current.Year, raw=T) #preps data for each year 

RVinputs$NSRV_unsexed<-RVinputs$NSRV_combined-(RVinputs$NSRV_males+RVinputs$NSRV_females) # unsexed
NSRV_Index<-RVinputs$Index
NSRV_Lengths<-list(NSRV_males=RVinputs$NSRV_males,
                   NSRV_females=RVinputs$NSRV_females,
                   NSRV_combined=RVinputs$NSRV_combined,
                   NSRV_unsexed=RVinputs$NSRV_unsexed)
# 
#NSRV_Index$NPT[NSRV_Index$year==2021]<-NA

BubblePlotLF(NSRV_Lengths,bins=bins,yrs=1970:2023,path=file.path(wd,'figures'),filen='NLRVSurvey',inches=0.3)

nsrvl<-revNAL(NSRV_Lengths[-3])
YrsObs = 1994:2023 
BarPlotLF(nsrvl[which(names(nsrvl)%in%YrsObs)],yrs=YrsObs,rel=F,filen=file.path(wd,"figures","NSRVbarplot"),rows=15,graphic='png',ax=c(rep(2,15),rep(4,15)))

## data for comparing 4X and 4VSW

strata4x<-c(470:495)
RVinputs4X<-prepRVdata(RVdata,strat=strata4x,bins=bins,years=1970:Current.Year, raw=T)

strata4VSW<-c(440:469)
RVinputs4VSW<-prepRVdata(RVdata,strat=strata4VSW,bins=bins,years=c(1970:2017, 2019:2020,2023), raw=T)


NSRV_Lengths4X<-list(NSRV_males=RVinputs4X$NSRV_males,
                   NSRV_females=RVinputs4X$NSRV_females,
                   NSRV_combined=RVinputs4X$NSRV_combined)

NSRV_Lengths4VSW<-list(NSRV_males=RVinputs4VSW$NSRV_males,
                   NSRV_females=RVinputs4VSW$NSRV_females,
                   NSRV_combined=RVinputs4VSW$NSRV_combined)

```
# compare 4X and 4VSW in RV survey
```{r}
png(file.path(wd,'figures',"RVsurvey4Xvs4VSW.png"), width =8, height = 6,units='in',pointsize=12, res=300,type='cairo')
par(mfcol=c(2,1), mar = c(0,4,0,0.5), omi = c(0.5, 0.25, 0.5, 0.25),las=1)

plot(B/tunits~year,RVinputs4X$Index, ylab="Kg / tow", xlab = '', ylim=c(0,5), type='o',pch=16,xlim=c(1970,Current.Year),xaxt='n')
lines(B/tunits~year,RVinputs4VSW$Index,type='o',pch=15,col=2)


plot(NPT~year,RVinputs4X$Index, ylab="Number / tow", xlab = '', ylim=c(0,max(NPT,na.rm=T)*1.2), type='o',pch=16,xlim=c(1970,Current.Year))
lines(NPT~year,RVinputs4VSW$Index,type='o',pch=15,col=2)

dev.off()


```

#Halibut Longline Survey

```{r}

  HSFixed_Index<-FixedSurveyIndex(datadir=datadir,yrs=1998:2022)
  HSRandom_Index_WZS<-StratifiedRandomSurveyIndex(datadir=datadir,yrs=2017:Current.Year)
  HSRandom_Index_WOZS<-StratifiedRandomSurveyIndex(datadir=datadir,yrs=2017:Current.Year, ZeroStrata=F) 
  ##if looking for findPolys try require(PBSmapping)
  
  with(subset(HSRandom_Index_WZS,Year==Current.Year),c(mean3_biomass,mean3_biomass-KgPKHse*1.96,mean3_biomass+KgPKHse*1.96))/2
  with(subset(HSRandom_Index_WZS,Year==Current.Year),c(KgPKH,KgPKH-KgPKHse*1.96,KgPKH+KgPKHse*1.96))/2
  
  
```

# Harvest Control Rule
```{r}
value <- HSRandom_Index_WZS$mean3_biomass[HSRandom_Index_WZS$Year==Current.Year] 
HCR(value)


#HCR(value, q=0.0019) # when using calculated weight...but HCR function uses q = 0.002

```


## plot survey indices SSR version
```{r}
#RVinputs$Index


png(file.path(wd,'figures',"ModelindicesSSR_WZS_wo2021.png"), width =8, height = 6,units='in',pointsize=12, res=300,type='cairo')
par(mfcol=c(2,1), mar = c(0,4,0,0.5), omi = c(0.5, 0.25, 0.5, 0.25),las=1)

plot(NPT~year,NSRV_Index_pre2021, ylab="Number / tow", xlab = '', ylim=c(0,max(NPT,na.rm=T)*1.2), type='o',pch=16,xlim=c(1970,Current.Year),xaxt='n')
axis(1,lab=F)
lines(NPT~year,NSRV_Index_post2021, ylab="Number / tow", xlab = '', ylim=c(0,max(NPT,na.rm=T)*1.2), type='o',pch=16,xlim=c(1970,Current.Year),xaxt='n')
abline(h = 0.3, lty = "dashed")
points(c(2018,2022),NSRV_Index$NPT[NSRV_Index$year%in%c(2018,2022)],col='red',cex=2)
legend(x="topleft",   # Coordinates (x also accepts keywords)
       legend = c("RV Survey"),
       text.col = "black",
       bty = "n",

)

plot(KgPKH~Year,HSFixed_Index,type='o',pch=16,col='grey40',ylim=c(0,255),xlim=c(1970,Current.Year), ylab="Kg / 1,000 hooks")
lines(mean3_biomass~Year,HSFixed_Index, lwd = 3, col = 'grey40')
lines(KgPKH~Year,HSRandom_Index_WZS,type='o',pch=16,col='blue')
#lines(KgPKH+KgPKHse~Year,HSRandom_Index,lty=2,col='blue')
#lines(KgPKH-KgPKHse~Year,HSRandom_Index,lty=2,col='blue')
lines(mean3_biomass~Year,HSRandom_Index_WZS,lwd=3, col = "blue")
legend(x="topleft", inset = 0.05,   # Coordinates (x also accepts keywords)
       title = "Halibut Longline Survey",
       title.col = "black",
       legend = c("Fixed Station", "Stratified Random"), # Vector with the name of each group
       text.col = c("grey40", "blue"),
       bty = "n",

)

dev.off()


##checking out #/tow and biomass/tow
RVinputs <- data.frame(RVinputs)
RVinputs <- RVinputs %>% mutate(Index.BPT = (Index.B/Index.tunits))

png(file.path(wd,'figures',"RVsurvey_AbundanceVSBiomass.png"), width =8, height = 6,units='in',pointsize=12, res=300,type='cairo')

par(mfcol=c(2,1), mar = c(0,4,0,0.5), omi = c(0.5, 0.25, 0.5, 0.25),las=1)

plot(NPT~year,NSRV_Index, ylab="Number / tow", xlab = '', ylim=c(0,max(NPT,na.rm=T)*1.2), type='o',pch=16,xlim=c(1970,Current.Year),xaxt='n')
axis(1,lab=F)
points(c(2018,2021,2022),NSRV_Index$NPT[NSRV_Index$year%in%c(2018,2021,2022)],col='red',cex=2)

plot(Index.BPT~Index.year,RVinputs, ylab="Biomass / tow", xlab = '', ylim=c(0,max(Index.BPT,na.rm=T)*1.2), type='o',pch=16,xlim=c(1970,Current.Year))
axis(1,lab=F)
points(c(2018,2021,2022),RVinputs$Index.BPT[RVinputs$Index.year%in%c(2018,2021,2022)],col='red',cex=2)

dev.off()
```

# Comparing calculated weight vs estimated weight
```{r}
 HSRandom_Index<-StratifiedRandomSurveyIndex(datadir=datadir,yrs=2017:Current.Year,use.calc.weight = F)
 HSRandom_Index2<-StratifiedRandomSurveyIndex(datadir=datadir,yrs=2017:Current.Year,use.calc.weight = T)

```

```{r}
  HSRandom_Index<-StratifiedRandomSurveyIndex(datadir=datadir,yrs=2017:Current.Year,use.calc.weight = F)
  HSRandom_Index1<-StratifiedRandomSurveyIndex(datadir=datadir,yrs=2017:Current.Year,use.calc.weight = F)

```

##some playing around...trying to figure some stuff out
```{r}
 #index using calculated weight vs estimated weight (I think we use estimated weight)
 plot(KgPKH~Year,HSRandom_Index,ylim=c(0,150),type='l',lwd=2)
 with(HSRandom_Index,lines(Year,mean3_biomass,lty=2))
 with(HSRandom_Index2,lines(Year,KgPKH,lwd=2,col=2))
 with(HSRandom_Index2,lines(Year,mean3_biomass,lty=2,col=2))
 
 # Q-adjusted (using estimated weight) - plot of biomass rather than the index
 plot(KgPKH/0.002~Year,HSRandom_Index,ylim=c(0,60000),type='l',lwd=2,ylab="Biomass")
 with(HSRandom_Index,lines(Year,mean3_biomass/0.002,lty=2))
 with(HSRandom_Index2,lines(Year,KgPKH/0.0019,lwd=2,col=2))
 with(HSRandom_Index2,lines(Year,mean3_biomass/0.0019,lty=2,col=2))
 
 # Q-adjusted (adjusting Fixed with SR)
 plot(KgPKH/4.6~Year,HSFixed_Index,ylim=c(0,60),type='l',lwd=2,ylab="Biomas (kt)",las=1)
 #with(HSFixed_Index,lines(Year,mean3_biomass/0.0046,lty=2))
 with(HSRandom_Index,lines(Year,KgPKH/2,lwd=2,col=2))
# with(HSRandom_Index,lines(Year,mean3_biomass/0.002,lty=2,col=2))
 
 
 # comparing with and without zero strata
 plot(KgPKH~Year,HSRandom_Index_WZS,ylim=c(0,150),type='l',lwd=2)
 with(HSRandom_Index_WZS,lines(Year,mean3_biomass,lty=2)) #black line
 with(HSRandom_Index_WOZS,lines(Year,KgPKH,lwd=2,col=2)) #red line
 with(HSRandom_Index_WOZS,lines(Year,mean3_biomass,lty=2,col=2)) 
 
 # comparing with and without 5.2 and 5.3 combined...not sure this really works how i think it does...but lines ALMOST overlap...red line slightly more conservative
  HSRandom_Index_WOZS_5blend<-StratifiedRandomSurveyIndex(datadir=datadir,yrs=2017:Current.Year, ZeroStrata=F, restratify = T)
  HSRandom_Index_WOZS_test<-StratifiedRandomSurveyIndex(datadir=datadir,yrs=2017:Current.Year, ZeroStrata=F, restratify = F)
  HSRandom_Index5<-StratifiedRandomSurveyIndex(datadir=datadir,yrs=2017:Current.Year, ZeroStrata=T, restratify=F, combine.strata=52:53)
  HSRandom_Index_WZS_5_restrat<-StratifiedRandomSurveyIndex(datadir=datadir,yrs=2017:Current.Year, ZeroStrata=T, restratify=T, combine.strata=52:53)
 
png(file.path(wd,'figures',"LLindex_WZS_5combinedAll.png"), width =8, height = 6,units='in',pointsize=12, res=300,type='cairo')
 plot(KgPKH~Year,HSRandom_Index_WZS,ylim=c(0,150),type='l',lwd=2)
 with(HSRandom_Index_WZS,lines(Year,mean3_biomass,lty=2)) #black line
 with(HSRandom_Index5,lines(Year,KgPKH,lwd=2,col=2)) #red line
 with(HSRandom_Index5,lines(Year,mean3_biomass,lty=2,col=2))
 dev.off()
 
 ##for Bruce on zero strata
 png(file.path(wd,'figures',"LLindex_overtime_WandWOZS.png"), width =8, height = 6,units='in',pointsize=12, res=300,type='cairo')
 plot(KgPKH~Year,HSRandom_Index_WOZS,ylim=c(0,150),type='l',lwd=2)
 with(HSRandom_Index_WZS_5_restrat,lines(Year,KgPKH,lwd=2,col=2)) #red line
 dev.off()
```

##looking at #sets/area for each strata under different strata schemes and over time (to investigate the impact of zero strata)
```{r}
RSdata <- RandomSurveyData(sp=30, datadir, add.gear=F, add.LF=F, hook.data=F)

load(file.path(datadir, "Survey","SurveyStrata.RData"))
StrataAreas1 <-data.frame(StrataAreas)
surveyStrataPolyLL1 <- data.frame(surveyStrataPolyLL)

load(file.path(datadir, "Survey","SurveyStrata2022.RData"))
StrataAreas2022 <-data.frame(StrataAreas)
surveyStrataPolyLL2022 <- data.frame(surveyStrataPolyLL)

load(file.path(datadir, "Survey","SurveyStrata2023.RData"))
StrataAreas2023 <-data.frame(StrataAreas)
surveyStrataPolyLL2023 <- data.frame(surveyStrataPolyLL)

load(file.path(datadir, "Survey","SurveyStrata2023NoBO.RData"))
StrataAreas2023WOZS <-data.frame(StrataAreas)
surveyStrataPolyLL2023WOZS <- data.frame(surveyStrataPolyLL)

RSdata_original <- reStratify(sets = RSdata, strata = surveyStrataPolyLL1)
RSdata_2022 <- reStratify(sets = RSdata, strata = surveyStrataPolyLL2022)
RSdata_2023 <- reStratify(sets = RSdata, strata = surveyStrataPolyLL2023)
RSdata_2023WOZS <- reStratify(sets = RSdata, strata = surveyStrataPolyLL2023WOZS)

#looking using original strata scheme
RSdata_original_summary <- RSdata_original %>%
  group_by(YEAR, PID) %>%
  summarize(n = n())

RSdata_originalWarea <- RSdata_original_summary %>% left_join(StrataAreas1, by = "PID") %>% mutate(prop = signif(n/area)*10000, digits = 3)
RSdata_originalWarea_wide <- RSdata_originalWarea %>% select(PID, YEAR, prop) %>% spread(YEAR, prop)

#looking using 2022 strata scheme
RSdata_2022_summary <- RSdata_2022 %>%
  group_by(YEAR, PID) %>%
  summarize(n = n())

RSdata_originalWarea <- RSdata_original_summary %>% left_join(StrataAreas1, by = "PID") %>% mutate(prop = signif(n/area)*10000, digits = 3)
RSdata_originalWarea_wide <- RSdata_originalWarea %>% select(PID, YEAR, prop) %>% spread(YEAR, prop)


##density for new zero strata allocation design
load(file.path(datadir, "Survey","SurveyStrataOneZeroStrata.RData"))
StrataAreas_ZS2024 <-data.frame(StrataAreas)
StrataAreas_ZS2024$PID <- as.character(StrataAreas_ZS2024$PID)
surveyStrataPolyLL2024 <- data.frame(surveyStrataPolyLL)

allocation2024 <- read.csv("C:/Users/harperd/Documents/Halibut/RDataVault/Survey/allocation2024.csv")

allocation2024$SID <- as.character(gsub("\\.", "", (allocation2024$SID)))
allocation2024[18, "SID"] <- 60
allocation2024 <- allocation2024 %>% rename(PID = SID) %>% select(-Area)

allocation2024 <- allocation2024 %>% left_join(StrataAreas_ZS2024, by = "PID")

allocationdensity2024 <- allocation2024 %>% mutate(density = signif((Sets/area)*100, digits = 3))

allocationdensity2024_WOZS <- allocationdensity2024 %>% filter(PID != 60)
mean(allocationdensity2024_WOZS$density)
```
Survey update document
```{r}
ComIndex<-CommercialIndexData(datadir=datadir,bins=bins,by.sex=F,add.portsampling=F,add.LF = F)
FixedSurvey<-FixedSurveyData(datadir=datadir,by.sex=F, bins=bins, add.LF = F)
RandomSurvey<-RandomSurveyData(sp=30,datadir=datadir,add.gear=F,add.LF=F,bins=bins,by.sex=F,hook.data=F)

png(file.path(wd,'figures',"CI.png"), width =6, height = 4,units='in',pointsize=9, res=300,type='cairo')
   # par(mfcol=c(rows,ceiling(length(yrs)/rows)), mar = c(0,2,0,0.5), omi = c(0.85, 0.75, 0.75, 0.5))

ComIndex$KgPKH<-ComIndex$EST_COMBINED_WT/ComIndex$NUM_HOOK_HAUL*1000
dot <- ComIndex %>% group_by(YEAR) %>% summarize(mean = mean(KgPKH, na.rm = TRUE))
CI<-with(ComIndex,tapply(KgPKH,YEAR, mean, na.rm=T))
plot(as.numeric(names(CI)),CI,xlab = '',ylab = 'Kg / 1,000 Hooks',ylim=c(0,500),type='b',pch=16)
lines(KgPKH~Year,HSRandom_Index_WZS,type='o',pch=16,col='blue')
points(2023,dot$mean[dot$YEAR==2023],col='red',cex=2)



FixedSurvey$KgPKH<-FixedSurvey$EST_COMBINED_WT/FixedSurvey$NUM_HOOK_HAUL*1000
FS<-with(FixedSurvey,tapply(KgPKH,YEAR, mean, na.rm=T))
lines(as.numeric(names(FS)),FS,col='grey',lty=2)
legend(x="topleft", inset = 0.05,   # Coordinates (x also accepts keywords)
       title.col = "black",
       legend = c("Biological Sampling Program","Fixed Station", "Stratified Random"), # Vector with the name of each group
       text.col = c("black","grey40", "blue"),
       bty = "n",

)




dev.off()
```

