---
title: "Length Weight Relationship"
author: "Danni"
date: "February 16, 2021"
output: html_document
--- 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
 library(Mar.datawrangling)
 library(tidyverse)
 library(lubridate)
 library(stringr)

datadir="C:/Users/harperd/Documents/Halibut/RDataVault"
wd="C:/Users/harperd/Documents/Halibut/GitDataInputs/DataInputs/"
source(file.path(wd, ".passwords"))

```

Extract MAR RV survey data using RORACLE
```{r}
#get_data("rv", data.dir = "C:/Users/harperd/Documents/Halibut/RDataVault/", usepkg = "roracle",fn.oracle.username = "hubleyb", fn.oracle.password = "R4#vmxtas", fn.oracle.dsn = "PTRAN", force.extract = T)
```

Extract MAR RV survey data from existing dump
```{r}
get_data("rv", data.dir = datadir)
```

Pull halibut morphology data from the MAR RV survey since beginning of survey data set (1970?)
```{r}
hal <- GSDET %>%
  filter(SPEC == "30")
```

Scatter plot of Length and Weight of halibut from RV survey
```{r}
ggplot(hal, aes(FLEN, FWT)) +
  geom_point()
```
There are 1188 fish that have no weights associated with the lengths in the MAR RV survey data

Combining the hal data with the set info to determine strata
```{r}
setdata_halRV <- merge(hal, GSINF, by.x = c("MISSION","SETNO"), 
                       by.y = c("MISSION","SETNO"), all.x = TRUE, all.y = FALSE) %>% 
  select("MISSION", "SETNO", "SPEC", "FSHNO", "FLEN", "FSEX", "FMAT", "FWT", "AGMAT", "AGE", "CLEN", "REMARKS.x", "SIZE_CLASS", "SDATE", "STRAT", "SLAT", "SLONG", "ELAT", "ELONG", "ETIME", "REMARKS.y", "START_DEPTH", "END_DEPTH", "SURFACE_TEMPERATURE", "BOTTOM_TEMPERATURE", "BOTTOM_SALINITY", "STATION")
```

Joining the Stratum table to attach associated NAFO info
```{r}
setdata_halRV <- setdata_halRV %>%
  left_join(GSSTRATUM, by = "STRAT") %>%
  select(-"DMIN", -"DMAX", -"AREA", -"DEPTH")
```

Create year column and column of 5yr bins starting in 1970
```{r}
setdata_halRV <- setdata_halRV %>%
  mutate(YEAR = substr(SDATE, 1,4))

setdata_halRV <- setdata_halRV %>%
  mutate(yr_bin = 
           ifelse(YEAR %in% c(1970, 1971, 1972, 1973, 1974, 1975), "1970-1975", 
           ifelse(YEAR %in% c(1976, 1977, 1978, 1979, 1980), "1976-1980",
           ifelse(YEAR %in% c(1981, 1982, 1983, 1984, 1985), "1981-1985", 
           ifelse(YEAR %in% c(1986, 1987, 1988, 1989, 1990), "1986-1990",
           ifelse(YEAR %in% c(1991, 1992, 1993, 1994, 1995), "1991-1995",
           ifelse(YEAR %in% c(1996, 1997, 1998, 1999, 2000), "1996-2000",
           ifelse(YEAR %in% c(2001, 2002, 2003, 2004, 2005), "2001-2005",
           ifelse(YEAR %in% c(2006, 2007, 2008, 2009, 2010), "2006-2010",
           ifelse(YEAR %in% c(2011, 2012, 2013, 2014, 2015), "2011-2015",
           ifelse(YEAR %in% c(2016, 2017, 2018, 2019, 2020), "2016-2020", "Error")))))))))))

```


See NAFO zones represented in the hal data from the Mar RV survey data
```{r}
unique(setdata_halRV$NAME)
```
There are 10 fish where there is no corresponding NAFO (can likely just look it up using the coordinates?)

Reducing the fine scale of NAFO regions to the larger zones (e.g. 4Vs to 4V...) in new column NAFO
```{r}
setdata_halRV <- setdata_halRV %>%
  mutate(NAFO = substr(NAME, 1, 2))
```

Graphing the length by weight by larger NAFO zone
```{r}
ggplot(setdata_halRV, aes(FLEN, FWT)) +
  geom_point()+
  facet_wrap(~NAFO)
```
Graphing the length by weight for NAFO and year
```{r}
ggplot(setdata_halRV, aes(FLEN, FWT)) +
  geom_point()+
  facet_grid(YEAR~NAFO)
```

Graphing the length by weight for NAFO and 5yr bins (when colouring the points by sex we lose even more data (from 1188 to 1204 fish not included in the figure below...))
```{r}
ggplot(setdata_halRV, aes(FLEN, FWT, color = as.factor(FSEX))) +
  geom_point()+
  scale_color_manual(name = "FSEX",
                     values = c("0" = "black",
                                  "1" = "blue",
                                  "2" = "deeppink",
                                "NA" = "black"),
  labels = c("0", "male", "female", "NA"))+
  facet_grid(yr_bin~NAFO, scales = "free")
```
Checking how many fish sampled by year and NAFO
```{r}
NAFO_4_5 <- setdata_halRV %>%
  mutate(NAFO = substr(NAFO,1,1))


num_fish_sampled <- data.frame(table(NAFO_4_5$YEAR, NAFO_4_5$NAFO))

#this counts how many fish are in "numb_fish_sampled"
sum(num_fish_sampled$Freq)
```
There are 10 fewer fish in num_fish_sampled than in setdata_RVhal because 10 fish in the original data don't have an associated NAFO zone

Many of these don't match up for NAFO and year...(how is the data so different?)


####NEXT SECTION FOR ISDB DATA ####

Extract ISDB data using RORACLE
```{r}
#get_data("isdb", data.dir = "C:/Users/harperd/Documents/Halibut/RDataVault/", usepkg = "roracle",fn.oracle.username = "hubleyb", fn.oracle.password = "R4#vmxtas", fn.oracle.dsn = "PTRAN", force.extract = T)
```

Pull ISDB data from where it is stored locally (previously pulled from database using oracle)
```{r}
get_data("isdb", data.dir = datadir)
```

Pulling the halibut trips from the trips table (30 = halibut, 7057 = halibut longline survey, 7058 = halibut port sampling)
```{r}
hal_trips <- ISTRIPS %>%
  filter(TRIPCD_ID %in% c("30", "7057", "7058"))
```

Pulling all the sets that belong to the halibut trips identified above
```{r}
hal_fishsets <- hal_trips %>%
  left_join(ISFISHSETS, by = "TRIP_ID") %>%
  drop_na(FISHSET_ID) #need to drop NAs as there are 281 catches with NA for fish set in ISCATCHED that are being matched in the next join
```

Pulling out all the catches from the identified sets above
```{r}
hal_catches <- hal_fishsets %>%
  left_join(ISCATCHES, by = "FISHSET_ID")
```
**Weird that for species sought there are other species (30, 12, 15, 7099, 31, 23, 11, 10, 211, 201, 16, NA) even though the trip is identified as trip code 30...should I be isolating for only halibut sought trips???


Pulling out all fish caught in the catches from the halibut trips, and then filtering for just the halibut caught
```{r}
hal_fish <- hal_catches %>%
  left_join(ISFISH, by = "CATCH_ID") %>%
  filter(SPECCD_ID == "30")
```

First exploratory plot: plotting all length against width (some will not be plotted and provides idea of how many individuals are missing a length or weight measurement)
```{r}
ggplot(hal_fish, aes(FISH_LENGTH, FISH_WEIGHT)) +
  geom_point()
```
From looking at the data in hal_fish and the graph there are A LOT of issues: lots of missing lengths and weights, and weights that don't make sense (what's the units for the weights?)

Trimming down the dataframe to useful columns
```{r}


```

