---
title: "ModelRuns"
author: "Brad"
date: "8/17/2021"
output: html_document
---

## Use this script to update data for assessment model, check final year as you go through each chunk

```{r setup, include=FALSE}
library(ROracle)
library(devtools)
library(tidyverse)
library(SpatialHub)
source(file.path(getwd(), "directories.r"))
source(file.path(wd, "passwords.r")) 

scaldir<-file.path(datadir,"scal")
#scaldir<-"R:/Science/Population Ecology Division/Shared/Atlantic Halibut/Assessments & Updates/2021 Framework/Data/Model Inputs"
#scaldir<-"C:/Users/hubleyb/Documents/Halibut/git/siscal-ah/Data/AtlHal"

ds_all <- Mar.datawrangling::load_datasources()

devtools::load_all(".")


bins=seq(0,260,5)
# labelled by upper bound

```


## Update data from database 

```{r }


# Update data from database 

# Mar.datawrangling::get_data(db='isdb',data.dir=datadir,fn.oracle.username = uid, fn.oracle.password = pwd, fn.oracle.dsn = 'ptran',usepkg='roracle',force.extract=T)
# open and check setprofile_wide to confirm that you have year, latitude & longitude

# Mar.datawrangling::get_data_custom(schema="observer", data.dir = datadir, tables = c("ISFISHLENGTHS","ISSAMPLES"), usepkg = "roracle", fn.oracle.dsn= "PTRAN", fn.oracle.username=uid, fn.oracle.password =pwd)

#get_ps_data(data.dir=datadir,fn.oracle.username = uid, fn.oracle.password = pwd, fn.oracle.dsn = 'ptran',usepkg='roracle',force.extract=T)


#  Mar.datawrangling::get_data(db='rv',data.dir=datadir,fn.oracle.username = uid, fn.oracle.password = pwd, fn.oracle.dsn = 'ptran',usepkg='roracle',force.extract=T)


```


# read in data from old inputs

```{r }
old_data_inputs<-getOldScalInputs(scaldir=scaldir)

# Old Model Starting parameters
#old_ipars<-lisread(file.path(scaldir,'scal_base.pin'))

```



# Output biological parameters table
```{r }
bioPars<-old_data_inputs[c("lInf_m","lInf_f","vonK_m","vonK_f","wt_a","wt_b","aMat_50","aMat_95","prior_initM","dM")]
bioPars$dM2<-bioPars$dM[3]
bioPars$dM<-bioPars$dM[1]
bioPars$wt_a=0.006803616
bioPars$wt_b=3.119924
bioParsTab<-tibble(data.frame(Parameter=names(bioPars),Value=unlist(bioPars)))
write.csv(bioParsTab,file.path(scaldir,"bioPars.csv"),row.names=F)

BiologicalParameters<-bioPars
BiologicalParameters$bins<-bins
```

#NS RV survey

```{r}
RVdata<-get4VWXRV(uid, pwd, use.local=T,datadir=datadir)

# for spatial model
allStrata<-sort(unique(RVdata[[1]]$STRAT))
Strata4VW<-allStrata[1:25]
Strata4X<-allStrata[26:46]

RVinputs4X<-prepRVdata(RVdata,years=1970:2024,strat= Strata4X,bins=bins, raw=T)
RVinputs4X$NSRV_unsexed<-RVinputs4X$NSRV_combined-(RVinputs4X$NSRV_males+RVinputs4X$NSRV_females) # unsexed
NSRV_Index4X<-RVinputs4X$Index
NSRV_Lengths4X<-list(NSRV4X_males=RVinputs4X$NSRV_males,
                   NSRV4X_females=RVinputs4X$NSRV_females,
                   NSRV4X_combined=RVinputs4X$NSRV_combined,
                   NSRV4X_unsexed=RVinputs4X$NSRV_unsexed)



RVinputs4VW<-prepRVdata(RVdata,years=1970:2024,strat= Strata4VW,bins=bins, raw=T)
RVinputs4VW$NSRV_unsexed<-RVinputs4VW$NSRV_combined-(RVinputs4VW$NSRV_males+RVinputs4VW$NSRV_females) # unsexed
NSRV_Index4VW<-RVinputs4VW$Index
NSRV_Lengths4VW<-list(NSRV4VW_males=RVinputs4VW$NSRV_males,
                   NSRV4VW_females=RVinputs4VW$NSRV_females,
                   NSRV4VW_combined=RVinputs4VW$NSRV_combined,
                   NSRV4VW_unsexed=RVinputs4VW$NSRV_unsexed)



BubblePlotLF(NSRV_Lengths4X,bins=bins,yrs=1970:2024,path=file.path(wd,'figures'),filen='RVSurvey4X',inches=0.1)
BubblePlotLF(NSRV_Lengths4VW,bins=bins,yrs=1970:2024,path=file.path(wd,'figures'),filen='RVSurvey4VW',inches=0.1)
```

#NL RV Survey
```{r}

#NLRV_ny<-read.csv(file.path(datadir,"Newfoundland Data","NLsurvey_n_yr.csv"))
#NLRV_nys<-read.csv(file.path(datadir,"Newfoundland Data","NLsurvey_n_yrsex.csv"))

NLRV_males<-read.csv(file.path(datadir,"Newfoundland Data","NLsurvey_m_LF_obs.csv"))
NLRV_males<-data.frame(pivot_wider(NLRV_males,names_from = length,values_from = obs))
row.names(NLRV_males)<-NLRV_males$year
NLRV_males<-NLRV_males[,-1]
names(NLRV_males)<-paste0("L",bins[-1])
NLRV_females<-read.csv(file.path(datadir,"Newfoundland Data","NLsurvey_f_LF_obs.csv"))
NLRV_females<-data.frame(pivot_wider(NLRV_females,names_from = length,values_from = obs))
row.names(NLRV_females)<-NLRV_females$year
NLRV_females<-NLRV_females[,-1]
names(NLRV_females)<-paste0("L",bins[-1])
NLRV_combined<-read.csv(file.path(datadir,"Newfoundland Data","NLsurvey_mfu_LF_obs.csv"))
NLRV_combined<-data.frame(pivot_wider(NLRV_combined,names_from = length,values_from = obs))
row.names(NLRV_combined)<-NLRV_combined$year
NLRV_combined<-NLRV_combined[,-1]
names(NLRV_combined)<-paste0("L",bins[-1])
NLRV_unsexed<-NLRV_combined-(NLRV_males+NLRV_females) # unsexed

NLRV_Lengths3NOP<-list(NLRV3NOP_males=NLRV_males,
                   NLRV3NOP_females=NLRV_females,
                   NLRV3NOP_combined=NLRV_combined,
                   NLRV3NOP_unsexed=NLRV_unsexed)

NLRV_SCAL<-read.csv(file.path(scaldir,"NLSurveyCAL.csv"))
row.names(NLRV_SCAL)<-NLRV_SCAL$year
NLRV_SCAL<-NLRV_SCAL[,-1]
names(NLRV_SCAL)<-paste0("L",bins[-1])


#BubblePlotLF(NLRV_Lengths3NOP,bins=bins,yrs=1996:2019,path=file.path(wd,'figures'),filen='RVSurvey3NOP',inches=0.3)

NLRV_SumData<-read.csv(file.path(datadir,"Newfoundland Data","strat2_abundance_summary.csv"))
NLRV_Index3NOP<-subset(NLRV_SumData,select=c("survey.year","total","sample.units","mean","var"))
NLRV_Index3NOP$CV<-sqrt(NLRV_Index3NOP$var)/NLRV_Index3NOP$mean
#NLRV_Index<-rowSums(NLRV_SCAL)

```



#Halibut Survey Catch Composition
```{r}
  Yrs1 = 2000:2020        
  Yrs2 = 2017:2024        
 
   # for spatial model

  FixedSurvey <- FixedSurveyData(datadir=datadir, bins=bins,by.sex=T)
  nafo4X<-unique(FixedSurvey$NAFAREA_ID[substr(FixedSurvey$NAFAREA_ID,1,2)%in%c("5Z","5Y","4X")])
  nafo4VW<-unique(FixedSurvey$NAFAREA_ID[substr(FixedSurvey$NAFAREA_ID,1,2)%in%c("4V","4W")])
  nafo3NOP<-unique(FixedSurvey$NAFAREA_ID[substr(FixedSurvey$NAFAREA_ID,1,2)%in%c("3O","3N","3P")])
  
  bioMap(nafo = T)
  points(LAT1~LONG1,data=subset(FixedSurvey,NAFAREA_ID%in%nafo4X),col=rgb(0,0,1,0.2),pch=16)
  points(LAT1~LONG1,data=subset(FixedSurvey,NAFAREA_ID%in%nafo4VW),col=rgb(1,0,0,0.2),pch=16)
  points(LAT1~LONG1,data=subset(FixedSurvey,NAFAREA_ID%in%nafo3NOP),col=rgb(0,1,0,0.2),pch=16)

  
  RandomSurvey <- RandomSurveyData(datadir=datadir,bins=bins,by.sex=T)

  allStrata<-sort(unique(RandomSurvey$ASSIGNED_STRATUM_ID))
  Strata4VW<-allStrata[4:10]
  Strata4X<-allStrata[1:3]
  Strata3NOP<-allStrata[11:20]
  
  bioMap(nafo = T)
  points(LAT1~LONG1,data=subset(RandomSurvey,ASSIGNED_STRATUM_ID%in%Strata4X),col=rgb(0,0,1,0.2),pch=16)
  points(LAT1~LONG1,data=subset(RandomSurvey,ASSIGNED_STRATUM_ID%in%Strata4VW),col=rgb(1,0,0,0.2),pch=16)
  points(LAT1~LONG1,data=subset(RandomSurvey,ASSIGNED_STRATUM_ID%in%Strata3NOP),col=rgb(0,1,0,0.2),pch=16)

  

  # Construct CLF
  # Fixed survey numbers at length    
  FixedLF4X <- constructLF(LFdata=subset(FixedSurvey,NAFAREA_ID%in%nafo4X),bins=bins, Yrs = Yrs1)
  FixedLF4VW <- constructLF(LFdata=subset(FixedSurvey,NAFAREA_ID%in%nafo4VW),bins=bins, Yrs = Yrs1)
  FixedLF3NOP <- constructLF(LFdata=subset(FixedSurvey,NAFAREA_ID%in%nafo3NOP),bins=bins, Yrs = Yrs1)
  

  HSFixed_males<-getNAL(FixedLF4X,sex=1)
  HSFixed_females<-getNAL(FixedLF4X,sex=2)
  HSFixed_combined<-getNAL(FixedLF4X,sex='all')
  HSFixed_unsexed<-getNAL(FixedLF4X,sex=0)
  
  HSFixed_Lengths4X<-list(HSFixed_males=HSFixed_males,
                        HSFixed_females=HSFixed_females,
                        HSFixed_combined=HSFixed_combined,
                        HSFixed_unsexed=HSFixed_unsexed)
  
 
  HSFixed_males<-getNAL(FixedLF4VW,sex=1)
  HSFixed_females<-getNAL(FixedLF4VW,sex=2)
  HSFixed_combined<-getNAL(FixedLF4VW,sex='all')
  HSFixed_unsexed<-getNAL(FixedLF4VW,sex=0)
  
  HSFixed_Lengths4VW<-list(HSFixed_males=HSFixed_males,
                        HSFixed_females=HSFixed_females,
                        HSFixed_combined=HSFixed_combined,
                        HSFixed_unsexed=HSFixed_unsexed)
  
 
  HSFixed_males<-getNAL(FixedLF3NOP,sex=1)
  HSFixed_females<-getNAL(FixedLF3NOP,sex=2)
  HSFixed_combined<-getNAL(FixedLF3NOP,sex='all')
  HSFixed_unsexed<-getNAL(FixedLF3NOP,sex=0)
  
  HSFixed_Lengths3NOP<-list(HSFixed_males=HSFixed_males,
                        HSFixed_females=HSFixed_females,
                        HSFixed_combined=HSFixed_combined,
                        HSFixed_unsexed=HSFixed_unsexed)
  
  # Random survey numbers at length    
  RandomLF4VW <- constructLF(LFdata=subset(RandomSurvey,ASSIGNED_STRATUM_ID%in%Strata4VW),bins=bins, Yrs = Yrs2)
  RandomLF4X <- constructLF(LFdata=subset(RandomSurvey,ASSIGNED_STRATUM_ID%in%Strata4X),bins=bins, Yrs = Yrs2)
  RandomLF3NOP <- constructLF(LFdata=subset(RandomSurvey,ASSIGNED_STRATUM_ID%in%Strata3NOP),bins=bins, Yrs = Yrs2)
  
  HSRandom_males<-getNAL(RandomLF4VW,sex=1)
  HSRandom_females<-getNAL(RandomLF4VW,sex=2)
  HSRandom_combined<-getNAL(RandomLF4VW,sex='all')
  HSRandom_unsexed<-getNAL(RandomLF4VW,sex=0)
  
  HSRandom_Lengths4VW<-list(HSRandom_males=HSRandom_males,
                         HSRandom_females=HSRandom_females,
                         HSRandom_combined=HSRandom_combined,
                         HSRandom_unsexed=HSRandom_unsexed)
  
  HSRandom_males<-getNAL(RandomLF4X,sex=1)
  HSRandom_females<-getNAL(RandomLF4X,sex=2)
  HSRandom_combined<-getNAL(RandomLF4X,sex='all')
  HSRandom_unsexed<-getNAL(RandomLF4X,sex=0)
  
  HSRandom_Lengths4X<-list(HSRandom_males=HSRandom_males,
                         HSRandom_females=HSRandom_females,
                         HSRandom_combined=HSRandom_combined,
                         HSRandom_unsexed=HSRandom_unsexed)
  
  HSRandom_males<-getNAL(RandomLF3NOP,sex=1)
  HSRandom_females<-getNAL(RandomLF3NOP,sex=2)
  HSRandom_combined<-getNAL(RandomLF3NOP,sex='all')
  HSRandom_unsexed<-getNAL(RandomLF3NOP,sex=0)
  
  HSRandom_Lengths3NOP<-list(HSRandom_males=HSRandom_males,
                         HSRandom_females=HSRandom_females,
                         HSRandom_combined=HSRandom_combined,
                         HSRandom_unsexed=HSRandom_unsexed)
  
  #HSRandom_Index<-data.frame(Year= 2017:2020,
   #                          Index= c(2.204668e-05, 3.263344e-05, 3.004447e-05, 1.819697e-05), 
   #                          se= c(1.075617e-06, 1.291483e-06, 1.437159e-06, 8.738080e-07))
  
  HSFixed_Index4X<-FixedSurveyIndex(datadir=datadir,yrs=1998:2022,nafo=nafo4X)
  HSFixed_Index4VW<-FixedSurveyIndex(datadir=datadir,yrs=1998:2022,nafo=nafo4VW)
  HSFixed_Index3NOP<-FixedSurveyIndex(datadir=datadir,yrs=1998:2022,nafo=nafo3NOP)
  
  
  HSRandom_Index4VW<-StratifiedRandomSurveyIndex(datadir=datadir,yrs=2017:2023,use.calc.weight = F, select.strata = as.numeric(substr(Strata4VW,2,3)))
  HSRandom_Index4X<-StratifiedRandomSurveyIndex(datadir=datadir,yrs=2017:2023,use.calc.weight = F, select.strata = as.numeric(substr(Strata4X,2,3)))
  HSRandom_Index3NOP<-StratifiedRandomSurveyIndex(datadir=datadir,yrs=2017:2023,use.calc.weight = F, select.strata = as.numeric(substr(Strata3NOP,2,3)))

 # Cusk
  HSRandom_IndexCusk<-StratifiedRandomSurveyIndex(datadir=datadir,yrs=2017:2023,use.calc.weight = F, select.strata = as.numeric(substr(allStrata[1:10],2,3)),sp=15) 

  BubblePlotLF(Fishery_Lengths,bins=bins,yrs=CCyrs,path=file.path(wd,'figures'),filen='Commercial',inches=0.1)

  
```



#Landings
```{r}
#Landings_t<-read.csv(file.path(scaldir,"nafoSCAL.csv"))
#Landings_t<-rbind(Landings_t,data.frame(X=52,Year=2021,Landings_t[51,3:6]))

# updated to 2021
Landings_t<-read.csv(file.path(scaldir,"nafoSISCAL.csv"))
```


#Fishery Catch Composition
```{r}
CCyrs=1978:2024

 nafo3NOP = c( "3N", "3NA", "3NB", "3NC", "3ND", "3NE", "3NF", "3O", "3OA", "3OC", "3OD", "3OE", "3PS", "3U")
 nafo4VW = c("4V","4VB", "4VC", "4VN", "4VNW", "4VS", "4VSB", "4VSC","4VSW","4VW","4VX", "4W", "4WD", "4WE", "4WF", "4WG", "4WH", "4WJ", "4WK", "4WL", "4WM","4WX")
 nafo4X5Z = c("4XMNOPQ", "4X", "4XL", "4XM", "4XN", "4XO", "4XP", "4XQ", "4XR", "4XU", "4XS", "4XRS", "5YB", "5ZJ", "5ZM", "5ZN","5ZEJ")

PortData<-PortSampleData(datadir=datadir,by.sex=T, bins=bins)
portNAFO3NOPOT=constructLF(LFdata=subset(PortData,NAFO%in%nafo3NOP&GEAR=="OT"),bins=bins, Yrs = CCyrs)
portNAFO4X5ZOT=constructLF(LFdata=subset(PortData,NAFO%in%nafo4X5Z&GEAR=="OT"),bins=bins, Yrs = CCyrs)
portNAFO4VWOT=constructLF(LFdata=subset(PortData,NAFO%in%nafo4VW&GEAR=="OT"),bins=bins, Yrs = CCyrs)
portNAFO3NOPLL=constructLF(LFdata=subset(PortData,NAFO%in%nafo3NOP&GEAR=="LL"),bins=bins, Yrs = CCyrs)
portNAFO4X5ZLL=constructLF(LFdata=subset(PortData,NAFO%in%nafo4X5Z&GEAR=="LL"),bins=bins, Yrs = CCyrs)
portNAFO4VWLL=constructLF(LFdata=subset(PortData,NAFO%in%nafo4VW&GEAR=="LL"),bins=bins, Yrs = CCyrs)


ObserverData<-ObsData(datadir=datadir,by.sex=T, bins=bins,min.length=0)
obsNAFO3NOPOT=constructLF(LFdata=subset(ObserverData,NAFO%in%nafo3NOP&GEAR=="OT"),bins=bins, Yrs = CCyrs)
obsNAFO4X5ZOT=constructLF(LFdata=subset(ObserverData,NAFO%in%nafo4X5Z&GEAR=="OT"),bins=bins, Yrs = CCyrs)
obsNAFO4VWOT=constructLF(LFdata=subset(ObserverData,NAFO%in%nafo4VW&GEAR=="OT"),bins=bins, Yrs = CCyrs)
obsNAFO3NOPLL=constructLF(LFdata=subset(ObserverData,NAFO%in%nafo3NOP&GEAR=="LL"),bins=bins, Yrs = CCyrs)
obsNAFO4X5ZLL=constructLF(LFdata=subset(ObserverData,NAFO%in%nafo4X5Z&GEAR=="LL"),bins=bins, Yrs = CCyrs)
obsNAFO4VWLL=constructLF(LFdata=subset(ObserverData,NAFO%in%nafo4VW&GEAR=="LL"),bins=bins, Yrs = CCyrs)


# Long Line
LLNafo4X5Z_males<-getNAL(portNAFO4X5ZLL,sex=1)+getNAL(obsNAFO4X5ZLL,sex=1)
LLNafo4X5Z_females<-getNAL(portNAFO4X5ZLL,sex=2)+getNAL(obsNAFO4X5ZLL,sex=2)
LLNafo4X5Z_combined<-getNAL(portNAFO4X5ZLL,sex='all')+getNAL(obsNAFO4X5ZLL,sex='all')
LLNafo4X5Z_unsexed<-getNAL(portNAFO4X5ZLL,sex=0)+getNAL(obsNAFO4X5ZLL,sex=0)

LLNafo4VW_males<-getNAL(portNAFO4VWLL,sex=1)+getNAL(obsNAFO4VWLL,sex=1)
LLNafo4VW_females<-getNAL(portNAFO4VWLL,sex=2)+getNAL(obsNAFO4VWLL,sex=2)
LLNafo4VW_combined<-getNAL(portNAFO4VWLL,sex='all')+getNAL(obsNAFO4VWLL,sex='all')
LLNafo4VW_unsexed<-getNAL(portNAFO4VWLL,sex=0)+getNAL(obsNAFO4VWLL,sex=0)

# add NL observer data
NLmalesLL<-read.csv(file.path(datadir,"Newfoundland Data/NLmalesLLobs_binned2022.csv"))
NLmalesLL<-merge(data.frame(year=CCyrs),NLmalesLL,all=T)[,-1]
NLmalesLL[is.na(NLmalesLL)]<-0
dimnames(NLmalesLL)<-list(CCyrs,paste0("L",bins[-1]))
NLfemalesLL<-read.csv(file.path(datadir,"Newfoundland Data/NLfemalesLLobs_binned2022.csv"))
NLfemalesLL<-merge(data.frame(year=CCyrs),NLfemalesLL,all=T)[,-1]
NLfemalesLL[is.na(NLfemalesLL)]<-0
dimnames(NLfemalesLL)<-list(CCyrs,paste0("L",bins[-1]))
NLunknownLL<-read.csv(file.path(datadir,"Newfoundland Data/NLunknownLLobs_binned2022.csv"))
NLunknownLL<-merge(data.frame(year=CCyrs),NLunknownLL,all=T)[,-1]
NLunknownLL[is.na(NLunknownLL)]<-0
dimnames(NLunknownLL)<-list(CCyrs,paste0("L",bins[-1]))

LLNafo3NOP_males<-getNAL(portNAFO3NOPLL,sex=1)+getNAL(obsNAFO3NOPLL,sex=1)+NLmalesLL
LLNafo3NOP_females<-getNAL(portNAFO3NOPLL,sex=2)+getNAL(obsNAFO3NOPLL,sex=2)+NLfemalesLL
LLNafo3NOP_combined<-getNAL(portNAFO3NOPLL,sex='all')+getNAL(obsNAFO3NOPLL,sex='all')+NLmalesLL+NLfemalesLL+NLunknownLL
LLNafo3NOP_unsexed<-getNAL(portNAFO3NOPLL,sex=0)+getNAL(obsNAFO3NOPLL,sex=0)+NLunknownLL

# Otter Trawl
OTNafo4X5Z_males<-getNAL(portNAFO4X5ZOT,sex=1)+getNAL(obsNAFO4X5ZOT,sex=1)
OTNafo4X5Z_females<-getNAL(portNAFO4X5ZOT,sex=2)+getNAL(obsNAFO4X5ZOT,sex=2)
OTNafo4X5Z_combined<-getNAL(portNAFO4X5ZOT,sex='all')+getNAL(obsNAFO4X5ZOT,sex='all')
OTNafo4X5Z_unsexed<-getNAL(portNAFO4X5ZOT,sex=0)+getNAL(obsNAFO4X5ZOT,sex=0)

OTNafo4VW_males<-getNAL(portNAFO4VWOT,sex=1)+getNAL(obsNAFO4VWOT,sex=1)
OTNafo4VW_females<-getNAL(portNAFO4VWOT,sex=2)+getNAL(obsNAFO4VWOT,sex=2)
OTNafo4VW_combined<-getNAL(portNAFO4VWOT,sex='all')+getNAL(obsNAFO4VWOT,sex='all')
OTNafo4VW_unsexed<-getNAL(portNAFO4VWOT,sex=0)+getNAL(obsNAFO4VWOT,sex=0)

# add NL observer data
NLmalesOT<-read.csv(file.path(datadir,"Newfoundland Data/NLmalesOTobs_binned2022.csv"))
NLmalesOT<-merge(data.frame(year=CCyrs),NLmalesOT,all=T)[,-1]
NLmalesOT[is.na(NLmalesOT)]<-0
dimnames(NLmalesOT)<-list(CCyrs,paste0("L",bins[-1]))
NLfemalesOT<-read.csv(file.path(datadir,"Newfoundland Data/NLfemalesOTobs_binned2022.csv"))
NLfemalesOT<-merge(data.frame(year=CCyrs),NLfemalesOT,all=T)[,-1]
NLfemalesOT[is.na(NLfemalesOT)]<-0
dimnames(NLfemalesOT)<-list(CCyrs,paste0("L",bins[-1]))
NLunknownOT<-read.csv(file.path(datadir,"Newfoundland Data/NLunknownOTobs_binned2022.csv"))
NLunknownOT<-merge(data.frame(year=CCyrs),NLunknownOT,all=T)[,-1]
NLunknownOT[is.na(NLunknownOT)]<-0
dimnames(NLunknownOT)<-list(CCyrs,paste0("L",bins[-1]))

OTNafo3NOP_males<-getNAL(portNAFO3NOPOT,sex=1)+getNAL(obsNAFO3NOPOT,sex=1)+NLmalesOT
OTNafo3NOP_females<-getNAL(portNAFO3NOPOT,sex=2)+getNAL(obsNAFO3NOPOT,sex=2)+NLfemalesOT
OTNafo3NOP_combined<-getNAL(portNAFO3NOPOT,sex='all')+getNAL(obsNAFO3NOPOT,sex='all')+NLmalesOT+NLfemalesOT+NLunknownOT
OTNafo3NOP_unsexed<-getNAL(portNAFO3NOPOT,sex=0)+getNAL(obsNAFO3NOPOT,sex=0)+NLunknownOT

Fishery_Lengths<-list(LLNafo4X5Z_males=LLNafo4X5Z_males,
                      LLNafo4X5Z_females=LLNafo4X5Z_females,
                      LLNafo4X5Z_combined=LLNafo4X5Z_combined,
                      LLNafo4X5Z_unsexed=LLNafo4X5Z_unsexed,
                      
                      LLNafo4VW_males=LLNafo4VW_males,
                      LLNafo4VW_females=LLNafo4VW_females,
                      LLNafo4VW_combined=LLNafo4VW_combined,
                      LLNafo4VW_unsexed=LLNafo4VW_unsexed,
                      
                      LLNafo3NOP_males=LLNafo3NOP_males,
                      LLNafo3NOP_females=LLNafo3NOP_females,
                      LLNafo3NOP_combined=LLNafo3NOP_combined,
                      LLNafo3NOP_unsexed=LLNafo3NOP_unsexed,
                      
                      OTNafo4X5Z_males=OTNafo4X5Z_males,
                      OTNafo4X5Z_females=OTNafo4X5Z_females,
                      OTNafo4X5Z_combined=OTNafo4X5Z_combined, 
                      OTNafo4X5Z_unsexed=OTNafo4X5Z_unsexed,
                      
                      OTNafo4VW_males=OTNafo4VW_males,
                      OTNafo4VW_females=OTNafo4VW_females,
                      OTNafo4VW_combined=OTNafo4VW_combined, 
                      OTNafo4VW_unsexed=OTNafo4VW_unsexed,
                      
                      OTNafo3NOP_males=OTNafo3NOP_males,
                      OTNafo3NOP_females=OTNafo3NOP_females,
                      OTNafo3NOP_combined=OTNafo3NOP_combined, 
                      OTNafo3NOP_unsexed=OTNafo3NOP_unsexed) 

BubblePlotLF(Fishery_Lengths,bins=bins,yrs=CCyrs,path=file.path(wd,'figures'),filen='Commercial',inches=0.1)

```
```{r}
HAST<-read.csv(file.path(datadir,"HASTdataJan2025.csv"))
HAST$rel_date<-as.Date(HAST$rel_date,"%d/%m/%Y")
HAST$rel_date<-as.Date(HAST$rec_date,"%d/%m/%Y")
HAST$rel_lon<-HAST$rel_lon*-1
HAST$rel_lon<-HAST$rec_lon*-1


```


# Put it all together
```{r}

save(BiologicalParameters,NSRV_Index,NSRV_Lengths,NLRV_Index,NLRV_Lengths,HSFixed_Index,HSFixed_Lengths,HSRandom_Index,HSRandom_Lengths,Landings_t,Fishery_Lengths,file=file.path(scaldir,"SCALmodeldata2024.rdata"))

load(file=file.path(scaldir,"SCALmodeldata2024.rdata"))

```


# Bubble plots
```{r}

BubblePlotLF(NSRV_Lengths,bins=bins,yrs=1970:2020,path=file.path(wd,'figures'),filen='NSRVSurvey',inches=0.3,graphic='png')
BubblePlotLF(list(NLRV_Lengths),bins=bins,yrs=1996:2019,path=file.path(wd,'figures'),filen='NLRVSurvey',inches=0.3,graphic='png')
BubblePlotLF(HSFixed_Lengths,bins=bins,yrs=2000:2022,path=file.path(wd,'figures'),filen='FixedSurvey',inches=0.3,graphic='png')
BubblePlotLF(HSRandom_Lengths,bins=bins,yrs=2017:2022,path=file.path(wd,'figures'),filen='RandomSurvey',inches=0.3,graphic='png')
BubblePlotLF(Fishery_Lengths,bins=bins,yrs=1978:2022,path=file.path(wd,'figures'),filen='Commercial',inches=0.3,graphic='png')


```

# bubble for doc
```{r}
combinedLengths<-list(Fishery_Lengths$LLNafo3_combined,Fishery_Lengths$LLNafo4_combined,Fishery_Lengths$OTNafo3_combined,Fishery_Lengths$OTNafo4_combined,NSRV_Lengths$NSRV_combined,NLRV_Lengths$NLRV_combined,HSFixed_Lengths$HSFixed_combined,HSRandom_Lengths$HSRandom_combined)
names(combinedLengths)<-c("LL_NAFO3","LL_NAFO4","OT_NAFO3","OT_NAFO4","RV_4VWX","RV_3NOP","HS_Fixed","HS_Random")

BubblePannelPlotLF(combinedLengths,bins=bins,yrs=1970:2024,path=file.path(wd,'figures'),filen='Combined',inches=0.1,graphic='png',ht=11,wd=8)

# French version
BubblePannelPlotLF(combinedLengths,bins=bins,yrs=1970:2022,path=file.path(wd,'figures'),filen='CombinedFr',inches=0.1,graphic='png',ht=11,wd=8,lang='fr')

```

# Plots for Data Inputs Doc
```{r}
    
    pltYrs1=2000:2022
    pltYrs2=2017:2024
   
    BarPlotLF(FixedLF4X[which(names(FixedLF4X)%in%pltYrs1)],yrs=pltYrs1,rel=F,filen=file.path(wd,"figures","HSLFfixed4X"),graphic='png')
  
    BarPlotLF(FixedLF4VW[which(names(FixedLF4VW)%in%pltYrs1)],yrs=pltYrs1,rel=F,filen=file.path(wd,"figures","HSLFfixed4VW"),graphic='png')
  
    BarPlotLF(FixedLF3NOP[which(names(FixedLF3NOP)%in%pltYrs1)],yrs=pltYrs1,rel=F,filen=file.path(wd,"figures","HSLFfixed3NOP"),graphic='png')
  
    BarPlotLF(RandomLF4X[which(names(RandomLF4X)%in%pltYrs2)],yrs=pltYrs2,rel=F,ht=6,filen=file.path(wd,"figures","HSLFrandom4X"),graphic='png')

    BarPlotLF(RandomLF4VW[which(names(RandomLF4VW)%in%pltYrs2)],yrs=pltYrs2,rel=F,ht=6,filen=file.path(wd,"figures","HSLFrandom4VW"),graphic='png')

    BarPlotLF(RandomLF3NOP[which(names(RandomLF3NOP)%in%pltYrs2)],yrs=pltYrs2,rel=F,ht=6,filen=file.path(wd,"figures","HSLFrandom3NOP"),graphic='png')

    
        #BarPlotLF(FixedLF[which(names(FixedLF)%in%pltYrs1)],yrs=pltYrs1,rel=F,filen=file.path(wd,"figures","HSLFfixed_fr"),graphic='png',xlab="Longueur (cm)",ylab="Nombre de flétans")
  
    #BarPlotLF(RandomLF[which(names(RandomLF)%in%pltYrs2)],yrs=pltYrs2,rel=F,ht=6,filen=file.path(wd,"figures","HSLFrandom_fr"),graphic='png',xlab="Longueur (cm)",ylab="Nombre de flétans")

```
## plot all survey indices used in model ResDoc
```{r}
# switch for language version
lang="en"
#lang="fr"

if(lang=="en"){
  ylab1="Number / tow"
  ylab2="Kg / 1000 hooks"
  options(OutDec = ".")
}

if(lang=="fr"){
  ylab1="Nombre / trait"
  ylab2="Kg / 1 000 hameçons"
  options(OutDec = ",")
}

png(file.path(wd,'figures',paste0("ModelindicesSpatial",lang,".png")), width =8, height = 8,units='in',pointsize=12, res=300,type='cairo')
par(mfcol=c(2,1), mar = c(0,4,0,0.5), omi = c(0.5, 0.25, 0.5, 0.25),las=1)

plot(NPT~year,subset(NSRV_Index,year<2021), ylab=ylab1, xlab = '', ylim=c(0,max(NPT)*1.2), type='b',pch=16,xlim=c(1970,2021),xaxt='n')
axis(1,lab=F)
points(2018,NSRV_Index$NPT[NSRV_Index$year==2018],col='red',cex=2)
#points(2021,NSRV_Index$NPT[NSRV_Index$year==2021],col='red',cex=2)
#lines(mean~survey.year,NLRV_Index,type='b',pch=16,col='green3')
plot(KgPKH~Year,subset(HSFixed_Index,Year<2022),type='b',pch=16,col='red',ylim=c(0,255),xlim=c(1970,2021), ylab=ylab2)
lines(KgPKH~Year,subset(HSRandom_Index,Year<2022),type='b',pch=16,col='blue')
#lines(KgPKH+KgPKHse~Year,HSRandom_Index,lty=2,col='blue')
#lines(KgPKH-KgPKHse~Year,HSRandom_Index,lty=2,col='blue')


dev.off()

```

## plot all survey indices used in model SAR version
```{r}
# switch for language version
lang="en"
#lang="fr"

if(lang=="en"){
  ylab1="Number / tow"
  ylab2="Kg / 1000 hooks"
  options(OutDec = ".")
}

if(lang=="fr"){
  ylab1="Nombre / trait"
  ylab2="Kg / 1 000 hameçons"
  options(OutDec = ",")
}

png(file.path(wd,'figures',paste0("ModelindicesSpatial",lang,".png")), width =8, height = 6,units='in',pointsize=12, res=300,type='cairo')
par(mfcol=c(2,1), mar = c(0,4,0,0.5), omi = c(0.5, 0.25, 0.5, 0.25),las=1)

plot(NPT~year,NSRV_Index4X, ylab=ylab1, xlab = '', ylim=c(0,1.25), type='b', pch=16, xaxt='n', col='blue',xlim=c(1970,2024))
axis(1,lab=F)
points(NPT~year,NSRV_Index4VW,type='b',col='red',pch=15)
points(mean~survey.year,NLRV_Index3NOP,type='b',col='green',pch=17)


plot(KgPKH~Year,HSFixed_Index4X,type='b',pch=1,lty=3,col='blue',ylim=c(0,320),xlim=c(1970,2024), ylab=ylab2)
lines(KgPKH~Year,HSFixed_Index4VW,type='b',pch=0,lty=3,col='red')
lines(KgPKH~Year,HSFixed_Index3NOP,type='b',pch=2,lty=3,col='green')


lines(KgPKH~Year,HSRandom_Index4X,type='b',pch=16,col='blue')
lines(KgPKH~Year,HSRandom_Index4VW,type='b',pch=15,col='red')
lines(KgPKH~Year,HSRandom_Index3NOP,type='b',pch=17,col='green')


dev.off()

```

HCR plot
```{r}


  refs=c(LRP=7.88,  USR=20,  BMSY=25,  BMSY1.2=30)
  
  hcr1=data.frame(B=c(0,refs,40),F=c(0.05,0.05,0.1,0.1,0.1,0.1))
  hcr2=data.frame(B=c(0,refs,40),F=c(0.05,0.05,0.14,0.14,0.14,0.14))
  
  plot(F~B,hcr1,type='l',lwd=2,ylim=c(0,0.15))
  #grid()
  abline(v=refs,lty=2,col=c('red','yellow','green','blue'))

  plot(F~B,hcr2,type='l',lwd=2,ylim=c(0,0.15))
  #grid()
  abline(v=refs,lty=2,col=c('red','yellow','green','blue'))
  
```

