---
title: "NL Observer Data"
author: "Danni Harper"
date: "1/7/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
 library(Mar.datawrangling)
 library(tidyverse)
 library(lubridate)
 library(stringr)
 library(data.table)
 library(SpatialHub)

#datadir="C:/Users/harperd/Documents/Halibut/RDataVault/NLdata_received2021/"
#wd=file.path(datadir,"Newfoundland Data/DataInputs/"
source(file.path(getwd(), "directories.r"))
source(file.path(wd, "passwords.r"))
datadir<-file.path(datadir,"Newfoundland Data")
```


Changing LF data from text file to data frame
```{r}
NLobs_LF24 <- read.fwf(file.path(datadir,"OBS 2024 update/maritimes_Ahal_LF_21-24.lfdata"), widths=c(1,4,2,4,1,3,8,1,2,3,3,2,5,3,4,1,2,2,2,1,3,4,6,4,3,1,4,1,rep((3),times=250)))
                  

colnames(NLobs_LF24) <- c("data_source", "set_no", "country_q", "sample_spp", "sex", "deployment_no", "vessel", "ton_class", "gear", "mesh_size", "NAFO_div", "unit_area", "ASOC", "num_measured", "start_time", "sample_type", "year", "month", "day", "quarter", "country", "depth_m", "total_wt_kg", "sample_wt_kg", "otoliths", "len_gr_int", "spp_sought", "blank",  seq(1,250))

write.csv(NLobs_LF24, file.path(datadir,"OBS 2024 update/NLobserver_lf24.csv"))
NLobs_LF24<-read.csv( file.path(datadir,"OBS 2024 update/NLobserver_lf24.csv"))
```

Changing set info from text file to data frame
```{r}
NLobs_set24 <- read.fwf(file.path(datadir,"OBS 2024 update/maritimes_Ahal_SC_21_24.sdet"), widths=c(2,2,10,4,4,3,2,2,3,3,3,1,3,4,1,1,2,2,2,3,2,4,4,4,4,4,1,4,2,2,2,3,4,1,1,1,4,5,4,1,3,3,3,3,3,3,8,3,5,6,6,6,6,6,5,5,2,2,1,3,5))

colnames(NLobs_set24) <- c("year", "country", "vessel", "hpwr","deployment_no", "set_no","vessel_class", "gear", "codend_mesh", "body_mesh_size", "roller_size", "chafing_gear", "num_gillnets", "num_hooks", "net_damage", "in_out_eez", "day", "month", "NAFO_div", "unit_area", "tow_speed", "lat_ddmm", "long_ddmm", "av_fishing_depth_m", "start_time", "set_dur", "data_type", "directed_spp", "num_spp", "quota", "trip_type", "av_len_gillnet", "num_pots", "window", "num_windows", "record_type", "species", "kept_wt_kg", "discard_wt_kg", "blank1", "num_ind", "blank2", "blank3", "blank4", "blank5", "blank6", "stratarea", "pots", "est_wt_crab", "ASOC", "start_lat_ddmmmm", "start_long_ddmmmm", "end_lat_ddmmmm", "end_long_ddmmmm", "logbook_id", "logbook_id2", "haul_day", "haul_month", "catch_comp", "subtrip_no", "ASO_company")

write.csv(NLobs_set24, file.path(datadir,"OBS 2024 update/NLobserver_setdetail24.csv"))
NLobs_set24<-read.csv(file.path(datadir,"OBS 2024 update/NLobserver_setdetail24.csv"))
```

NAFO divisions and unit areas need fixing for NLobs_set (DO WE NEED UNIT AREAS(subdivisions within NAFO divisions??)
```{r}
NLobs_set24 <- NLobs_set24 %>%
  mutate(NAFO_div= replace(NAFO_div, NAFO_div == 34, "3N")) %>%
  mutate(NAFO_div= replace(NAFO_div, NAFO_div == 35, "3O")) %>%
  mutate(NAFO_div= replace(NAFO_div, NAFO_div == 36, "3Pn")) %>%
  mutate(NAFO_div= replace(NAFO_div, NAFO_div == 37, "3Ps")) %>%
  mutate(NAFO_div= replace(NAFO_div, NAFO_div == 44, "4Vn")) %>%
  mutate(NAFO_div= replace(NAFO_div, NAFO_div == 45, "4Vs")) %>%
  mutate(NAFO_div= replace(NAFO_div, NAFO_div == 46, "4W"))

```

Convert coordinates from ddmm.mm to dec deg (NLobs_set)...position data more available for just ddmm
ddmm.mm = NA for 10,381 out of 20,681
ddmm = NA for 4 out of 20,681
worth it to use less precise to have more available data?
```{r}
#using spatialHub
NLobs_set24$slat_deg<-SpatialHub::convertDecDeg(NLobs_set24$start_lat_ddmmmm)
NLobs_set24$slon_deg<-SpatialHub::convertDecDeg(NLobs_set24$start_long_ddmmmm ) *-1
NLobs_set24$elat_deg<-SpatialHub::convertDecDeg(NLobs_set24$end_lat_ddmmmm )
NLobs_set24$elon_deg<-SpatialHub::convertDecDeg(NLobs_set24$end_long_ddmmmm ) *-1

#column of just degrees

#column that has mmmm

#adding decimal into mm.mm

#dividing mm.mm by 60 to get decimal column

#combining degree and decimal column to get decimal degrees

# plot map to see where they are
bioMap()
with(NLobs_set24,points(slon_deg,slat_deg))

```

DH - Combine set and LF info into 1 useful DF
```{r}
NL_updatedcombined <- NLobs_LF24 %>%
  left_join(NLobs_set24, by = c("year", "deployment_no", "set_no", "vessel"))
```

Check LF against #measured (I assume they created the "# measured" column by just adding the LF by we'll see...)
```{r}
NLobs_LF24$total<-rowSums(NLobs_LF24[,which(names(NLobs_LF24)%in%paste0("X",1:250))],na.rm=T)
NLobs_LF24$total-NLobs_LF24$num_measured
NLobs_LF24$total.undersize<-rowSums(NLobs_LF24[,which(names(NLobs_LF24)%in%paste0("X",1:83))],na.rm=T)

```


Question: there are 20,681 set observations and only 13,440 LF observations. Assuming each row or each table corresponds to a set, where are there more sets than LF? Did they include all sets for all trips that caught halibut (including sets on those trips that didn't catch halibut?)

Question: LF goes to 250...ours goes to 260, do we need the extra columns here to make this work?

Question: do you want it binned like how our other data is binned (right now it's in 1cm intervals)


Data for Brad (3 tables (one for each sex category), years as row, columns are LF bins)

## Otter Trawl

```{r}
NLobs_lf24<-NLobs_LF24
NLobs_malesOT24 <- NLobs_lf24 %>% 
  filter(sex == 1 & gear == 1)
NLobs_femalesOT24 <- NLobs_lf24 %>% 
  filter(sex == 5 & gear == 1)
NLobs_unknownOT24 <- NLobs_lf24 %>%
  filter(is.na(sex) & gear == 1)


#males
NLmalesOT_LF24 <- NLobs_malesOT24 %>% 
  select(-"data_source", -"set_no", -"country_q", -"sample_spp", -"sex", -"deployment_no", -"vessel", -"ton_class", -"gear", -"mesh_size", -"NAFO_div", -"unit_area", -"ASOC", -"num_measured", -"start_time", -"sample_type", -"month", -"day", -"quarter", -"country", -"depth_m", -"sample_wt_kg", -"total_wt_kg", -"otoliths", -"len_gr_int", -"spp_sought", -"blank")

NLmalesOT_LF24 <- NLmalesOT_LF24 %>%
  group_by(year) %>%
  summarise_each(funs(sum))

NLmalesOT_LF24 <- NLmalesOT_LF24 %>%
  mutate(year= replace(year, year == 21, 2021)) %>%
  mutate(year= replace(year, year == 22, 2022)) %>%
  mutate(year= replace(year, year == 23, 2023)) %>%
  mutate(year= replace(year, year == 24, 2024))

#females
NLfemalesOT_LF24 <- NLobs_femalesOT24 %>% 
  select(-"data_source", -"set_no", -"country_q", -"sample_spp", -"sex", -"deployment_no", -"vessel", -"ton_class", -"gear", -"mesh_size", -"NAFO_div", -"unit_area", -"ASOC", -"num_measured", -"start_time", -"sample_type", -"month", -"day", -"quarter", -"country", -"depth_m", -"sample_wt_kg", -"total_wt_kg", -"otoliths", -"len_gr_int", -"spp_sought", -"blank")

NLfemalesOT_LF24 <- NLfemalesOT_LF24 %>%
  group_by(year) %>%
  summarise_each(funs(sum))

NLfemalesOT_LF24 <- NLfemalesOT_LF24 %>%
  mutate(year= replace(year, year == 21, 2021)) %>%
  mutate(year= replace(year, year == 22, 2022)) %>%
  mutate(year= replace(year, year == 23, 2023)) %>%
  mutate(year= replace(year, year == 24, 2024))

##unknown sex
NLunknownOT_LF24 <- NLobs_unknownOT24 %>% 
  select(-"data_source", -"set_no", -"country_q", -"sample_spp", -"sex", -"deployment_no", -"vessel", -"ton_class", -"gear", -"mesh_size", -"NAFO_div", -"unit_area", -"ASOC", -"num_measured", -"start_time", -"sample_type", -"month", -"day", -"quarter", -"country", -"depth_m", -"sample_wt_kg", -"total_wt_kg", -"otoliths", -"len_gr_int", -"spp_sought", -"blank")

NLunknownOT_LF24 <- NLunknownOT_LF24 %>%
  group_by(year) %>%
  summarise_each(funs(sum))

NLunknownOT_LF24 <- NLunknownOT_LF24 %>%
  mutate(year= replace(year, year == 21, 2021)) %>%
  mutate(year= replace(year, year == 22, 2022)) %>%
  mutate(year= replace(year, year == 23, 2023)) %>%
  mutate(year= replace(year, year == 24, 2024))

write.csv(NLmalesOT_LF24, file.path(datadir,"OBS 2024 update/NLmaleOTLF2024.csv"),row.names=F)
write.csv(NLfemalesOT_LF24, file.path(datadir,"OBS 2024 update/NLfemaleOTLF2024.csv"),row.names=F)
write.csv(NLunknownOT_LF24, file.path(datadir,"OBS 2024 update/NLunknownOTLF2024.csv"),row.names=F)
NLunknownOT_LF24<-read.csv(file.path(datadir,"OBS 2024 update/NLunknownOTLF2024.csv"))


##bin by 5cm (0,250) 

## added this in case no observation in a year
    yrs<-2021:2024
    NLmalesOT_LF24 <- merge(NLmalesOT_LF24,data.frame(year=yrs),all=T)
    NLmalesOT_LF24[is.na(NLmalesOT_LF24)]<-0

#male
NLmales2_24 <- transpose(NLmalesOT_LF24, keep.names = "", make.names = "year") 
NLmales2_24 <-NLmales2_24 %>% rename_with( ~ paste0("yr", .x))
NLmales2_24 <- NLmales2_24 %>%
  rename(length = yrV1)

NLmales2_24$length <- as.numeric(NLmales2_24$length)

## changed code here to do it as a loop
    NLmalesOT24<-list()
    for(i in 1:length(yrs)){
      NLmalesOT24[[i]]<-binNumAtLen(num.at.len = NLmales2_24[,paste0("yr",yrs[i])], len = NLmales2_24$length, bins = seq(0, 260, 5))
    }
    NLmalesOT.tab24 <- data.frame(do.call("cbind",NLmalesOT24))
    names(NLmalesOT.tab24)<-yrs
    
## to here

NLmalesOT.tab24 <- tibble::rownames_to_column(NLmalesOT.tab24, "length")

NLmalesOT.tab24$length<-gsub("L"," ", as.character(NLmalesOT.tab24$length))

NLmalesOT.tab24$length <- as.numeric(NLmalesOT.tab24$length)

NLmalesOT.tab24 <- transpose(NLmalesOT.tab24, keep.names = "yr", make.names = "length")

NLmalesOT.tab24$yr <- gsub("yr"," ", as.character(NLmalesOT.tab24$yr))

NLmalesOT.tab24 <- NLmalesOT.tab24 %>%
   rename(year = yr)

write.csv(NLmalesOT.tab24, file.path(datadir,"OBS 2024 update/NLmalesOTobs_binned2024.csv"), row.names = FALSE)

#female


## added this in case no observation in a year
    yrs<-2021:2024
    NLfemalesOT_LF24 <- merge(NLfemalesOT_LF24,data.frame(year=yrs),all=T)
    NLfemalesOT_LF24[is.na(NLfemalesOT_LF24)]<-0

NLfemales2_24 <- transpose(NLfemalesOT_LF24, keep.names = "", make.names = "year") 
NLfemales2_24 <-NLfemales2_24 %>% rename_with( ~ paste0("yr", .x))
NLfemales2_24 <- NLfemales2_24 %>%
  rename(length = yrV1)

NLfemales2_24$length <- as.numeric(NLfemales2_24$length)


## changed code here to do it as a loop
    NLfemalesOT24<-list()
    for(i in 1:length(yrs)){
      NLfemalesOT24[[i]]<-binNumAtLen(num.at.len = NLfemales2_24[,paste0("yr",yrs[i])], len = NLfemales2_24$length, bins = seq(0, 260, 5))
    }
    NLfemalesOT.tab24 <- data.frame(do.call("cbind",NLfemalesOT24))
    names(NLfemalesOT.tab24)<-yrs
    
## to here


NLfemalesOT.tab24 <- tibble::rownames_to_column(NLfemalesOT.tab24, "length")

NLfemalesOT.tab24$length<-gsub("L"," ", as.character(NLfemalesOT.tab24$length))

NLfemalesOT.tab24$length <- as.numeric(NLfemalesOT.tab24$length)

NLfemalesOT.tab24 <- transpose(NLfemalesOT.tab24, keep.names = "yr", make.names = "length")

NLfemalesOT.tab24$yr <- gsub("yr"," ", as.character(NLfemalesOT.tab24$yr))

NLfemalesOT.tab24 <- NLfemalesOT.tab24 %>%
   rename(year = yr)

write.csv(NLfemalesOT.tab24, file.path(datadir,"OBS 2024 update/NLfemalesOTobs_binned2024.csv"), row.names = FALSE)

#unknown sex

## added this in case no observation in a year
    yrs<-2021:2024
    NLunknownOT_LF24 <- merge(NLunknownOT_LF24,data.frame(year=yrs),all=T)
    NLunknownOT_LF24[is.na(NLunknownOT_LF24)]<-0

NLunknown2_24 <- transpose(NLunknownOT_LF24, keep.names = "", make.names = "year") 
NLunknown2_24 <-NLunknown2_24 %>% rename_with( ~ paste0("yr", .x))
NLunknown2_24 <- NLunknown2_24 %>%
  rename(length = yrV1)

NLunknown2_24$length <- as.numeric(NLunknown2_24$length)


## changed code here to do it as a loop
    NLunknownOT24<-list()
    for(i in 1:length(yrs)){
      NLunknownOT24[[i]]<-binNumAtLen(num.at.len = NLunknown2_24[,paste0("yr",yrs[i])], len = NLunknown2_24$length, bins = seq(0, 260, 5))
    }
    NLunknownOT.tab24 <- data.frame(do.call("cbind",NLunknownOT24))
    names(NLunknownOT.tab24)<-yrs
    
## to here

NLunknownOT.tab24 <- tibble::rownames_to_column(NLunknownOT.tab24, "length")

NLunknownOT.tab24$length<-gsub("L"," ", as.character(NLunknownOT.tab24$length))

NLunknownOT.tab24$length <- as.numeric(NLunknownOT.tab24$length)

NLunknownOT.tab24 <- transpose(NLunknownOT.tab24, keep.names = "yr", make.names = "length")

NLunknownOT.tab24$yr <- gsub("yr"," ", as.character(NLunknownOT.tab24$yr))

NLunknownOT.tab24 <- NLunknownOT.tab24 %>%
   rename(year = yr)

write.csv(NLunknownOT.tab24, file.path(datadir,"OBS 2024 update/NLunknownOTobs_binned2024.csv"), row.names = FALSE)
```

## Longline

Data for Brad (3 tables (one for each sex category), years as row, columns are LF bins)
```{r}

NLobs_malesLL24 <- NLobs_lf24 %>% 
  filter(sex == 1 & gear == 7)
NLobs_femalesLL24 <- NLobs_lf24 %>% 
  filter(sex == 5 & gear == 7)
NLobs_unknownLL24 <- NLobs_lf24 %>%
  filter(is.na(sex) & gear == 7)


#males
NLmalesLL_LF24 <- NLobs_malesLL24 %>% 
  select(-"data_source", -"set_no", -"country_q", -"sample_spp", -"sex", -"deployment_no", -"vessel", -"ton_class", -"gear", -"mesh_size", -"NAFO_div", -"unit_area", -"ASOC", -"num_measured", -"start_time", -"sample_type", -"month", -"day", -"quarter", -"country", -"depth_m", -"sample_wt_kg", -"total_wt_kg", -"otoliths", -"len_gr_int", -"spp_sought", -"blank")

NLmalesLL_LF24 <- NLmalesLL_LF24 %>%
  group_by(year) %>%
  summarise_each(funs(sum))

NLmalesLL_LF24 <- NLmalesLL_LF24 %>%
  mutate(year= replace(year, year == 21, 2021)) %>%
  mutate(year= replace(year, year == 22, 2022)) %>%
  mutate(year= replace(year, year == 23, 2023)) %>%
  mutate(year= replace(year, year == 24, 2024)) 

#females
NLfemalesLL_LF24 <- NLobs_femalesLL24 %>% 
  select(-"data_source", -"set_no", -"country_q", -"sample_spp", -"sex", -"deployment_no", -"vessel", -"ton_class", -"gear", -"mesh_size", -"NAFO_div", -"unit_area", -"ASOC", -"num_measured", -"start_time", -"sample_type", -"month", -"day", -"quarter", -"country", -"depth_m", -"sample_wt_kg", -"total_wt_kg", -"otoliths", -"len_gr_int", -"spp_sought", -"blank")

NLfemalesLL_LF24 <- NLfemalesLL_LF24 %>%
  group_by(year) %>%
  summarise_each(funs(sum))

NLfemalesLL_LF24 <- NLfemalesLL_LF24 %>%
    mutate(year= replace(year, year == 21, 2021)) %>%
  mutate(year= replace(year, year == 22, 2022)) %>%
  mutate(year= replace(year, year == 23, 2023)) %>%
  mutate(year= replace(year, year == 24, 2024)) 

##unknown sex
NLunknownLL_LF24 <- NLobs_unknownOT24 %>% 
  select(-"data_source", -"set_no", -"country_q", -"sample_spp", -"sex", -"deployment_no", -"vessel", -"ton_class", -"gear", -"mesh_size", -"NAFO_div", -"unit_area", -"ASOC", -"num_measured", -"start_time", -"sample_type", -"month", -"day", -"quarter", -"country", -"depth_m", -"sample_wt_kg", -"total_wt_kg", -"otoliths", -"len_gr_int", -"spp_sought", -"blank")

NLunknownLL_LF24 <- NLunknownLL_LF24 %>%
  group_by(year) %>%
  summarise_each(funs(sum))

NLunknownLL_LF24 <- NLunknownLL_LF24 %>%
    mutate(year= replace(year, year == 21, 2021)) %>%
  mutate(year= replace(year, year == 22, 2022)) %>%
  mutate(year= replace(year, year == 23, 2023)) %>%
  mutate(year= replace(year, year == 24, 2024)) 

write.csv(NLmalesLL_LF24, file.path(datadir,"OBS 2024 update/NLmaleLLLF2024.csv"))
write.csv(NLfemalesLL_LF24, file.path(datadir,"OBS 2024 update/NLfemaleLLLF2024.csv"))
write.csv(NLunknownLL_LF24, file.path(datadir,"OBS 2024 update/NLunknownLLLF2024.csv"))


##bin by 5cm (0,250) yrs<-1995:2020 NLmalesOT_LF <- merge(NLmalesOT_LF,data.frame(year=yrs),all=T) NLmalesOT_LF[is.na(NLmalesOT_LF)]<-0

#male

## added this in case no observation in a year
    yrs<-2021:2024
    NLmalesLL_LF24 <- merge(NLmalesLL_LF24,data.frame(year=yrs),all=T)
    NLmalesLL_LF24[is.na(NLmalesLL_LF24)]<-0

#male
NLmales2_24 <- transpose(NLmalesLL_LF24, keep.names = "", make.names = "year") 
NLmales2_24 <-NLmales2_24 %>% rename_with( ~ paste0("yr", .x))
NLmales2_24 <- NLmales2_24 %>%
  rename(length = yrV1)

NLmales2_24$length <- as.numeric(NLmales2_24$length)

## changed code here to do it as a loop
    NLmalesLL24<-list()
    for(i in 1:length(yrs)){
      NLmalesLL24[[i]]<-binNumAtLen(num.at.len = NLmales2_24[,paste0("yr",yrs[i])], len = NLmales2_24$length, bins = seq(0, 260, 5))
    }
    NLmalesLL.tab24 <- data.frame(do.call("cbind",NLmalesLL24))
    names(NLmalesLL.tab24)<-yrs
    
## to here
    
NLmalesLL.tab24 <- tibble::rownames_to_column(NLmalesLL.tab24, "length")

NLmalesLL.tab24$length<-gsub("L"," ", as.character(NLmalesLL.tab24$length))

NLmalesLL.tab24$length <- as.numeric(NLmalesLL.tab24$length)

NLmalesLL.tab24 <- transpose(NLmalesLL.tab24, keep.names = "yr", make.names = "length")

NLmalesLL.tab24$yr <- gsub("yr"," ", as.character(NLmalesLL.tab24$yr))

NLmalesLL.tab24 <- NLmalesLL.tab24 %>%
   rename(year = yr)

write.csv(NLmalesLL.tab24, file.path(datadir,"OBS 2024 update/NLmalesLLobs_binned2024.csv"), row.names = FALSE)

#female

## added this in case no observation in a year
    yrs<-2021:2024
    NLfemalesLL_LF24 <- merge(NLfemalesLL_LF24,data.frame(year=yrs),all=T)
    NLfemalesLL_LF24[is.na(NLfemalesLL_LF24)]<-0

#female
NLfemales2_24 <- transpose(NLfemalesLL_LF24, keep.names = "", make.names = "year") 
NLfemales2_24 <-NLfemales2_24 %>% rename_with( ~ paste0("yr", .x))
NLfemales2_24 <- NLfemales2_24 %>%
  rename(length = yrV1)

NLfemales2_24$length <- as.numeric(NLfemales2_24$length)

## changed code here to do it as a loop
    NLfemalesLL24<-list()
    for(i in 1:length(yrs)){
      NLfemalesLL24[[i]]<-binNumAtLen(num.at.len = NLfemales2_24[,paste0("yr",yrs[i])], len = NLfemales2_24$length, bins = seq(0, 260, 5))
    }
    NLfemalesLL.tab24 <- data.frame(do.call("cbind",NLfemalesLL24))
    names(NLfemalesLL.tab24)<-yrs
    
## to here
    
NLfemalesLL.tab24 <- tibble::rownames_to_column(NLfemalesLL.tab24, "length")

NLfemalesLL.tab24$length<-gsub("L"," ", as.character(NLfemalesLL.tab24$length))

NLfemalesLL.tab24$length <- as.numeric(NLfemalesLL.tab24$length)

NLfemalesLL.tab24 <- transpose(NLfemalesLL.tab24, keep.names = "yr", make.names = "length")

NLfemalesLL.tab24$yr <- gsub("yr"," ", as.character(NLfemalesLL.tab24$yr))

NLfemalesLL.tab24 <- NLfemalesLL.tab24 %>%
   rename(year = yr)

write.csv(NLfemalesLL.tab24, file.path(datadir,"OBS 2024 update/NLfemalesLLobs_binned2024.csv"), row.names = FALSE)

#unknown sex

## added this in case no observation in a year
    yrs<-2021:2024
    NLunknownLL_LF24 <- merge(NLunknownLL_LF24,data.frame(year=yrs),all=T)
    NLunknownLL_LF24[is.na(NLunknownLL_LF24)]<-0

#unknown
NLunknown2_24 <- transpose(NLunknownLL_LF24, keep.names = "", make.names = "year") 
NLunknown2_24 <-NLunknown2_24 %>% rename_with( ~ paste0("yr", .x))
NLunknown2_24 <- NLunknown2_24 %>%
  rename(length = yrV1)

NLunknown2_24$length <- as.numeric(NLunknown2_24$length)

## changed code here to do it as a loop
    NLunknownLL24<-list()
    for(i in 1:length(yrs)){
      NLunknownLL24[[i]]<-binNumAtLen(num.at.len = NLunknown2_24[,paste0("yr",yrs[i])], len = NLunknown2_24$length, bins = seq(0, 260, 5))
    }
    NLunknownLL.tab24 <- data.frame(do.call("cbind",NLunknownLL24))
    names(NLunknownLL.tab24)<-yrs
    
## to here
    
NLunknownLL.tab24 <- tibble::rownames_to_column(NLunknownLL.tab24, "length")

NLunknownLL.tab24$length<-gsub("L"," ", as.character(NLunknownLL.tab24$length))

NLunknownLL.tab24$length <- as.numeric(NLunknownLL.tab24$length)

NLunknownLL.tab24 <- transpose(NLunknownLL.tab24, keep.names = "yr", make.names = "length")

NLunknownLL.tab24$yr <- gsub("yr"," ", as.character(NLunknownLL.tab24$yr))

NLunknownLL.tab24 <- NLunknownLL.tab24 %>%
   rename(year = yr)

write.csv(NLunknownLL.tab24, file.path(datadir,"OBS 2024 update/NLunknownLLobs_binned2024.csv"), row.names = FALSE)
```


