---
title: "NL Observer Data"
author: "Danni Harper"
date: "1/7/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
 library(Mar.datawrangling)
 library(tidyverse)
 library(lubridate)
 library(stringr)
 library(data.table)

#datadir="C:/Users/harperd/Documents/Halibut/RDataVault/NLdata_received2021/"
#wd=file.path(datadir,"Newfoundland Data/DataInputs/"
source(file.path(getwd(), "directories.r"))
source(file.path(wd, "passwords.r"))
```


Changing LF data from text file to data frame
```{r}
NLobs_LF <- read.fwf(file.path(datadir,"Newfoundland Data/maritimes_Ahal_LF_95_20.lfdata"), widths=c(1,4,2,4,1,3,8,1,2,3,3,2,5,3,4,1,2,2,2,1,3,4,6,4,3,1,4,1,rep((3),times=250)))
                  

colnames(NLobs_LF) <- c("data_source", "set_no", "country_q", "sample_spp", "sex", "deployment_no", "vessel", "ton_class", "gear", "mesh_size", "NAFO_div", "unit_area", "ASOC", "num_measured", "start_time", "sample_type", "year", "month", "day", "quarter", "country", "depth_m", "total_wt_kg", "sample_wt_kg", "otoliths", "len_gr_int", "spp_sought", "blank",  seq(1,250))

write.csv(NLobs_LF, file.path(datadir,"Newfoundland Data/NLobserver_lf.csv"))
```

Changing set info from text file to data frame
```{r}
NLobs_set <- read.fwf(file.path(datadir,"Newfoundland Data/maritimes_Ahal_SC_95_20.sdet"), widths=c(2,2,10,4,4,3,2,2,3,3,3,1,3,4,1,1,2,2,2,3,2,4,4,4,4,4,1,4,2,2,2,3,4,1,1,1,4,5,4,1,3,3,3,3,3,3,8,3,5,6,6,6,6,6,5,5,2,2,1,3,5))

colnames(NLobs_set) <- c("year", "country", "side_no", "hpwr","deployment_no", "set_no","vessel_class", "gear", "codend_mesh", "body_mesh_size", "roller_size", "chafing_gear", "num_gillnets", "num_hooks", "net_damage", "in_out_eez", "day", "month", "NAFO_div", "unit_area", "tow_speed", "lat_ddmm", "long_ddmm", "av_fishing_depth_m", "start_time", "set_dur", "data_type", "directed_spp", "num_spp", "quota", "trip_type", "av_len_gillnet", "num_pots", "window", "num_windows", "record_type", "species", "kept_wt_kg", "discard_wt_kg", "blank1", "num_ind", "blank2", "blank3", "blank4", "blank5", "blank6", "stratarea", "pots", "est_wt_crab", "ASOC", "start_lat_ddmmmm", "start_long_ddmmmm", "end_lat_ddmmmm", "end_long_ddmmmm", "logbook_id", "logbook_id2", "haul_day", "haul_month", "catch_comp", "subtrip_no", "ASO_company")

write.csv(NLobs_set, file.path(datadir,"Newfoundland Data/NLobserver_setdetail.csv"))
```

NAFO divisions and unit areas need fixing for NLobs_set (DO WE NEED UNIT AREAS(subdivisions within NAFO divisions??)
```{r}
NLobs_set <- NLobs_set %>%
  mutate(NAFO_div= replace(NAFO_div, NAFO_div == 34, "3N")) %>%
  mutate(NAFO_div= replace(NAFO_div, NAFO_div == 35, "3O")) %>%
  mutate(NAFO_div= replace(NAFO_div, NAFO_div == 36, "3Pn")) %>%
  mutate(NAFO_div= replace(NAFO_div, NAFO_div == 37, "3Ps")) %>%
  mutate(NAFO_div= replace(NAFO_div, NAFO_div == 44, "4Vn")) %>%
  mutate(NAFO_div= replace(NAFO_div, NAFO_div == 45, "4Vs"))  ##%>%
  ##mutate(NAFO_div= replace(NAFO_div, NAFO_div == 46, "4W"))

```

Convert coordinates from ddmm.mm to dec deg (NLobs_set)...position data more available for just ddmm
ddmm.mm = NA for 10,381 out of 20,681
ddmm = NA for 4 out of 20,681
worth it to use less precise to have more available data?
```{r}
#column of just degrees
NLobs_set <- NLobs_set %>%
  mutate(lat_deg = substr(lat_ddmm, 1, 2))

#column that has mmmm

#adding decimal into mm.mm

#dividing mm.mm by 60 to get decimal column

#combining degree and decimal column to get decimal degrees


```

Combine set and LF info into 1 useful DF with only useful columns
```{r}

```

Check LF against #measured (I assume they created the "# measured" column by just adding the LF by we'll see...)
```{r}

```

Question: there are 20,681 set observations and only 13,440 LF observations. Assuming reach row or each table corresponds to a set, where are there more sets than LF? Did they include all sets for all trips that caught halibut (including sets on those trips that didn't catch halibut?)

Question: LF goes to 250...ours goes to 260, do we need the extra columns here to make this work?

Question: do you want it binned like how our other data is binned (right now it's in 1cm intervals)


Data for Brad (3 tables (one for each sex category), years as row, columns are LF bins)

## Otter Trawl

```{r}
NLobs_lf<-NLobs_LF
NLobs_malesOT <- NLobs_lf %>% 
  filter(sex == 1 & gear == 1)
NLobs_femalesOT <- NLobs_lf %>% 
  filter(sex == 5 & gear == 1)
NLobs_unknownOT <- NLobs_lf %>%
  filter(is.na(sex) & gear == 1)


#males
NLmalesOT_LF <- NLobs_malesOT %>% 
  select(-"data_source", -"set_no", -"country_q", -"sample_spp", -"sex", -"deployment_no", -"vessel", -"ton_class", -"gear", -"mesh_size", -"NAFO_div", -"unit_area", -"ASOC", -"num_measured", -"start_time", -"sample_type", -"month", -"day", -"quarter", -"country", -"depth_m", -"sample_wt_kg", -"total_wt_kg", -"otoliths", -"len_gr_int", -"spp_sought", -"blank")

NLmalesOT_LF <- NLmalesOT_LF %>%
  group_by(year) %>%
  summarise_each(funs(sum))

NLmalesOT_LF <- NLmalesOT_LF %>%
  mutate(year= replace(year, year == 0, 2000)) %>%
  mutate(year= replace(year, year == 1, 2001)) %>%
  mutate(year= replace(year, year == 2, 2002)) %>%
  mutate(year= replace(year, year == 3, 2003)) %>%
  mutate(year= replace(year, year == 4, 2004)) %>%
  mutate(year= replace(year, year == 5, 2005)) %>%
  mutate(year= replace(year, year == 6, 2006)) %>%
  mutate(year= replace(year, year == 7, 2007)) %>%
  mutate(year= replace(year, year == 8, 2008)) %>%
  mutate(year= replace(year, year == 9, 2009)) %>%
  mutate(year= replace(year, year == 10, 2010)) %>%
  mutate(year= replace(year, year == 11, 2011)) %>%
  mutate(year= replace(year, year == 12, 2012)) %>%
  mutate(year= replace(year, year == 13, 2013)) %>%
  mutate(year= replace(year, year == 14, 2014)) %>%
  mutate(year= replace(year, year == 15, 2015)) %>%
  mutate(year= replace(year, year == 16, 2016)) %>%
  mutate(year= replace(year, year == 17, 2017)) %>%
  mutate(year= replace(year, year == 18, 2018)) %>%
  mutate(year= replace(year, year == 19, 2019)) %>%
  mutate(year= replace(year, year == 20, 2020)) %>%
  mutate(year= replace(year, year == 95, 1995)) %>%
  mutate(year= replace(year, year == 96, 1996)) %>%
  mutate(year= replace(year, year == 97, 1997)) %>%
  mutate(year= replace(year, year == 98, 1998)) %>%
  mutate(year= replace(year, year == 99, 1999)) 

#females
NLfemalesOT_LF <- NLobs_femalesOT %>% 
  select(-"data_source", -"set_no", -"country_q", -"sample_spp", -"sex", -"deployment_no", -"vessel", -"ton_class", -"gear", -"mesh_size", -"NAFO_div", -"unit_area", -"ASOC", -"num_measured", -"start_time", -"sample_type", -"month", -"day", -"quarter", -"country", -"depth_m", -"sample_wt_kg", -"total_wt_kg", -"otoliths", -"len_gr_int", -"spp_sought", -"blank")

NLfemalesOT_LF <- NLfemalesOT_LF %>%
  group_by(year) %>%
  summarise_each(funs(sum))

NLfemalesOT_LF <- NLfemalesOT_LF %>%
  mutate(year= replace(year, year == 0, 2000)) %>%
  mutate(year= replace(year, year == 1, 2001)) %>%
  mutate(year= replace(year, year == 2, 2002)) %>%
  mutate(year= replace(year, year == 3, 2003)) %>%
  mutate(year= replace(year, year == 4, 2004)) %>%
  mutate(year= replace(year, year == 5, 2005)) %>%
  mutate(year= replace(year, year == 6, 2006)) %>%
  mutate(year= replace(year, year == 7, 2007)) %>%
  mutate(year= replace(year, year == 8, 2008)) %>%
  mutate(year= replace(year, year == 9, 2009)) %>%
  mutate(year= replace(year, year == 10, 2010)) %>%
  mutate(year= replace(year, year == 11, 2011)) %>%
  mutate(year= replace(year, year == 12, 2012)) %>%
  mutate(year= replace(year, year == 13, 2013)) %>%
  mutate(year= replace(year, year == 14, 2014)) %>%
  mutate(year= replace(year, year == 15, 2015)) %>%
  mutate(year= replace(year, year == 16, 2016)) %>%
  mutate(year= replace(year, year == 17, 2017)) %>%
  mutate(year= replace(year, year == 18, 2018)) %>%
  mutate(year= replace(year, year == 19, 2019)) %>%
  mutate(year= replace(year, year == 20, 2020)) %>%
  mutate(year= replace(year, year == 95, 1995)) %>%
  mutate(year= replace(year, year == 96, 1996)) %>%
  mutate(year= replace(year, year == 97, 1997)) %>%
  mutate(year= replace(year, year == 98, 1998)) %>%
  mutate(year= replace(year, year == 99, 1999))

##unknown sex
NLunknownOT_LF <- NLobs_unknownOT %>% 
  select(-"data_source", -"set_no", -"country_q", -"sample_spp", -"sex", -"deployment_no", -"vessel", -"ton_class", -"gear", -"mesh_size", -"NAFO_div", -"unit_area", -"ASOC", -"num_measured", -"start_time", -"sample_type", -"month", -"day", -"quarter", -"country", -"depth_m", -"sample_wt_kg", -"total_wt_kg", -"otoliths", -"len_gr_int", -"spp_sought", -"blank")

NLunknownOT_LF <- NLunknownOT_LF %>%
  group_by(year) %>%
  summarise_each(funs(sum))

NLunknownOT_LF <- NLunknownOT_LF %>%
  mutate(year= replace(year, year == 0, 2000)) %>%
  mutate(year= replace(year, year == 1, 2001)) %>%
  mutate(year= replace(year, year == 2, 2002)) %>%
  mutate(year= replace(year, year == 3, 2003)) %>%
  mutate(year= replace(year, year == 4, 2004)) %>%
  mutate(year= replace(year, year == 5, 2005)) %>%
  mutate(year= replace(year, year == 6, 2006)) %>%
  mutate(year= replace(year, year == 7, 2007)) %>%
  mutate(year= replace(year, year == 8, 2008)) %>%
  mutate(year= replace(year, year == 9, 2009)) %>%
  mutate(year= replace(year, year == 10, 2010)) %>%
  mutate(year= replace(year, year == 11, 2011)) %>%
  mutate(year= replace(year, year == 12, 2012)) %>%
  mutate(year= replace(year, year == 13, 2013)) %>%
  mutate(year= replace(year, year == 14, 2014)) %>%
  mutate(year= replace(year, year == 15, 2015)) %>%
  mutate(year= replace(year, year == 16, 2016)) %>%
  mutate(year= replace(year, year == 17, 2017)) %>%
  mutate(year= replace(year, year == 18, 2018)) %>%
  mutate(year= replace(year, year == 19, 2019)) %>%
  mutate(year= replace(year, year == 20, 2020)) %>%
  mutate(year= replace(year, year == 95, 1995)) %>%
  mutate(year= replace(year, year == 96, 1996)) %>%
  mutate(year= replace(year, year == 97, 1997)) %>%
  mutate(year= replace(year, year == 98, 1998)) %>%
  mutate(year= replace(year, year == 99, 1999))

write.csv(NLmalesOT_LF, file.path(datadir,"Newfoundland Data/NLmaleOTLF2022.csv"),row.names=F)
write.csv(NLfemalesOT_LF, file.path(datadir,"Newfoundland Data/NLfemaleOTLF2022.csv"),row.names=F)
write.csv(NLunknownOT_LF, file.path(datadir,"Newfoundland Data/NLunknownOTLF2022.csv"),row.names=F)


##bin by 5cm (0,250) 

## added this in case no observation in a year
    yrs<-1995:2020
    NLmalesOT_LF <- merge(NLmalesOT_LF,data.frame(year=yrs),all=T)
    NLmalesOT_LF[is.na(NLmalesOT_LF)]<-0

#male
NLmales2 <- transpose(NLmalesOT_LF, keep.names = "", make.names = "year") 
NLmales2 <-NLmales2 %>% rename_with( ~ paste0("yr", .x))
NLmales2 <- NLmales2 %>%
  rename(length = yrV1)

NLmales2$length <- as.numeric(NLmales2$length)

## changed code here to do it as a loop
    NLmalesOT<-list()
    for(i in 1:length(yrs)){
      NLmalesOT[[i]]<-binNumAtLen(num.at.len = NLmales2[,paste0("yr",yrs[i])], len = NLmales2$length, bins = seq(0, 260, 5))
    }
    NLmalesOT.tab <- data.frame(do.call("cbind",NLmalesOT))
    names(NLmalesOT.tab)<-yrs
    
## to here

NLmalesOT.tab <- tibble::rownames_to_column(NLmalesOT.tab, "length")

NLmalesOT.tab$length<-gsub("L"," ", as.character(NLmalesOT.tab$length))

NLmalesOT.tab$length <- as.numeric(NLmalesOT.tab$length)

NLmalesOT.tab <- transpose(NLmalesOT.tab, keep.names = "yr", make.names = "length")

NLmalesOT.tab$yr <- gsub("yr"," ", as.character(NLmalesOT.tab$yr))

NLmalesOT.tab <- NLmalesOT.tab %>%
   rename(year = yr)

write.csv(NLmalesOT.tab, file.path(datadir,"Newfoundland Data/NLmalesOTobs_binned2022.csv"), row.names = FALSE)

#female


## added this in case no observation in a year
    yrs<-1995:2020
    NLfemalesOT_LF <- merge(NLfemalesOT_LF,data.frame(year=yrs),all=T)
    NLfemalesOT_LF[is.na(NLfemalesOT_LF)]<-0

NLfemales2 <- transpose(NLfemalesOT_LF, keep.names = "", make.names = "year") 
NLfemales2 <-NLfemales2 %>% rename_with( ~ paste0("yr", .x))
NLfemales2 <- NLfemales2 %>%
  rename(length = yrV1)

NLfemales2$length <- as.numeric(NLfemales2$length)


## changed code here to do it as a loop
    NLfemalesOT<-list()
    for(i in 1:length(yrs)){
      NLfemalesOT[[i]]<-binNumAtLen(num.at.len = NLfemales2[,paste0("yr",yrs[i])], len = NLfemales2$length, bins = seq(0, 260, 5))
    }
    NLfemalesOT.tab <- data.frame(do.call("cbind",NLfemalesOT))
    names(NLfemalesOT.tab)<-yrs
    
## to here


NLfemalesOT.tab <- tibble::rownames_to_column(NLfemalesOT.tab, "length")

NLfemalesOT.tab$length<-gsub("L"," ", as.character(NLfemalesOT.tab$length))

NLfemalesOT.tab$length <- as.numeric(NLfemalesOT.tab$length)

NLfemalesOT.tab <- transpose(NLfemalesOT.tab, keep.names = "yr", make.names = "length")

NLfemalesOT.tab$yr <- gsub("yr"," ", as.character(NLfemalesOT.tab$yr))

NLfemalesOT.tab <- NLfemalesOT.tab %>%
   rename(year = yr)

write.csv(NLfemalesOT.tab, file.path(datadir,"Newfoundland Data/NLfemalesOTobs_binned2022.csv"), row.names = FALSE)

#unknown sex

## added this in case no observation in a year
    yrs<-1995:2020
    NLunknownOT_LF <- merge(NLunknownOT_LF,data.frame(year=yrs),all=T)
    NLunknownOT_LF[is.na(NLunknownOT_LF)]<-0

NLunknown2 <- transpose(NLunknownOT_LF, keep.names = "", make.names = "year") 
NLunknown2 <-NLunknown2 %>% rename_with( ~ paste0("yr", .x))
NLunknown2 <- NLunknown2 %>%
  rename(length = yrV1)

NLunknown2$length <- as.numeric(NLunknown2$length)


## changed code here to do it as a loop
    NLunknownOT<-list()
    for(i in 1:length(yrs)){
      NLunknownOT[[i]]<-binNumAtLen(num.at.len = NLunknown2[,paste0("yr",yrs[i])], len = NLunknown2$length, bins = seq(0, 260, 5))
    }
    NLunknownOT.tab <- data.frame(do.call("cbind",NLunknownOT))
    names(NLunknownOT.tab)<-yrs
    
## to here

NLunknownOT.tab <- tibble::rownames_to_column(NLunknownOT.tab, "length")

NLunknownOT.tab$length<-gsub("L"," ", as.character(NLunknownOT.tab$length))

NLunknownOT.tab$length <- as.numeric(NLunknownOT.tab$length)

NLunknownOT.tab <- transpose(NLunknownOT.tab, keep.names = "yr", make.names = "length")

NLunknownOT.tab$yr <- gsub("yr"," ", as.character(NLunknownOT.tab$yr))

NLunknownOT.tab <- NLunknownOT.tab %>%
   rename(year = yr)

write.csv(NLunknownOT.tab, file.path(datadir,"Newfoundland Data/NLunknownOTobs_binned2022.csv"), row.names = FALSE)
```

## Longline

Data for Brad (3 tables (one for each sex category), years as row, columns are LF bins)
```{r}

NLobs_malesLL <- NLobs_lf %>% 
  filter(sex == 1 & gear == 7)
NLobs_femalesLL <- NLobs_lf %>% 
  filter(sex == 5 & gear == 7)
NLobs_unknownLL <- NLobs_lf %>%
  filter(is.na(sex) & gear == 7)


#males
NLmalesLL_LF <- NLobs_malesLL %>% 
  select(-"data_source", -"set_no", -"country_q", -"sample_spp", -"sex", -"deployment_no", -"vessel", -"ton_class", -"gear", -"mesh_size", -"NAFO_div", -"unit_area", -"ASOC", -"num_measured", -"start_time", -"sample_type", -"month", -"day", -"quarter", -"country", -"depth_m", -"sample_wt_kg", -"total_wt_kg", -"otoliths", -"len_gr_int", -"spp_sought", -"blank")

NLmalesLL_LF <- NLmalesLL_LF %>%
  group_by(year) %>%
  summarise_each(funs(sum))

NLmalesLL_LF <- NLmalesLL_LF %>%
  mutate(year= replace(year, year == 0, 2000)) %>%
  mutate(year= replace(year, year == 1, 2001)) %>%
  mutate(year= replace(year, year == 2, 2002)) %>%
  mutate(year= replace(year, year == 3, 2003)) %>%
  mutate(year= replace(year, year == 4, 2004)) %>%
  mutate(year= replace(year, year == 5, 2005)) %>%
  mutate(year= replace(year, year == 6, 2006)) %>%
  mutate(year= replace(year, year == 7, 2007)) %>%
  mutate(year= replace(year, year == 8, 2008)) %>%
  mutate(year= replace(year, year == 9, 2009)) %>%
  mutate(year= replace(year, year == 10, 2010)) %>%
  mutate(year= replace(year, year == 11, 2011)) %>%
  mutate(year= replace(year, year == 12, 2012)) %>%
  mutate(year= replace(year, year == 13, 2013)) %>%
  mutate(year= replace(year, year == 14, 2014)) %>%
  mutate(year= replace(year, year == 15, 2015)) %>%
  mutate(year= replace(year, year == 16, 2016)) %>%
  mutate(year= replace(year, year == 17, 2017)) %>%
  mutate(year= replace(year, year == 18, 2018)) %>%
  mutate(year= replace(year, year == 19, 2019)) %>%
  mutate(year= replace(year, year == 20, 2020)) %>%
  mutate(year= replace(year, year == 95, 1995)) %>%
  mutate(year= replace(year, year == 96, 1996)) %>%
  mutate(year= replace(year, year == 97, 1997)) %>%
  mutate(year= replace(year, year == 98, 1998)) %>%
  mutate(year= replace(year, year == 99, 1999)) 

#females
NLfemalesLL_LF <- NLobs_femalesLL %>% 
  select(-"data_source", -"set_no", -"country_q", -"sample_spp", -"sex", -"deployment_no", -"vessel", -"ton_class", -"gear", -"mesh_size", -"NAFO_div", -"unit_area", -"ASOC", -"num_measured", -"start_time", -"sample_type", -"month", -"day", -"quarter", -"country", -"depth_m", -"sample_wt_kg", -"total_wt_kg", -"otoliths", -"len_gr_int", -"spp_sought", -"blank")

NLfemalesLL_LF <- NLfemalesLL_LF %>%
  group_by(year) %>%
  summarise_each(funs(sum))

NLfemalesLL_LF <- NLfemalesLL_LF %>%
  mutate(year= replace(year, year == 0, 2000)) %>%
  mutate(year= replace(year, year == 1, 2001)) %>%
  mutate(year= replace(year, year == 2, 2002)) %>%
  mutate(year= replace(year, year == 3, 2003)) %>%
  mutate(year= replace(year, year == 4, 2004)) %>%
  mutate(year= replace(year, year == 5, 2005)) %>%
  mutate(year= replace(year, year == 6, 2006)) %>%
  mutate(year= replace(year, year == 7, 2007)) %>%
  mutate(year= replace(year, year == 8, 2008)) %>%
  mutate(year= replace(year, year == 9, 2009)) %>%
  mutate(year= replace(year, year == 10, 2010)) %>%
  mutate(year= replace(year, year == 11, 2011)) %>%
  mutate(year= replace(year, year == 12, 2012)) %>%
  mutate(year= replace(year, year == 13, 2013)) %>%
  mutate(year= replace(year, year == 14, 2014)) %>%
  mutate(year= replace(year, year == 15, 2015)) %>%
  mutate(year= replace(year, year == 16, 2016)) %>%
  mutate(year= replace(year, year == 17, 2017)) %>%
  mutate(year= replace(year, year == 18, 2018)) %>%
  mutate(year= replace(year, year == 19, 2019)) %>%
  mutate(year= replace(year, year == 20, 2020)) %>%
  mutate(year= replace(year, year == 95, 1995)) %>%
  mutate(year= replace(year, year == 96, 1996)) %>%
  mutate(year= replace(year, year == 97, 1997)) %>%
  mutate(year= replace(year, year == 98, 1998)) %>%
  mutate(year= replace(year, year == 99, 1999))

##unknown sex
NLunknownLL_LF <- NLobs_unknownOT %>% 
  select(-"data_source", -"set_no", -"country_q", -"sample_spp", -"sex", -"deployment_no", -"vessel", -"ton_class", -"gear", -"mesh_size", -"NAFO_div", -"unit_area", -"ASOC", -"num_measured", -"start_time", -"sample_type", -"month", -"day", -"quarter", -"country", -"depth_m", -"sample_wt_kg", -"total_wt_kg", -"otoliths", -"len_gr_int", -"spp_sought", -"blank")

NLunknownLL_LF <- NLunknownLL_LF %>%
  group_by(year) %>%
  summarise_each(funs(sum))

NLunknownLL_LF <- NLunknownLL_LF %>%
  mutate(year= replace(year, year == 0, 2000)) %>%
  mutate(year= replace(year, year == 1, 2001)) %>%
  mutate(year= replace(year, year == 2, 2002)) %>%
  mutate(year= replace(year, year == 3, 2003)) %>%
  mutate(year= replace(year, year == 4, 2004)) %>%
  mutate(year= replace(year, year == 5, 2005)) %>%
  mutate(year= replace(year, year == 6, 2006)) %>%
  mutate(year= replace(year, year == 7, 2007)) %>%
  mutate(year= replace(year, year == 8, 2008)) %>%
  mutate(year= replace(year, year == 9, 2009)) %>%
  mutate(year= replace(year, year == 10, 2010)) %>%
  mutate(year= replace(year, year == 11, 2011)) %>%
  mutate(year= replace(year, year == 12, 2012)) %>%
  mutate(year= replace(year, year == 13, 2013)) %>%
  mutate(year= replace(year, year == 14, 2014)) %>%
  mutate(year= replace(year, year == 15, 2015)) %>%
  mutate(year= replace(year, year == 16, 2016)) %>%
  mutate(year= replace(year, year == 17, 2017)) %>%
  mutate(year= replace(year, year == 18, 2018)) %>%
  mutate(year= replace(year, year == 19, 2019)) %>%
  mutate(year= replace(year, year == 20, 2020)) %>%
  mutate(year= replace(year, year == 95, 1995)) %>%
  mutate(year= replace(year, year == 96, 1996)) %>%
  mutate(year= replace(year, year == 97, 1997)) %>%
  mutate(year= replace(year, year == 98, 1998)) %>%
  mutate(year= replace(year, year == 99, 1999))

write.csv(NLmalesLL_LF, file.path(datadir,"Newfoundland Data/NLmaleLLLF2022.csv"))
write.csv(NLfemalesLL_LF, file.path(datadir,"Newfoundland Data/NLfemaleLLLF2022.csv"))
write.csv(NLunknownLL_LF, file.path(datadir,"Newfoundland Data/NLunknownLLLF2022.csv"))


##bin by 5cm (0,250) yrs<-1995:2020 NLmalesOT_LF <- merge(NLmalesOT_LF,data.frame(year=yrs),all=T) NLmalesOT_LF[is.na(NLmalesOT_LF)]<-0

#male

## added this in case no observation in a year
    yrs<-1995:2020
    NLmalesLL_LF <- merge(NLmalesLL_LF,data.frame(year=yrs),all=T)
    NLmalesLL_LF[is.na(NLmalesLL_LF)]<-0

#male
NLmales2 <- transpose(NLmalesLL_LF, keep.names = "", make.names = "year") 
NLmales2 <-NLmales2 %>% rename_with( ~ paste0("yr", .x))
NLmales2 <- NLmales2 %>%
  rename(length = yrV1)

NLmales2$length <- as.numeric(NLmales2$length)

## changed code here to do it as a loop
    NLmalesLL<-list()
    for(i in 1:length(yrs)){
      NLmalesLL[[i]]<-binNumAtLen(num.at.len = NLmales2[,paste0("yr",yrs[i])], len = NLmales2$length, bins = seq(0, 260, 5))
    }
    NLmalesLL.tab <- data.frame(do.call("cbind",NLmalesLL))
    names(NLmalesLL.tab)<-yrs
    
## to here
    
NLmalesLL.tab <- tibble::rownames_to_column(NLmalesLL.tab, "length")

NLmalesLL.tab$length<-gsub("L"," ", as.character(NLmalesLL.tab$length))

NLmalesLL.tab$length <- as.numeric(NLmalesLL.tab$length)

NLmalesLL.tab <- transpose(NLmalesLL.tab, keep.names = "yr", make.names = "length")

NLmalesLL.tab$yr <- gsub("yr"," ", as.character(NLmalesLL.tab$yr))

NLmalesLL.tab <- NLmalesLL.tab %>%
   rename(year = yr)

write.csv(NLmalesLL.tab, file.path(datadir,"Newfoundland Data/NLmalesLLobs_binned2022.csv"), row.names = FALSE)

#female

## added this in case no observation in a year
    yrs<-1995:2020
    NLfemalesLL_LF <- merge(NLfemalesLL_LF,data.frame(year=yrs),all=T)
    NLfemalesLL_LF[is.na(NLfemalesLL_LF)]<-0

#female
NLfemales2 <- transpose(NLfemalesLL_LF, keep.names = "", make.names = "year") 
NLfemales2 <-NLfemales2 %>% rename_with( ~ paste0("yr", .x))
NLfemales2 <- NLfemales2 %>%
  rename(length = yrV1)

NLfemales2$length <- as.numeric(NLfemales2$length)

## changed code here to do it as a loop
    NLfemalesLL<-list()
    for(i in 1:length(yrs)){
      NLfemalesLL[[i]]<-binNumAtLen(num.at.len = NLfemales2[,paste0("yr",yrs[i])], len = NLfemales2$length, bins = seq(0, 260, 5))
    }
    NLfemalesLL.tab <- data.frame(do.call("cbind",NLfemalesLL))
    names(NLfemalesLL.tab)<-yrs
    
## to here
    
NLfemalesLL.tab <- tibble::rownames_to_column(NLfemalesLL.tab, "length")

NLfemalesLL.tab$length<-gsub("L"," ", as.character(NLfemalesLL.tab$length))

NLfemalesLL.tab$length <- as.numeric(NLfemalesLL.tab$length)

NLfemalesLL.tab <- transpose(NLfemalesLL.tab, keep.names = "yr", make.names = "length")

NLfemalesLL.tab$yr <- gsub("yr"," ", as.character(NLfemalesLL.tab$yr))

NLfemalesLL.tab <- NLfemalesLL.tab %>%
   rename(year = yr)

write.csv(NLfemalesLL.tab, file.path(datadir,"Newfoundland Data/NLfemalesLLobs_binned2022.csv"), row.names = FALSE)

#unknown sex

## added this in case no observation in a year
    yrs<-1995:2020
    NLunknownLL_LF <- merge(NLunknownLL_LF,data.frame(year=yrs),all=T)
    NLunknownLL_LF[is.na(NLunknownLL_LF)]<-0

#unknown
NLunknown2 <- transpose(NLunknownLL_LF, keep.names = "", make.names = "year") 
NLunknown2 <-NLunknown2 %>% rename_with( ~ paste0("yr", .x))
NLunknown2 <- NLunknown2 %>%
  rename(length = yrV1)

NLunknown2$length <- as.numeric(NLunknown2$length)

## changed code here to do it as a loop
    NLunknownLL<-list()
    for(i in 1:length(yrs)){
      NLunknownLL[[i]]<-binNumAtLen(num.at.len = NLunknown2[,paste0("yr",yrs[i])], len = NLunknown2$length, bins = seq(0, 260, 5))
    }
    NLunknownLL.tab <- data.frame(do.call("cbind",NLunknownLL))
    names(NLunknownLL.tab)<-yrs
    
## to here
    
NLunknownLL.tab <- tibble::rownames_to_column(NLunknownLL.tab, "length")

NLunknownLL.tab$length<-gsub("L"," ", as.character(NLunknownLL.tab$length))

NLunknownLL.tab$length <- as.numeric(NLunknownLL.tab$length)

NLunknownLL.tab <- transpose(NLunknownLL.tab, keep.names = "yr", make.names = "length")

NLunknownLL.tab$yr <- gsub("yr"," ", as.character(NLunknownLL.tab$yr))

NLunknownLL.tab <- NLunknownLL.tab %>%
   rename(year = yr)

write.csv(NLunknownLL.tab, file.path(datadir,"Newfoundland Data/NLunknownLLobs_binned2022.csv"), row.names = FALSE)
```


