---
title: "SurveyData"
author: "Brad"
date: "January 12, 2021"
output: html_document
---

```{r setup, include=FALSE}
library(devtools)
#install_github("Maritimes/Mar.datawrangling")
#install_github("Maritimes/Mar.bycatch")
 library(Mar.datawrangling)
 library(Mar.bycatch)
 library(tidyverse)

datadir="C:/Users/harperd/Documents/Halibut/RDataVault"
wd="C:/Users/harperd/Documents/Halibut/GitDataInputs/DataInputs/"
source(file.path(wd, ".passwords"))

```

## Update data from database 

```{r}

# Update data from database 

#get_data(db='isdb',data.dir=datadir,fn.oracle.username = uid, fn.oracle.password = pwd, fn.oracle.dsn = 'ptran',usepkg='roracle',force.extract=T)

get_data(db='isdb',data.dir=datadir,fn.oracle.username = uid, fn.oracle.password = pwd, fn.oracle.dsn = 'ptran',usepkg='roracle',force.extract=T)
get_data(db='rv',data.dir=datadir,fn.oracle.username = uid, fn.oracle.password = pwd, fn.oracle.dsn = 'ptran',usepkg='roracle',force.extract=T)
```

### SUrvey ABundance Indices
## prepare Data for hook occupancy model

```{r }
HookOccupancyData<-PrepareDataHookModel()

```

### Size Composition 
## Prepare data for Length Frequemcy plots

```{r, echo=FALSE}
bins=seq(0,260,5)

FixedSurvey<-FixedSurveyData(datadir=datadir)
RandomSurvey<-RandomSurveyData(datadir=datadir)
RandomSurvey<-RandomSurveyData(sp=30,datadir=datadir,add.gear=T,add.LF=T,bins=seq(5,260,5),by.sex=T,hook.data=T)
ComIndex<-CommercialIndexData(datadir=datadir,bins=seq(0,260,5),add.portsampling=T)
```

## Construct LF for plots
```{r}
        
  #subset by nafo
  #nafos<-unique(FixedSurvey$NAFAREA_ID)
  #nafo4x <- nafos[c(grep('4X',nafos),grep('4V',nafos),grep('4W',nafos))]
  #FixedSurvey4X <-subset(FixedSurvey,NAFAREA_ID%in%nafo4x)
  
  #scale LF
  #FixedSurvey$N_MEASURED <- rowSums(FixedSurvey[,bins])
  #x$scaler = x$NUM_HOOK_HAUL/1000 

  # Construct CLF
  fixedLF=constructLF(LFdata=FixedSurvey,bins=bins, Yrs = 2009:2020)
  RandomLF=constructLF(LFdata=RandomSurvey,bins=bins, Yrs = 2017:2020)
  ComIndexLF=constructLF(LFdata=ComIndex,bins=bins, Yrs = 2009:2020)
<<<<<<< HEAD
  

```


## LF plots
```{r}
  
    BarPlotLF(fixedLF,yrs=Yrs,rel=T,filen=file.path(wd,"figures","fixedLF1"))
    BarPlotLF(fixedLF,yrs=Yrs,rel=F,filen=file.path(wd,"figures","fixedLF2"),ymax=320)
  
    BarPlotLF(RandomLF,yrs=Yrs,rel=T,filen=file.path(wd,"figures","randomLF1"))
    BarPlotLF(RandomLF,yrs=Yrs,rel=F,filen=file.path(wd,"figures","randomLF2"),ymax=150)
    
    BarPlotLF(ComIndexLF,yrs=2009:2020,rel=T,filen=file.path(wd,"figures","comindexLF1"))
    BarPlotLF(ComIndexLF,yrs=2009:2020,rel=F,filen=file.path(wd,"figures","comindexLF2"),ymax=1400)
=======
  
>>>>>>> f09e299c3523e6aa317c352a9135a4dc1d3521af

```

## Checking data for mis match in EST_NUM_CAUGHT and NUM_MEASURED

<<<<<<< HEAD
```{r}

<<<<<<< HEAD
SR20=subset(RandomSurvey,NUM_MEASURED!=EST_NUM_CAUGHT&YEAR==2020)

#FS20=subset(FixedSurvey, NUM_MEASURED!=EST_NUM_CAUGHT&YEAR==2020) 
##maybe they don't have to measure all ones caught in the fixed survey stations?
=======
RandomSurvey$N_MEASURED <- rowSums(RandomSurvey[,paste0("L",bins[-1])])
x20=subset(RandomSurvey,NUM_MEASURED!=EST_NUM_CAUGHT&YEAR==2020)
>>>>>>> f09e299c3523e6aa317c352a9135a4dc1d3521af

SR20 <- SR20 %>%
  select("FISHSET_ID", "SET_NO", "TRIP_ID", "STATION", "EST_NUM_CAUGHT", "EST_COMBINED_WT", "SEXCD_ID", "NUM_MEASURED", "TRIP", "VESS_ID", "VESSEL_NAME", "CFV")
print(SR20)
=======
## LF plots
```{r}
  
    BarPlotLF(fixedLF,yrs=Yrs,rel=T,filen=file.path(wd,"figures","fixedLF1"))
    BarPlotLF(fixedLF,yrs=Yrs,rel=F,filen=file.path(wd,"figures","fixedLF2"),ymax=320)
  
    BarPlotLF(RandomLF,yrs=Yrs,rel=T,filen=file.path(wd,"figures","randomLF1"))
    BarPlotLF(RandomLF,yrs=Yrs,rel=F,filen=file.path(wd,"figures","randomLF2"),ymax=150)
    
    BarPlotLF(ComIndexLF,yrs=2009:2020,rel=T,filen=file.path(wd,"figures","comindexLF1"))
    BarPlotLF(ComIndexLF,yrs=2009:2020,rel=F,filen=file.path(wd,"figures","comindexLF2"),ymax=1400)
>>>>>>> f09e299c3523e6aa317c352a9135a4dc1d3521af

```

Checking length against weight for random survey halibut


```{r}
halibut <- ISSETPROFILE_WIDE %>%
  filter(YEAR == 2020) %>%
  left_join(ISFISHSETS, by = "FISHSET_ID")%>%
  filter(SETCD_ID == 5) %>%
  left_join(ISCATCHES, by = "FISHSET_ID") %>%
  filter(SPECCD_ID == 30) %>%
  left_join(ISFISH, by = "CATCH_ID")

halibut$SEXCD_ID <- as.character(halibut$SEXCD_ID)

ggplot(halibut, aes(x=FISH_LENGTH, fill=SEXCD_ID, color=SEXCD_ID)) +
  geom_histogram(binwidth = 1, position="identity", alpha=0.5) +
  theme_classic()
```

<<<<<<< HEAD
## RV survey

```{r}
get_data(db='rv',data.dir=datadir)
=======
RandomSurvey$N_MEASURED <- rowSums(RandomSurvey[,paste0("L",bins[-1])])
x20=subset(RandomSurvey,NUM_MEASURED!=EST_NUM_CAUGHT&YEAR==2020)
>>>>>>> f09e299c3523e6aa317c352a9135a4dc1d3521af


```

## RV survey

```{r}
get_data(db='rv',data.dir=datadir)


```

