---
title: "SurveyData"
author: "Brad"
date: "January 12, 2021"
output: html_document
---

```{r setup, include=FALSE}
library(devtools)

# install_github("Maritimes/Mar.datawrangling")
# install_github("Maritimes/Mar.fleets")
#install_github("Maritimes/Mar.utils")
 library(Mar.datawrangling)
 #library(Mar.fleets)
 library(tidyverse)
 library(SpatialHub)
library(PBSmappingb)
 library(RColorBrewer)

#install_github("BradHubley/SpatialHub")
source(file.path(getwd(), "directories.r"))
#datadir="C:/Users/hubleyb/Documents/Halibut/data"
#wd="C:/Users/hubleyb/Documents/Halibut/Git/DataInputs/"
datadir="C:/Users/harperd/Documents/Halibut/RDataVault"
wd="C:/Users/harperd/Documents/Halibut/GitDataInputs/DataInputs/"

source(file.path(wd, "passwords.r")) 
ds_all <- Mar.datawrangling::load_datasources()
devtools::load_all(".")
bins=seq(0,260,5)

library(conflicted)
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")

# png(paste0(filen,".png"), width = wd, height = ht,units='in',pointsize=12, res=300,type='cairo')

```

## Update data from database 

```{r}

# Update data from database 

# get_data(db='isdb',data.dir=datadir,fn.oracle.username = uid, fn.oracle.password = pwd, fn.oracle.dsn = 'ptran',usepkg='roracle',force.extract=T)
# open and check setprofile_wide to confirm that you have year, latitude & longitude

# get_data(db='rv',data.dir=datadir,fn.oracle.username = uid, fn.oracle.password = pwd, fn.oracle.dsn = 'ptran',usepkg='roracle',force.extract=T)


#get_data_custom(schema="observer", data.dir = datadir, tables = c("ISFISHLENGTHS","ISSAMPLES"), usepkg = "roracle",  fn.oracle.dsn= "PTRAN", fn.oracle.username=uid, fn.oracle.password =pwd)


#get_data_custom(schema="isdb", data.dir = datadir, tables = c("ISFISHSETGEARGROUPS","ISGEARCOMPONENTS","ISGEARCOMPONENTSBAIT"), usepkg = "roracle", fn.oracle.dsn= "PTRAN", fn.oracle.username=uid, fn.oracle.password =pwd)



```

### SUrvey ABundance Indices
## prepare Data for hook occupancy model

```{r }
#HookOccupancyData<-PrepareDataHookModel(datadir=datadir)
#HookOccupancyDataRaw<-PrepareDataHookModel(datadir=datadir, getrawdata=TRUE)

#sp=220 #dogfish
#sp=15 #cusk
sp=30 #halibut
#sp=10 #cod


RandomSurveyRaph<-RandomSurveyData(sp=sp,datadir=datadir,by.sex=F,add.LF=F, hook.data=F)
RandomSurveyRaph$avg_weight<-RandomSurveyRaph$EST_COMBINED_WT / RandomSurveyRaph$total_target_species
RandomSurveyRaph$STATION <- RandomSurveyRaph$ASSIGNED_STATION
RandomSurveyRaph$DEPTH[RandomSurveyRaph$DEPTH==0]<-NA

colNames1 <- c("FISHSET_ID","SET_NO","TRIP","DATE_TIME3","SOAKMINP3P1","DEPTH","LATITUDE","LONGITUDE","YEAR","STATION","STRATUM_ID","NAFAREA_ID","NUM_HOOK_HAUL","CFV","SPECCD_ID","total_other_species","total_target_species","broken_hook","empty_baited","empty_unbaited","other_species","target_species","missing_hook","total_sampled","avg_weight")

RandomSurvey <- RandomSurveyRaph[,colNames1]
names(RandomSurveyRaph)

FixedSurveyRaph<-FixedSurveyData(sp=sp,datadir=datadir,by.sex=F,add.LF=F, hook.data=T)
FixedSurveyRaph$avg_weight<-FixedSurveyRaph$EST_COMBINED_WT / FixedSurveyRaph$total_target_species
FixedSurveyRaph$DEPTH[FixedSurveyRaph$DEPTH==0]<-NA

 # strata for fixed
 load(file.path(datadir,"Survey","SurveyStrata.rdata")) # These are the original strata (2017-2021)
 FixedSurveyRaph<-reStratify(FixedSurveyRaph,surveyStrataPolyLL,lines=T)
 FixedSurveyRaph$STRATUM_ID<-paste0("H",FixedSurveyRaph$PID)

colNames2 <- c("FISHSET_ID","SET_NO","TRIP","DATE_TIME3","SOAKMINP3P1","DEPTH","LATITUDE","LONGITUDE","YEAR","STATION","STRATUM_ID","NAFAREA_ID","NUM_HOOK_HAUL","CFV","SPECCD_ID","total_other_species","total_target_species","avg_weight")

FixedSurvey <- FixedSurveyRaph[,colNames2]


#write.csv(FixedSurvey,file.path(datadir,"FixedSurveyCusk2.csv"),row.names=F)
write.csv(FixedSurvey,file.path(datadir,"FixedSurveyHalibut.csv"),row.names=F)
write.csv(RandomSurvey,file.path(datadir,"RandomSurveyHalibut.csv"),row.names=F)

FixedSurvey<-read.csv(file.path(datadir,"FixedSurveyHalibut.csv"))
RandomSurvey<-read.csv(file.path(datadir,"RandomSurveyHalibut.csv"))

```

```{r}
FixedSurveyTest<-FixedSurveyData(sp=30,datadir=datadir,by.sex=F,add.LF=F)
```

# data request
```{r}

#species of interest
sp<-c(10, #cod
      15, #cusk
      30, #halibut
      220)#dogfish

HLLRS<-list()#HalibutLongLineRandomSurvey     

for(i in 1:legth(sp)){
  
  
  RS<-RandomSurveyData(sp=sp[i],datadir=datadir,by.sex=F,add.LF=F, hook.data=F)
  RS$STATION <- RS$ASSIGNED_STATION
  RS$DEPTH[RS$DEPTH==0]<-NA
  
  colNames1 <- c("FISHSET_ID","SET_NO","TRIP","DATE_TIME3","SOAKMINP3P1","DEPTH","LATITUDE","LONGITUDE","YEAR","STATION","STRATUM_ID","NAFAREA_ID","NUM_HOOK_HAUL","SPECCD_ID","EST_NUM_CAUGHT", "EST_COMBINED_WT")
  
HLLRS[[i]] <- RS[,colNames1]
#write.csv(HLLRS[[i]],paste0("HalibutLongLineSurveySp",sp[i],".csv"))
}
names(HLLRS)<-sp


```



# data checking random survey
```{r}
RandomSurvey1<-RandomSurveyData(sp=30,datadir=datadir,add.gear=F,add.LF=T,bins=bins,by.sex=F, LF.from = "ISFISHLENGTHS")

RandomSurvey2<-RandomSurveyData(sp=30,datadir=datadir,add.gear=F,add.LF=T,bins=bins,by.sex=F, LF.from = "ISFISH")

RandomSurvey1$diff_count<-with(RandomSurvey1,EST_NUM_CAUGHT -NUM_MEASURED )
RandomSurvey1$diff_weight<-with(RandomSurvey1,EST_COMBINED_WT -calc_weight )

RandomSurvey2$diff_count<-with(RandomSurvey2,EST_NUM_CAUGHT -NUM_MEASURED )
RandomSurvey2$diff_weight<-with(RandomSurvey2,EST_COMBINED_WT -calc_weight )

subset(RandomSurvey1,abs(diff_count)>1&YEAR==2020)
subset(RandomSurvey1,abs(diff_count)>1&YEAR==2021&!is.na(L5))
subset(RandomSurvey2,abs(diff_count)>1&YEAR==2021)

subset(RandomSurvey1,diff_count==0&diff_weight>100)
subset(RandomSurvey2,diff_count==0&diff_weight>100)
```

Making spreadsheet of data what needs to be fixed for Stratified Random Survey (whole time series)
```{r}
#weight from both datasets (using ISFISH and ISFISHLENGTHS)
rs_ISFISH_w <- subset(RandomSurvey2,diff_count==0&diff_weight>100)
rs_ISFISHLENGTHS_w <- subset(RandomSurvey1,diff_count==0&diff_weight>100)
rs_w <- rs_ISFISHLENGTHS_w %>% bind_rows(.,rs_ISFISH_w)
rs_w <- rs_w[!duplicated(rs_w$FISHSET_ID), ]

write.csv(rs_w, "C:/Users/harperd/Documents/Halibut/GitDataInputs/randomsurvey_weightissues_2024.csv", row.names = FALSE)

#length from both datasets (using ISFISH and ISFISHLENGTHS)
rs_ISFISH_l <- subset(RandomSurvey2,abs(diff_count)>0)
rs_ISFISHLENGTHS_l <- subset(RandomSurvey1,abs(diff_count)>0)
rs_l <- rs_ISFISHLENGTHS_l %>% bind_rows(.,rs_ISFISH_l)
rs_l <- rs_l[!duplicated(rs_l$FISHSET_ID), ]

write.csv(rs_l, "C:/Users/harperd/Documents/Halibut/GitDataInputs/randomsurvey_lengthissues_2024.csv", row.names = FALSE)
```


Making spreadsheet of data what needs to be fixed for Fixed Survey (whole time series)
```{r}
#weight from both datasets (using ISFISH and ISFISHLENGTHS)
fs_ISFISH_w <- subset(FixedSurvey2,diff_count==0&diff_weight>100)
fs_ISFISHLENGTHS_w <- subset(FixedSurvey1,diff_count==0&diff_weight>100)
fs_w <- fs_ISFISHLENGTHS_w %>% bind_rows(.,fs_ISFISH_w)
fs_w <- fs_w[!duplicated(fs_w$FISHSET_ID), ]

write.csv(fs_w, "C:/Users/harperd/Documents/Halibut/GitDataInputs/fixedsurvey_weightissues.csv", row.names = FALSE)

#length from both datasets (using ISFISH and ISFISHLENGTHS)
fs_ISFISH_l <- subset(FixedSurvey2,abs(diff_count)>1)
fs_ISFISHLENGTHS_l <- subset(FixedSurvey1,abs(diff_count)>1)
fs_l <- fs_ISFISHLENGTHS_l %>% bind_rows(.,fs_ISFISH_l)
fs_l <- fs_l[!duplicated(fs_l$FISHSET_ID), ]

write.csv(fs_l, "C:/Users/harperd/Documents/Halibut/GitDataInputs/fixedsurvey_lengthissues.csv", row.names = FALSE)
```


# data checking fixed survey
```{r}
FixedSurvey1<-FixedSurveyData(sp=30,datadir=datadir,add.gear=F,add.LF=T,bins=bins,by.sex=F, LF.from = "ISFISHLENGTHS")
FixedSurvey2<-FixedSurveyData(sp=30,datadir=datadir,add.gear=F,add.LF=T,bins=bins,by.sex=F, LF.from = "ISFISH")

FixedSurvey1$diff_count<-with(FixedSurvey1,EST_NUM_CAUGHT -NUM_MEASURED )
FixedSurvey1$diff_weight<-abs(with(FixedSurvey1,EST_COMBINED_WT -calc_weight ))

FixedSurvey2$diff_count<-abs(with(FixedSurvey2,EST_NUM_CAUGHT -NUM_MEASURED ))
FixedSurvey2$diff_weight<-abs(with(FixedSurvey2,EST_COMBINED_WT -calc_weight ))

subset(FixedSurvey1,diff_count>1&YEAR==2021)
subset(FixedSurvey2,diff_count>1&YEAR==2021)

FS_weight1 <- subset(FixedSurvey1,diff_count==0&diff_weight>100)
FS_weight2 <- subset(FixedSurvey2,diff_count==0&diff_weight>100)

```


# data checking fixed survey
```{r}
FixedSurvey1<-FixedSurveyData(sp=30,datadir=datadir,add.gear=F,add.LF=T,bins=bins,by.sex=F, LF.from = "ISFISHLENGTHS")
FixedSurvey2<-FixedSurveyData(sp=30,datadir=datadir,add.gear=F,add.LF=T,bins=bins,by.sex=F, LF.from = "ISFISH")

FixedSurvey1$diff_count<-with(FixedSurvey1,EST_NUM_CAUGHT -NUM_MEASURED )
FixedSurvey1$diff_weight<-abs(with(FixedSurvey1,EST_COMBINED_WT -calc_weight ))

FixedSurvey2$diff_count<-abs(with(FixedSurvey2,EST_NUM_CAUGHT -NUM_MEASURED ))
FixedSurvey2$diff_weight<-abs(with(FixedSurvey2,EST_COMBINED_WT -calc_weight ))

subset(FixedSurvey1,diff_count>1&YEAR==2021)
subset(FixedSurvey2,diff_count>1&YEAR==2021)

subset(FixedSurvey1,diff_count==0&diff_weight>100)

FS_weight_diff <- anti_join(FS_weight1, FS_weight2, by = c("SET_NO", "TRIP"))
FS_weight_diff2 <- anti_join(FS_weight2, FS_weight1, by = c("SET_NO", "TRIP"))
```


### Size Composition 
## Prepare data for Length Frequemcy plots

```{r, echo=FALSE}
bins=seq(0,260,5)

#FixedSurvey<-FixedSurveyData(datadir=datadir,by.sex=T, bins=bins, LF.from = "ISFISH")
FixedSurvey<-FixedSurveyData(datadir=datadir,by.sex=T, bins=bins, LF.from = "ISFISHLENGTHS")
FixedSurvey2<-FixedSurveyData(datadir=datadir,by.sex=F, bins=bins, LF.from = "ISFISHLENGTHS")

RandomSurvey<-RandomSurveyData(sp=30,datadir=datadir,add.gear=F,add.LF=T,bins=bins,by.sex=F,hook.data=F, LF.from = "ISFISH")
RandomSurvey_FL<-RandomSurveyData(sp=30,datadir=datadir,add.gear=F,add.LF=T,bins=bins,by.sex=F,hook.data=F, LF.from = "ISFISHLENGTHS")
RandomSurvey2<-RandomSurveyData(sp=30,datadir=datadir,add.LF=T,bins=bins,by.sex=F)
RandomSurveyAudit<-RandomSurveyData(sp=30,datadir=datadir,add.LF=F,bins=bins,by.sex=F,add.gear = T)

ComIndex<-CommercialIndexData(datadir=datadir,bins=bins,add.portsampling=F)
```

## Map Fixed stations
```{r}
xl=c(-69,-47)
yl=c(40,48)    
stations_100 <- read.csv(file.path(wd,'data',"FixedStation100.csv"))

  pdf(file.path(wd,'figures','FixedStationSurveyMap.pdf'),width=10,height=7)
                         
SpatialHub::bioMap(xlim = xl, ylim = yl, mapRes="MR", title = "Fixed Station Survey")
points(LATITUDE~LONGITUDE,FixedSurvey2,pch=16,col=rgb(0,0,0,0.2),cex=0.5)
points(lat.DecDeg~lon.DecDeg,stations_100,pch=21,bg=rgb(1,1,0,0.8))

dev.off()



```

## Map Random stations
```{r}
xl=c(-69,-47)
yl=c(40,48)    
#stations_100 <- read.csv(file.path(wd,'data',"FixedStation100.csv"))

  #pdf(file.path(wd,'figures','StratfiedRandomSurveyMap2020.pdf'),width=10,height=7)

png(file.path(wd,'figures',"StratfiedRandomSurveyMap2020.png"), width =8, height = 6,units='in',pointsize=9, res=300,type='cairo')

png(file.path(wd,'figures',paste0("StratfiedRandomSurveyMap",yr,".png")), width =8, height = 6,units='in',pointsize=9, res=300,type='cairo')

                         
SpatialHub::bioMap(xlim = xl, ylim = yl, mapRes="MR", title = "Stratified Random Survey",nafo=T)
points(LAT1~LONG1,subset(RandomSurvey,YEAR==2021),pch=21,bg=rgb(1,1,0,1))

dev.off()



```

```{r}
# data
load(file.path(datadir,"Survey","SurveyStrata.RData"))
strata = sort(c(1:5+.1,1:5+.2,1:5+.3))

#NAFO <- rgdal::readOGR("data/Divisions/Divisions.shp")

# colours
cols = c(brewer.pal(3,"Blues"),brewer.pal(3,"Reds"),brewer.pal(3,"Purples"),brewer.pal(3,"Greens"),brewer.pal(3,"PuRd"))
props=data.frame(PID=strata*10,col=cols,border=NA)

# map
png(file.path(wd,'figures',"SurveyStrata2020.png"), width =8, height = 6,units='in',pointsize=9, res=300,type='cairo')
#pdf("SurveyStrata2.pdf", width = 11, height = 7)
bioMap(xlim=c(-68,-47),ylim=c(41,49),poly.lst = list(surveyStrataPolyLL,props),plot.rivers=F,isobaths = NULL,mapRes="MR",nafo=T)
points(LAT1~LONG1,subset(RandomSurvey,YEAR==2020),pch=21,bg=rgb(1,1,0,1))
#plot(NAFO,add=T)
text(coordinates(NAFO)[,1], coordinates(NAFO)[,2],NAFO$ZONE)
dev.off()


```

## Construct LF for plots
```{r}
  Yrs1 = 2000:2020        
  Yrs2 = 2017:2021        
  
  #subset by nafo
  #nafos<-unique(FixedSurvey$NAFAREA_ID)
  #nafo4x <- nafos[c(grep('4X',nafos),grep('4V',nafos),grep('4W',nafos))]
  #FixedSurvey4X <-subset(FixedSurvey,NAFAREA_ID%in%nafo4x)
  
  #scale LF
  #FixedSurvey$N_MEASURED <- rowSums(FixedSurvey[,bins])
  #x$scaler = x$NUM_HOOK_HAUL/1000 


  FixedSurvey<-FixedSurveyData(datadir=datadir,by.sex=T, bins=bins, LF.from = "ISFISHLENGTHS")
  
  # Construct CLF
  fixedLF=constructLF(LFdata=FixedSurvey,bins=bins, Yrs = 1998:2009)
  fixedLF2=constructLF(LFdata=FixedSurvey,bins=bins, Yrs = 2010:2021)
    
  BarPlotLF(fixedLF,yrs=1998:2009,rel=F,filen=file.path(wd,"figures","fixedLF1"),graphic='png')
  BarPlotLF(fixedLF2,yrs=2010:2021,rel=F,filen=file.path(wd,"figures","fixedLF2"),graphic='png')
  
  RandomSurvey<-RandomSurveyData(sp=30,datadir=datadir,add.gear=F,add.LF=T,bins=bins, LF.from = "ISFISHLENGTHS",by.sex=T)
  
  RandomLF=constructLF(LFdata=RandomSurvey,bins=bins, Yrs = Yrs2)
  
    BarPlotLF(RandomLF,yrs=Yrs2,rel=F,filen=file.path(wd,"figures","randomLF1"),graphic='png',ht=6)
  
    
    
    BarPlotLF(fixedLF,yrs=Yrs1,rel=F,filen=file.path(wd,"figures","fixedLF1"),graphic='png')
    
    
 matplot(t(SRfromLF(fixedLF)),pch=1,ylab="Proportion Female")   
   
```


```{r}

  ComIndexLF=constructLF(LFdata=ComIndex,bins=bins, Yrs = Yrs1)

  

```


## LF plots
```{r}

    BarPlotLF(fixedLF,yrs=Yrs1,rel=T,filen=file.path(wd,"figures","fixedLF1"))
    BarPlotLF(fixedLF2,yrs=Yrs1,rel=T,filen=file.path(wd,"figures","fixedLF2"))
    BarPlotLF(fixedLF,yrs=Yrs1,rel=F,filen=file.path(wd,"figures","fixedLF"),ymax=320)
  
    BarPlotLF(RandomLF,yrs=Yrs2,rel=F,filen=file.path(wd,"figures","randomLF1"))
    BarPlotLF(RandomLF,yrs=Yrs2,rel=F,filen=file.path(wd,"figures","randomLF2"),ymax=150)
    
    BarPlotLF(ComIndexLF,yrs=Yrs1,rel=T,filen=file.path(wd,"figures","comindexLF1"))
    BarPlotLF(ComIndexLF,yrs=Yrs1,rel=F,filen=file.path(wd,"figures","comindexLF2"),ymax=1400)


```
## Checking data for mis match in EST_NUM_CAUGHT and NUM_MEASURED

```{r}


SR20=subset(RandomSurvey,NUM_MEASURED!=EST_NUM_CAUGHT&YEAR==2020)

#FS20=subset(FixedSurvey, NUM_MEASURED!=EST_NUM_CAUGHT&YEAR==2020) 
##maybe they don't have to measure all ones caught in the fixed survey stations?

#RandomSurvey$N_MEASURED <- rowSums(RandomSurvey[,paste0("L",bins[-1])])
#x20=subset(RandomSurvey,NUM_MEASURED!=EST_NUM_CAUGHT&YEAR==2020)


SR20 <- SR20 %>%
  select("FISHSET_ID", "SET_NO", "TRIP_ID", "STATION", "EST_NUM_CAUGHT", "EST_COMBINED_WT", "SEXCD_ID", "NUM_MEASURED", "TRIP", "VESS_ID", "VESSEL_NAME", "CFV")
print(SR20)
```

## LF plots
```{r}
  
    BarPlotLF(fixedLF,yrs=Yrs,rel=T,filen=file.path(wd,"figures","fixedLF1"))
    BarPlotLF(fixedLF,yrs=Yrs,rel=F,filen=file.path(wd,"figures","fixedLF2"),ymax=320)
  
    BarPlotLF(RandomLF,yrs=Yrs,rel=T,filen=file.path(wd,"figures","randomLF1"))
    BarPlotLF(RandomLF,yrs=Yrs,rel=F,filen=file.path(wd,"figures","randomLF2"),ymax=150)
    
    BarPlotLF(ComIndexLF,yrs=2009:2020,rel=T,filen=file.path(wd,"figures","comindexLF1"))
    BarPlotLF(ComIndexLF,yrs=2009:2020,rel=F,filen=file.path(wd,"figures","comindexLF2"),ymax=1400)


```

# commercial Index comparison to survey
```{r}
ComIndex<-CommercialIndexData(datadir=datadir,bins=bins,by.sex=F,add.portsampling=F,add.LF = F)
FixedSurvey<-FixedSurveyData(datadir=datadir,by.sex=F, bins=bins, add.LF = F)
RandomSurvey<-RandomSurveyData(sp=30,datadir=datadir,add.gear=F,add.LF=F,bins=bins,by.sex=F,hook.data=F)
```

# plot
```{r}

png(file.path(wd,'figures',"CI.png"), width =6, height = 4,units='in',pointsize=9, res=300,type='cairo')
   # par(mfcol=c(rows,ceiling(length(yrs)/rows)), mar = c(0,2,0,0.5), omi = c(0.85, 0.75, 0.75, 0.5))



ComIndex$KgPKH<-ComIndex$EST_COMBINED_WT/ComIndex$NUM_HOOK_HAUL*1000
CI<-with(ComIndex,tapply(KgPKH,YEAR, mean, na.rm=T))
plot(as.numeric(names(CI)),CI,xlab = '',ylab = 'KgPKH',ylim=c(0,500),type='b',pch=16)

FixedSurvey$KgPKH<-FixedSurvey$EST_COMBINED_WT/FixedSurvey$NUM_HOOK_HAUL*1000
FS<-with(FixedSurvey,tapply(KgPKH,YEAR, mean, na.rm=T))
lines(as.numeric(names(FS)),FS,col='grey',lty=2)


dev.off()

```



```{r}
RandomSurveyHO<-RandomSurveyData(sp=30,datadir=datadir,add.gear=F,add.LF=T,bins=bins,by.sex=F,hook.data=T)

with(RandomSurveyHO,plot((other_species+empty_unbaited)/(other_species+empty_unbaited+empty_baited),avglength,xlab="competition"))
with(RandomSurveyHO,abline(lm(avglength~(other_species+empty_unbaited)/(other_species+empty_unbaited+empty_baited))))

```


## RV survey

```{r}
#RVSurvey<-RVSurveyData(uid, pwd, bins=bins, by.sex=T,use.local=F,datadir=datadir)
RVSurvey<-RVSurveyData(uid, pwd, bins=bins, by.sex=T,use.local=T,datadir=datadir)


rvLF=constructLF(LFdata=RVSurvey,bins=bins, Yrs = Yrs1)
    
BarPlotLF(rvLF,yrs=Yrs1,rel=T,filen=file.path(wd,"figures","RVLF1"))
BarPlotLF(rvLF,yrs=Yrs1,rel=F,filen=file.path(wd,"figures","RVLF2"),ymax=50)


#RVnumbers<-RVindex(uid, pwd, use.local=F,plot=F,datadir=datadir)

RVdata<-get4VWXRV(uid, pwd, use.local=T,datadir=datadir)
RVinputs<-prepRVdata(RVdata,bins=bins,years=1970:2020)

```
## plot RVindex
```{r}
#RVinputs$Index
png(file.path(wd,'figures',"NSRVindex.png"), width =6, height = 8,units='in',pointsize=12, res=300,type='cairo')
par(mfcol=c(3,1), mar = c(0,4,0,0.5), omi = c(0.5, 0.25, 0.5, 0.25),las=1)


plot(N/10^6~year,RVinputs$Index, ylab="Numbers (millions)", xlab = '', ylim=c(0,max(N/10^6)*1.2), type='b',pch=16)
points(2018,RVinputs$Index$N[RVinputs$Index$year==2018]/10^6,col='red',cex=2)
plot(B/10^6~year,RVinputs$Index, ylab="Biomass (kt)", xlab = '', ylim=c(0,max(B/10^6)*1.2), type='b',pch=16)
points(2018,RVinputs$Index$B[RVinputs$Index$year==2018]/10^6,col='red',cex=2)
plot(NPT~year,RVinputs$Index, ylab="Number / tow", xlab = '', ylim=c(0,max(NPT)*1.2), type='b',pch=16)
points(2018,RVinputs$Index$NPT[RVinputs$Index$year==2018],col='red',cex=2)


dev.off()

#RVinputs$Index
png(file.path(wd,'figures',"NSRVindex_fr.png"), width =6, height = 8,units='in',pointsize=12, res=300,type='cairo')
par(mfcol=c(3,1), mar = c(0,4,0,0.5), omi = c(0.5, 0.25, 0.5, 0.25),las=1)


plot(N/10^6~year,RVinputs$Index, ylab="Nombre d’individus (millions)", xlab = '', ylim=c(0,max(N/10^6)*1.2), type='b',pch=16)
points(2018,RVinputs$Index$N[RVinputs$Index$year==2018]/10^6,col='red',cex=2)
plot(B/10^6~year,RVinputs$Index, ylab="Biomasse (kt)", xlab = '', ylim=c(0,max(B/10^6)*1.2), type='b',pch=16)
points(2018,RVinputs$Index$B[RVinputs$Index$year==2018]/10^6,col='red',cex=2)
plot(NPT~year,RVinputs$Index, ylab="Nombre / trait", xlab = '', ylim=c(0,max(NPT)*1.2), type='b',pch=16)
points(2018,RVinputs$Index$NPT[RVinputs$Index$year==2018],col='red',cex=2)


dev.off()


```
# plot RV LF
```{r}
Yrs1 = 1970:1995        
Yrs2 = 1996:2020        

rvLF<-rearrangeLFlist(RVinputs[2:4])

BarPlotLF(rvLF[1:26],yrs=Yrs1,rel=F,filen=file.path(wd,"figures","RVLF1"),rows=13,graphic='png',ax=c(rep(2,13),rep(4,13)),ylp=0.6)

BarPlotLF(rvLF[27:51],yrs=Yrs2,rel=F,filen=file.path(wd,"figures","RVLF2"),rows=13,graphic='png',ax=c(rep(2,13),rep(4,13)),ylp=0.6)

BarPlotLF(rvLF[1:26],yrs=Yrs1,rel=F,filen=file.path(wd,"figures","RVLF1_fr"),rows=13,graphic='png',ax=c(rep(2,13),rep(4,13)),ylp=0.6,xlab="Longueur (cm)",ylab="Nombre de flétans")

BarPlotLF(rvLF[27:51],yrs=Yrs2,rel=F,filen=file.path(wd,"figures","RVLF2_fr"),rows=13,graphic='png',ax=c(rep(2,13),rep(4,13)),ylp=0.6,xlab="Longueur (cm)",ylab="Nombre de flétans")


```


```{r}

xl=c(-69,-47)
yl=c(40,48)    
Yrs1 = 2009:2020        
  Yrs2 = 2017:2020        

for(i in Yrs2){
  
  
	# interpolate abundance
		interp.data<-subset(RandomSurvey2,YEAR==i,c('FISHSET_ID','LONGITUDE','LATITUDE','EST_NUM_CAUGHT'))
		interp.data$EST_NUM_CAUGHT[is.na(interp.data$EST_NUM_CAUGHT)]<-0
		hal.contours<-interpolation(interp.data,ticks='define',place=3,nstrata=5,str.min=0,interp.method='gstat',blank=T,res=0.1,smooth=F,idp=3.5)

		# define contour lines
		print(hal.contours$str.def)
		lvls=c(1, 2, 5, 10, 20, 50, 100)

	  cont.lst<-contourGen(hal.contours$image.dat,lvls,col="YlGn",colorAdj=1)
	  
    pdf(file.path(wd,'figures',paste0('StratfiedRandomSurveyMap',i,'.pdf')),width=10,height=7)
                             
    SpatialHub::bioMap(xlim = xl, ylim = yl,contours=cont.lst, mapRes="MR", title = paste("Stratified Random Survey",i))
    points(LATITUDE~LONGITUDE,subset(RandomSurvey2,YEAR==2020),pch=21,bg=rgb(1,1,0,1))

dev.off()
}

```

## 4VN Sentinel Survey

```{r}


#Sentinel4VN<-Sentinel4VNData(datadir=datadir,bins=bins,by.sex=F,add.LF = F)

Sentinel4VN$KGPHH<-Sentinel4VN$EST_COMBINED_WT/Sentinel4VN$NUM_HOOK_HAUL
SI<-with(Sentinel4VN,tapply(KGPHH,YEAR, mean, na.rm=T))
plot(as.numeric(names(SI)),SI,xlab = '',ylim=c(0,max(SI)),type='b')

totalWT<-with(Sentinel4VN,tapply(EST_COMBINED_WT,YEAR, sum, na.rm=T))
plot(as.numeric(names(totalWT)),totalWT,xlab = '',ylim=c(0,max(totalWT)),type='b')

totalN<-with(Sentinel4VN,tapply(EST_NUM_CAUGHT,YEAR, sum, na.rm=T))
plot(as.numeric(names(totalN)),totalN,xlab = '',ylim=c(0,max(totalN)),type='b')


totalH<-with(Sentinel4VN,tapply(NUM_HOOK_HAUL,YEAR, sum, na.rm=T))
plot(as.numeric(names(totalH)),totalH,xlab = '',ylim=c(0,max(totalH)),type='b')
plot(as.numeric(names(totalH)),totalWT/totalH*1000,xlab = '',ylim=c(0,max(totalWT/totalH*1000)),type='b')

meanWT<-with(Sentinel4VN,tapply(EST_COMBINED_WT,YEAR, mean, na.rm=T))
plot(as.numeric(names(meanWT)),meanWT,xlab = '',ylim=c(0,max(meanWT)),type='b')

meanN<-with(Sentinel4VN,tapply(EST_NUM_CAUGHT,YEAR, mean, na.rm=T))
plot(as.numeric(names(meanN)),meanN,xlab = '',ylim=c(0,max(meanN)),type='b')

bioMap(xlim=c(-60.5,-58.5),ylim=c(45,48))
points(LATITUDE~LONGITUDE,  Sentinel4VN         )


```

# plot for document
```{r}

png(file.path(wd,'figures',"Sentinal.png"), width =6, height = 6,units='in',pointsize=9, res=300,type='cairo')
par(mfcol=c(2,1), mar = c(0,4,0,0.5), omi = c(0.5, 0.25, 0.5, 0.25),las=1)

plot(as.numeric(names(totalH)),totalWT/totalH*1000, ylab="KgPKH", xlab = '', ylim=c(0,max(totalWT/totalH*1000)*1.2), type='b',pch=16)

plot(as.numeric(names(totalH)),totalN/totalH*1000, ylab="NPKH", xlab = '', ylim=c(0,max(totalN/totalH*1000)*1.2), type='b',pch=16)

dev.off()

```


# FS INDEX PLOT for Update

```{r}

HSFixed_Index<-FixedSurveyIndex(datadir=datadir,yrs=1998:2021)
HSFixed_Index2<-FixedSurveyIndex(datadir=datadir,yrs=1998:2021,use.calc.wt=T)

plot(KgPKH~Year,HSFixed_Index,type='b',pch=16,ylim=c(0,max(KgPKH)))
lines(mean3_biomass~Year,HSFixed_Index,col='red')

plot(KgPKH~Year,HSFixed_Index2,type='b',pch=16,ylim=c(0,max(KgPKH)))
lines(mean3_biomass~Year,HSFixed_Index2,col='red')


HSFixed_Index_old<-FixedSurveyIndex(datadir=datadir,yrs=1998:2021,old.model=T)

plot(predicted~year,HSFixed_Index_old,type='b',pch=16)
lines(mean3_biomass~year,HSFixed_Index_old,col='red')

plot(predicted~year,HSFixed_Index_old,type='l',lwd=2,ylim=c(0,250))
for(i in 1:7){
  temp<-FixedSurveyIndex(datadir=datadir,yrs=1998:(2021-i),old.model=T)
  lines(predicted~year,temp,col=i+1)
}

```





# RS INDEX PLOT for Update

```{r}

HSRandom_Index<-StratifiedRandomSurveyIndex(datadir=datadir,yrs=2017:2021)

#HSFixed_Index<-FixedSurveyIndex(datadir=datadir,yrs=2000:2021)

plot(KgPKH~Year,HSFixed_Index,type='l',col='grey',ylim=c(0,250))

lines(KgPKH~Year,HSRandom_Index,type='b',pch=16)
lines(KgPKH+KgPKHse*1.96~Year,HSRandom_Index,lty=2)
lines(KgPKH-KgPKHse*1.96~Year,HSRandom_Index,lty=2)


plot(NPKH~Year,HSFixed_Index,type='l',col='grey',ylim=c(0,21))

lines(NPKH~Year,HSRandom_Index,type='b',pch=16)
lines(NPKH+NPKHse*1.96~Year,HSRandom_Index,lty=2)
lines(NPKH-NPKHse*1.96~Year,HSRandom_Index,lty=2)



```


```{r}

#ggplot() + 
#  geom_line(data = HSFixed_Index, aes(x = Year, y = KgPKH   ), size = 1) + 
#  geom_line(data = HSFixed_Index, aes(x = Year, y = KgPKH - KgPKHse  ), linetype = 2, size = 1) + 
#  geom_line(data = HSFixed_Index, aes(x = Year, y = KgPKH + KgPKHse), linetype = 2, size = 1) +
#  geom_line(aes(x = HSFixed_Index$Year, y = mean3_biomass), size = 1, colour = "red") +  xlim(1998, AY+1) +
#  theme(panel.background = element_rect(fill = "white", colour = "black"), panel.grid = element_blank()) + 
#  ylab("Kg per 1000 hooks") + xlab("Year")

```

```{r}
xl=c(-55,-47)
yl=c(42,46)    
                         
SpatialHub::bioMap(xlim = xl, ylim = yl, mapRes="MR", title = "Stratified Random Survey",nafo=T)
points(LAT1~LONG1,subset(RandomSurvey,YEAR==2019),pch=21,bg=rgb(1,0,0,1))
points(LAT1~LONG1,subset(RandomSurvey,YEAR==2020),pch=21,bg=rgb(1,1,0,1))
points(LAT1~LONG1,subset(RandomSurvey,YEAR==2021),pch=21,bg=rgb(0,1,0,1))


```


# look at sample size effect on SE
```{r}
Ns<-seq(150,250,25)
SEs<-c()

for(i in 1:length(Ns)){
  
 HSRandom_Index<-StratifiedRandomSurveyIndex(datadir=datadir,yrs=2017:2021,nadj=Ns[i]/150)
  SEs[i]<-HSRandom_Index$KgPKHse[1]
}

1-SEs/SEs[1]


 RSindexData <- RandomSurveyData(datadir=datadir, add.LF=F,by.sex=F)
 n<-with(RSindexData,tapply(EST_NUM_CAUGHT,YEAR,length))
 sd<-with(RSindexData,tapply(EST_NUM_CAUGHT,YEAR,sd))
 se<-sd/sqrt(n)
 niter=1000; yrs=2017:2021; sd100<-matrix(NA,niter,length(yrs))
 for(j in 1:niter){
   for(i in 1:length(yrs)){
     x<-subset(RSindexData,YEAR==yrs[i])
     x<-x[sample(1:nrow(x),100),]
     sd100[j,i]<-sd(x$EST_NUM_CAUGHT)
   }
 }
 se100<-colMeans(sd100)/sqrt(100)
 se100
 tabse<-cbind(seq(100,250,50),rbind(colMeans(sd100)/sqrt(100),sd/sqrt(n),sd/sqrt(200),sd/sqrt(250)))
 write.csv(tabse,file.path(datadir,"Survey","RSse.csv"),row.names=F)
 
 tabse<-read.csv(file.path(datadir,"Survey","RSse.csv"))

```

# a look at mean/variance in each strata
```{r}
RSindexData <- RandomSurveyData(datadir=datadir, add.LF=F,by.sex=F)
RSindexData$X <- rowMeans(RSindexData[,c("LONG1", "LONG2", "LONG3", "LONG4")])
RSindexData$Y <- rowMeans(RSindexData[,c("LAT1", "LAT2", "LAT3", "LAT4")])
#RSindexData <- assignStrata(RSindexData)
#RSindexData$ASSIGNED_STRATUM_ID[RSindexData$ASSIGNED_STRATUM_ID=="H33"&RSindexData$PID==34]<-"H34"
#RSindexData$ASSIGNED_STRATUM_ID[RSindexData$ASSIGNED_STRATUM_ID=="H43"&RSindexData$PID==34]<-"H44"

  RSindexData$STRAT <- as.numeric(substr(RSindexData$ASSIGNED_STRATUM_ID,2,3))
  RSindexData$WPKH <-  RSindexData$EST_COMBINED_WT/(RSindexData$NUM_HOOK_HAUL/1000)
  RSindexData$NPKH <-  RSindexData$EST_NUM_CAUGHT/(RSindexData$NUM_HOOK_HAUL/1000)
  yrs<-2017:2021
  RSw<-RSm<-RSv<-list()
  for(i in 1:length(yrs)){
    w<- with(subset(RSindexData,YEAR==yrs[i]),tapply(WPKH,STRAT,mean,na.rm=T))
    m<- with(subset(RSindexData,YEAR==yrs[i]),tapply(NPKH,STRAT,mean,na.rm=T))
    v<- with(subset(RSindexData,YEAR==yrs[i]),tapply(NPKH,STRAT,var,na.rm=T))
    RSw[[i]]<- w
    RSv[[i]]<- v
    RSm[[i]]<- m
    
  }
  
  RSvtab<-do.call("rbind",RSv)
  #round(colMeans(RSvtab))
  
  RSmtab<-do.call("rbind",RSm)
  write.csv(round(RSmtab),file.path(datadir,"Survey","RSN.csv"),row.names=F)
  stratN<-colMeans(RSmtab)
  #saveRDS(stratN,file.path(datadir,"Survey","stratN.rds"))

  
  RSwtab<-do.call("rbind",RSw)
 write.csv(round(RSwtab),file.path(datadir,"Survey","RSW.csv"),row.names=F)



```


# a look at mean/variance in each strata
```{r}
FSindexdata <- FixedSurveyData(sp=30,datadir=datadir,by.sex=F,add.LF=F)
FSindexdata$X <- rowMeans(FSindexdata[,c("LONG1", "LONG2", "LONG3", "LONG4")])
FSindexdata$Y <- rowMeans(FSindexdata[,c("LAT1", "LAT2", "LAT3", "LAT4")])
FSindexdata <- assignStrata(FSindexdata)


  FSindexdata$STRAT <- floor(FSindexdata$PID/10)
  FSindexdata$WPKH <-  FSindexdata$EST_COMBINED_WT/(FSindexdata$NUM_HOOK_HAUL/1000)
  FSindexdata$NPKH <-  FSindexdata$EST_NUM_CAUGHT/(FSindexdata$NUM_HOOK_HAUL/1000)
  yrs<-2011:2021
  FSw<-FSm<-FSv<-list()
  for(i in 1:length(yrs)){
    w<- with(subset(FSindexdata,YEAR==yrs[i]),tapply(WPKH,STRAT,mean))
    m<- with(subset(FSindexdata,YEAR==yrs[i]),tapply(NPKH,STRAT,mean))
    v<- with(subset(FSindexdata,YEAR==yrs[i]),tapply(NPKH,STRAT,var))
    FSw[[i]]<- w
    FSv[[i]]<- v
    FSm[[i]]<- m
    
  }
  
  FSvtab<-do.call("rbind",FSv)
  #round(colMeans(RSvtab))
  
  FSmtab<-do.call("rbind",FSm)
  write.csv(round(FSmtab),file.path(datadir,"Survey","FSN.csv"),row.names=F)
  round(colMeans(FSmtab))
  
  FSwtab<-do.call("rbind",FSw)
  write.csv(round(FSwtab),file.path(datadir,"Survey","FSW.csv"),row.names=F)
  colMeans(FSwtab)
  


```




# a look at mean/variance in each strata
```{r}
CIindexdata <- CommercialIndexData(sp=30,datadir=datadir,by.sex=F,add.LF=F)
#CIindexdata$X <- rowMeans(CIindexdata[,c("LONG1", "LONG2", "LONG3", "LONG4")])
#CIindexdata$Y <- rowMeans(CIindexdata[,c("LAT1", "LAT2", "LAT3", "LAT4")])
#CIindexdata <- assignStrata(CIindexdata)
#CIindexdata$STRAT <- floor(CIindexdata$PID/10)
CIindexdata$STRAT <- NA
CIindexdata$STRAT[substr(CIindexdata$NAFAREA_ID,1,2)=="4X"]<-1
CIindexdata$STRAT[substr(CIindexdata$NAFAREA_ID,1,2)=="4W"]<-2
CIindexdata$STRAT[substr(CIindexdata$NAFAREA_ID,1,2)=="4V"]<-3
CIindexdata$STRAT[substr(CIindexdata$NAFAREA_ID,1,2)=="3P"]<-4
CIindexdata$STRAT[substr(CIindexdata$NAFAREA_ID,1,2)%in%c("3N","3O")]<-5

  CIindexdata$WPKH <-  CIindexdata$EST_COMBINED_WT/(CIindexdata$NUM_HOOK_HAUL/1000)
  CIindexdata$NPKH <-  CIindexdata$EST_NUM_CAUGHT/(CIindexdata$NUM_HOOK_HAUL/1000)
  yrs<-2011:2019
  CIw<-CIm<-CIv<-list()
  for(i in 1:length(yrs)){
    w<- with(subset(CIindexdata,YEAR==yrs[i]),tapply(WPKH,STRAT,mean,na.rm=T))
    m<- with(subset(CIindexdata,YEAR==yrs[i]),tapply(NPKH,STRAT,mean,na.rm=T))
    v<- with(subset(CIindexdata,YEAR==yrs[i]),tapply(NPKH,STRAT,var,na.rm=T))
    CIw[[i]]<- w
    CIv[[i]]<- v
    CIm[[i]]<- m
    
  }
  
  CIvtab<-do.call("rbind",CIv)
  round(colMeans(CIvtab))
  
  CImtab<-do.call("rbind",CIm)
  write.csv(round(CImtab),file.path(datadir,"Survey","CIN.csv"),row.names=F)
  round(colMeans(CImtab))
  
  CIwtab<-do.call("rbind",CIw)
  write.csv(round(CIwtab),file.path(datadir,"Survey","CIW.csv"),row.names=F)
  colMeans(CIwtab)
  


```

```{r}
RSindexData <- RandomSurveyData(datadir=datadir, add.LF=F,by.sex=F)
RSindexData$STRAT <- as.numeric(substr(RSindexData$ASSIGNED_STRATUM_ID,2,3))
RSindexData$X <- rowMeans(RSindexData[,c("LONG1", "LONG2", "LONG3", "LONG4")])
RSindexData$Y <- rowMeans(RSindexData[,c("LAT1", "LAT2", "LAT3", "LAT4")])

    #load(file.path(datadir,"Survey","SurveyStrata.rdata"))
    #polys<-surveyStrataPolyLL

  
out1<-assignStrata(RSindexData,coords=c("X","Y"))
out1<-subset(out1,STRAT!=PID)
nrow(out1)


out2<-assignStrata(RSindexData, polys=surveyStrataPolyLL,coords=c("LONG2","LAT2"))
out2<-subset(out2,STRAT!=PID)
nrow(out2)


out3<-assignStrata(RSindexData, polys=surveyStrataPolyLL,coords=c("LONG3","LAT3"))
out3<-subset(out3,STRAT!=PID)
nrow(out3)


out4<-assignStrata(RSindexData, polys=surveyStrataPolyLL,coords=c("LONG4","LAT4"))
out4<-subset(out4,STRAT!=PID)
nrow(out4)


```

# plot stations
```{r}

    # map data
    load(file.path(datadir,"Survey","SurveyStrata2022.RData"))
    #strata = sort(c(1:5+.1,1:5+.2,1:5+.3))
    strata = sort(c(1:5+.1,1:5+.2,1:5+.3,3:4+.4))
    
    #NAFO <- rgdal::readOGR("data/Divisions/Divisions.shp")
    
    # colours
    cols = c(brewer.pal(3,"Blues"),
             brewer.pal(3,"Reds"),
             brewer.pal(3,"Purples"),
             brewer.pal(6,"Purples")[6],
             brewer.pal(3,"Greens"),
             brewer.pal(6,"Greens")[6],
             brewer.pal(3,"PuRd"))
    #den=c(rep(0,9),30,rep(0,3),30,rep(0,3))
    props=data.frame(PID=strata*10,col=cols)
    
    yrs<= 2017:2021
    
# which sets to plot
for(y in yrs){
sets<-subset(RSindexData,YEAR==y)$FISHSET_ID
pdf(file.path(wd,'figures',paste0("SurveySets",y,".pdf")),height=9)

for(i in 1:length(sets)){
  
  d<-subset(RSindexData,FISHSET_ID==sets[i])
  y<-with(d,c(LAT1, LAT2, LAT3, LAT4))
  x<-with(d,c(LONG1, LONG2, LONG3, LONG4))

  bioMap(xlim=c(min(x)-0.1,max(x)+0.1), ylim=c(min(y)-0.1,max(y)+0.1), isobaths=seq(100,1000,10), poly.lst=list(surveyStrataPolyLL,props), plot.rivers=F, mapRes="UR", LT=T)
  title(main=paste("Station #", d$STATION),sub=paste("Assigned Strata", d$STRAT), col.sub=props$col[props$PID==d$STRAT], cex.sub=0.9)
  lines(x,y,type='o',pch=16)
  points(x[1],y[1],pch=21,cex=1.5,bg='red')
  points(d$X,d$Y,pch=21,cex=1.5,bg='yellow')
  

  }

dev.off()
}

```

