---
title: "SurveyMetrics"
author: "Danni Harper"
date: "11/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE)
 library(Mar.datawrangling)
 library(tidyverse)
 library(stringr)

datadir="C:/Users/harperd/Documents/Halibut/RDataVault"
wd="C:/Users/harperd/Documents/Halibut/GitDataInputs/DataInputs/"
#source(file.path(wd, "passwords.r"))
```

Pulling the ISDB to access the survey data
```{r}
get_data("isdb", data.dir = datadir)
```
Separating out just halibut survey data (from fixed, stratified random, commercial index)
```{r}
haltrips <- ISTRIPS %>%
  filter(TRIPCD_ID %in% c(7057, 7058))

halsets <- ISFISHSETS %>%
  filter(SETCD_ID %in% c(4,5,10), SPECSCD_ID == 30)

halsets <- halsets %>%
  left_join(ISSETPROFILE_WIDE, by = "FISHSET_ID")

halsets <- halsets %>%
  left_join(haltrips, by = "TRIP_ID")

halsets <- halsets %>%
  mutate(NAFO = substr(NAFAREA_ID, 1, 2))

halsets <- halsets %>%
  left_join(ISGEARS, by = "GEAR_ID")

####SHOULD BE USING SOAKTIME 3-1???
halsets$SOAKMINP3P1 <- difftime(halsets$DATE_TIME3,  halsets$DATE_TIME1, units='min')

#only until 2020
halsets <- halsets %>% filter(YEAR < 2021)
```

separating the 3 different survey types
```{r}
sets_fs1 <- halsets %>%
  filter(SETCD_ID == 4)

sets_fs1 %>% filter(YEAR %in% c(1998, 1999))

sets_sr1 <- halsets %>%
  filter(SETCD_ID == 5)

sets_ci1 <- halsets %>%
  filter(SETCD_ID == 10)
```

Pulling the data
```{r}
sets_sr <- RandomSurveyData(sp=30, datadir, add.gear=T, add.LF=F, by.sex=F, hook.data=F)
sets_sr <- sets_sr %>%
  mutate(NAFO = substr(NAFAREA_ID, 1, 2))
sets_sr <- sets_sr %>% filter(YEAR < 2021)

sets_fs <- FixedSurveyData(sp=30, datadir, add.gear=T, add.LF=F, by.sex=F,correct.splitsets=T)
sets_fs <- sets_fs %>%
  mutate(NAFO = substr(NAFAREA_ID, 1, 2))
sets_fs <- sets_fs %>% filter(YEAR < 2021) %>% filter(YEAR > 2000)

sets_ci <- CommercialIndexData(sp=30, datadir, add.gear=T, add.LF=F,by.sex=T, add.portsampling=F)
sets_ci <- sets_ci %>%
  mutate(NAFO = substr(NAFAREA_ID, 1, 2))
sets_ci <- sets_ci %>% filter(YEAR < 2021)
```

Checking difference in pulls
```{r}
library(sqldf)

#difference between my pull and Brad's for Fixed
sr1 <- sets_sr1 %>% select("FISHSET_ID", "TRIP_ID.x", "SETCD_ID", "TRIPCD_ID", "STATION", "NAFAREA_ID", "HAULCCD_ID", "YEAR", "TRIP") %>%
  rename(TRIP_ID = TRIP_ID.x)
sr <- sets_sr %>% select("FISHSET_ID", "TRIP_ID","SETCD_ID", "TRIPCD_ID","STATION", "NAFAREA_ID", "HAULCCD_ID", "YEAR", "TRIP")
sr1NotInsr <- sqldf('SELECT * FROM sr1 EXCEPT SELECT * FROM sr')

#difference between my pull and Brad's for Stratified Random
fs1 <- sets_fs1 %>% select("FISHSET_ID", "TRIP_ID.x", "NUM_HOOK_HAUL", "SETCD_ID", "TRIPCD_ID", "STATION", "NAFAREA_ID", "HAULCCD_ID", "YEAR", "TRIP") %>%
  rename(TRIP_ID = TRIP_ID.x)
fs <- sets_fs %>% select("FISHSET_ID", "TRIP_ID","NUM_HOOK_HAUL", "SETCD_ID", "TRIPCD_ID","STATION", "NAFAREA_ID", "HAULCCD_ID", "YEAR", "TRIP")
fs1NotInfs <- sqldf('SELECT * FROM fs1 EXCEPT SELECT * FROM fs')

#difference between my pull and Brad's for Commercial Index
ci1 <- sets_ci1 %>% select("FISHSET_ID", "TRIP_ID.x", "SETCD_ID", "TRIPCD_ID", "STATION", "NAFAREA_ID", "YEAR", "TRIP") %>%
  rename(TRIP_ID = TRIP_ID.x)
ci <- sets_ci %>% select("FISHSET_ID", "TRIP_ID","SETCD_ID", "TRIPCD_ID","STATION", "NAFAREA_ID", "YEAR", "TRIP")
ci1NotInci <- sqldf('SELECT * FROM ci1 EXCEPT SELECT * FROM ci')#16 differences and don't know why, all from single trip
ciNotInci1 <- sqldf('SELECT * FROM ci EXCEPT SELECT * FROM ci1')#1 different and don't know why, maybe b/c no year?
```


Checking correct species (30)
```{r}
halsets %>% count(SPECSCD_ID)
```

Checking for port sampled (7058) vs sampled at sea (7057)
```{r}
sets_fs %>% count(TRIPCD_ID)
sets_sr %>% count(TRIPCD_ID)
sets_ci %>% count(TRIPCD_ID)

#SR station with trip code 30
sets_sr1 %>% filter(TRIPCD_ID == 30)
#from 2002...looks like wrong SETCD_ID (should be 10 or 1?)
```

Checking that all set codes are right by checking the year...?
```{r}
sets_fs %>% count(YEAR)
#2009 entry station 274 has no set info...what to do with that point? (trip:100015694, set:100061377) NOT in ISSETPROFILE_WIDE...why??
sets_sr %>% count(YEAR)
#station coming from 2002 appears to be a mistake. This is the only set from trip 100014065. Should is even be a halibut trip? (Trip: J02-0208)
sets_ci %>% count(YEAR)
#2011 entry has no set info...(set:100160079, trip:100028073) not in ISSETPROFILE_WIDE...why?
```
NAFO
```{r}
halsets %>% count(NAFO)
#10 stations come up with no NAFO
#1 station comes up in 4R (2018, trip: 100051912, it's commercial index and the rest of the sets are in 3N)
#14 stations come up in 4T (2013: station 316...the rest are commercial index over range of time 2007-2021)
##assume I remove all of these?

halsets %>% filter(NAFO == "4T")
```
adding soaktime
```{r}
#sets_ci$SOAKMINP3P1 <- difftime(sets_ci$DATE_TIME3,  sets_ci$DATE_TIME1, units='min')
```


Stations fished by year and NAFO
```{r}
nafotab_fs <- sets_fs %>%
  dplyr::select("YEAR", "NAFO") %>%
  group_by(YEAR, NAFO) %>%
  summarize(n=n())
nafotab_fs <- spread(nafotab_fs, NAFO, n)

nafotab_fs1 <- sets_fs1 %>%
  dplyr::select("YEAR", "NAFO") %>%
  group_by(YEAR, NAFO) %>%
  summarize(n=n())
nafotab_fs1 <- spread(nafotab_fs1, NAFO, n)

nafotab_sr <- sets_sr %>%
  dplyr::select("YEAR", "NAFO") %>%
  group_by(YEAR, NAFO) %>%
  summarize(n=n())
nafotab_sr <- spread(nafotab_sr, NAFO, n)

nafotab_ci <- sets_ci %>%
  dplyr::select("YEAR", "NAFO") %>%
  group_by(YEAR, NAFO) %>%
  summarize(n=n())
nafotab_ci <- spread(nafotab_ci, NAFO, n)

nafotab <- halsets %>%
  dplyr::select("YEAR", "NAFO") %>%
  group_by(YEAR, NAFO) %>%
  summarize(n=n())
nafotab <- spread(nafotab, NAFO, n)
```
Stations fished by hooksize by NAFO
```{r}
hooknafotab_fs <- sets_fs %>%
  dplyr::select("NAFO", "HOOKSIZE") %>%
  group_by(NAFO, HOOKSIZE) %>%
  summarize(n=n())
hooknafotab_fs <- spread(hooknafotab_fs, HOOKSIZE, n)

sets_fs1_early_hooks <- sets_fs1_early%>% 
  dplyr::select("NAFO", "HOOKSIZE") %>%
  group_by(NAFO, HOOKSIZE) %>%
  summarize(n=n())
sets_fs1_early_hooks <- spread(sets_fs1_early_hooks, HOOKSIZE, n)

  
hooknafotab_sr <- sets_sr %>%
  dplyr::select("NAFO", "HOOKSIZE") %>%
  group_by(NAFO, HOOKSIZE) %>%
  summarize(n=n())
hooknafotab_sr <- spread(hooknafotab_sr, HOOKSIZE, n)

hooknafotab_ci <- sets_ci %>%
  dplyr::select("NAFO", "HOOKSIZE") %>%
  group_by(NAFO, HOOKSIZE) %>%
  summarize(n=n())
hooknafotab_ci <- spread(hooknafotab_ci, HOOKSIZE, n)

hooknafotab <- halsets %>%
  dplyr::select("NAFO", "HOOKSIZE") %>%
  group_by(NAFO, HOOKSIZE) %>%
  summarize(n=n())
hooknafotab <- spread(hooknafotab, HOOKSIZE, n)
```
Stations fished by hooksize by YEAR
```{r}
hookyeartab_fs <- sets_fs %>%
  dplyr::select("YEAR", "HOOKSIZE") %>%
  group_by(YEAR, HOOKSIZE) %>%
  summarize(n=n())
hookyeartab_fs <- spread(hookyeartab_fs, HOOKSIZE, n)

hookyeartab_sr <- sets_sr %>%
  dplyr::select("YEAR", "HOOKSIZE") %>%
  group_by(YEAR, HOOKSIZE) %>%
  summarize(n=n())
hookyeartab_sr <- spread(hookyeartab_sr, HOOKSIZE, n)

hookyeartab_ci <- sets_ci %>%
  dplyr::select("YEAR", "HOOKSIZE") %>%
  group_by(YEAR, HOOKSIZE) %>%
  summarize(n=n())
hookyeartab_ci <- spread(hookyeartab_ci, HOOKSIZE, n)

hookyeartab <- halsets %>%
  dplyr::select("YEAR", "HOOKSIZE") %>%
  group_by(YEAR, HOOKSIZE) %>%
  summarize(n=n())
hookyeartab <- spread(hookyeartab, HOOKSIZE, n)
```

looking at sea observer or at sea observed by year
```{r}
observedyear_fs <- sets_fs %>%
  dplyr::select("YEAR", "TRIPCD_ID") %>%
  group_by(YEAR, TRIPCD_ID) %>%
  summarize(n=n())
observedyear_fs <- spread(observedyear_fs, TRIPCD_ID, n)
#all fixed stations are observed at sea(trip code 7057)

observedyear_sr <- sets_sr %>%
  dplyr::select("YEAR", "TRIPCD_ID") %>%
  group_by(YEAR, TRIPCD_ID) %>%
  summarize(n=n())
observedyear_sr <- spread(observedyear_sr, TRIPCD_ID, n)
#all but 1 stratified random stations are observed at sea. Other is trip cd 30...likely from setcd_id (shouldn't be 5)

observedyear_ci <- sets_ci %>%
  dplyr::select("YEAR", "TRIPCD_ID") %>%
  group_by(YEAR, TRIPCD_ID) %>%
  summarize(n=n())
observedyear_ci <- spread(observedyear_ci, TRIPCD_ID, n)

observedyear_ci <- observedyear_ci %>% rename(ASO = "7057", PSO = "7058")
observedyear_ci <- observedyear_ci %>% mutate(percent_ASO = ((ASO/PSO)*100))
mean(observedyear_ci$percent_ASO, na.rm=TRUE)
```

looking at shore observed or at sea observed by NAFO (all fixed and stratified random station are observed at sea so only looking at the commercial index stations)
```{r}
observednafo_ci <- sets_ci %>%
  dplyr::select("NAFO", "TRIPCD_ID") %>%
  group_by(NAFO, TRIPCD_ID) %>%
  summarize(n=n())
observednafo_ci <- spread(observednafo_ci, TRIPCD_ID, n)
#8 stations in 3N in 2018 with no at sea observer

observedboth_ci <- sets_ci %>%
  dplyr::select("NAFO", "YEAR", "TRIPCD_ID") %>%
  group_by(NAFO, YEAR, TRIPCD_ID) %>%
  summarize(n=n())
observedboth_ci <- spread(observedboth_ci, TRIPCD_ID, n)

##for table 22
observed <- sets_ci %>%
  dplyr::select("NAFO", "YEAR", "TRIPCD_ID") %>%
  group_by(NAFO, YEAR, TRIPCD_ID) %>% 
  summarize(n=n())
observed <- spread(observed, NAFO, n)
```

how often have stations been fished over the time series
```{r}
stations_over_time <- halsets %>%
  dplyr::select("STATION") %>%
  group_by(STATION) %>%
  summarize(n=n())

stations_over_time$STATION <- as.numeric(stations_over_time$STATION)

#station fished more than 5 times...?
halsets %>% filter(STATION == 512.0) #fished twice in 2018, one coded commercial index (check J18-0203, set 7(CI), check J18-0213 set 1)

```

histograms of soak time for the different survey types
```{r}
#FIXED STATION
ggplot(sets_fs, aes(DUR_41, stat = "count")) +
  geom_histogram(position = "identity", alpha = 0.2) +
  theme_classic()
#some durations below 0...meaning times were input wrong. Remove these?

sets_fs %>% filter(DUR_41 < 0) #looks like all are a case of the end time (DATE_TIME4) happening at or after midnight and using the same date as the set time (not the next day like it should be) 

sets_fs_dur <- sets_fs %>%
  filter(DUR_41 > 0)

sets_fs_dur %>% filter(DUR_41 < 320) #lots of these have low hook numbers (some below 100)

#STRATIFIED RANDOM
ggplot(sets_sr, aes(DUR_41, stat = "count")) +
  geom_histogram(position = "identity", alpha = 0.2) +
  theme_classic() 

##one station very low. Looking into it
sets_sr %>% filter(DUR_41 <320) %>% select(DATE_TIME4) #again a case of where the set ended after midnight but use the same date

#COMMERCIAL INDEX (most don't have a date time for 4)
ggplot(ci_dur_zr, aes(DUR_32, stat = "count")) +
  geom_histogram(position = "identity", alpha = 0.2) +
  theme_classic()

####SHOULD BE USING SOAKTIME 3-1???
ggplot(sets_fs, aes(SOAKMINP3P1, stat = "count")) +
  geom_histogram(position = "identity", alpha = 0.2) +
  theme_classic()

#just some kind of estimate...
soaktime_ci <- sets_ci %>% 
  filter(DATE_TIME3 <= as.Date("9998-01-01"), SOAKMINP3P1 > 0)
mean(soaktime_ci$SOAKMINP3P1)
median(soaktime_ci$SOAKMINP3P1)
```
Graphing histograms of hooks
```{r}
ggplot(sets_fs, aes(NUM_HOOK_HAUL, stat = "count")) +
  geom_histogram(position = "identity", alpha = 0.2) +
  theme_classic()

ggplot(sets_sr, aes(NUM_HOOK_HAUL, stat = "count")) +
  geom_histogram(position = "identity", alpha = 0.2) +
  theme_classic()
#looking at stations with less than 975
sets_sr %>% filter(NUM_HOOK_HAUL < 975)
#the one that's likely the wrong set code has 910 hooks
#most concerning is set with 450 hooks (2017, trip S17-0701, set 5, station 604)

ggplot(sets_ci, aes(NUM_HOOK_HAUL, stat = "count")) +
  geom_histogram(position = "identity", alpha = 0.2) +
  theme_classic()

mean(sets_ci$NUM_HOOK_HAUL, na.rm = TRUE)
median(sets_ci$NUM_HOOK_HAUL, na.rm = TRUE)

#1 station with 0 hooks hauled, 1 with 4 hook, 11 with less than 100 hooks
#hook counts in the 6000s are for whole fishing days (recorded as a single set)
```

split sets
```{r}
sets_fs_ss <- FixedSurveyData(sp=30, datadir, add.gear=T, add.LF=F, by.sex=F,correct.splitsets=F)
sets_fs_ss <- sets_fs_ss %>% filter(YEAR < 2021) %>% filter(YEAR > 2000)

sets_fs_ss_new <- sets_fs_ss %>% filter(YEAR > 2013)
```

Number of fixed stations overall
```{r}
sets_fs_fix <- sets_fs %>% filter(YEAR > 2000)

sets_fs1_early <- sets_fs1 %>% filter(YEAR %in% c(1998, 1999, 2000))

3831+566
```

Pulling halibut catch from all 3 surveys
```{r}
#fixed station
vessels_fs <- sets_fs %>%
  group_by(YEAR) %>%
  summarise(n_vessels = n_distinct(VESS_ID))

stations_fs <- sets_fs %>%
  group_by(YEAR) %>%
  summarise(n_station = n_distinct(STATION))

catch_fs <- sets_fs %>%
  group_by(YEAR) %>%
  summarise(catch = sum(EST_COMBINED_WT, na.rm = TRUE)) %>%
  mutate(catch = catch/1000)

fs_stats <- vessels_fs %>%
  left_join(stations_fs, by = "YEAR")
#fixed station summary
fs_stats <- fs_stats %>%
  left_join(catch_fs, by = "YEAR")

#stratified random
vessels_sr <- sets_sr %>%
  group_by(YEAR) %>%
  summarise(n_vessels = n_distinct(VESS_ID))

stations_sr <- sets_sr %>%
  group_by(YEAR) %>%
  summarise(n_station = n_distinct(STATION))

catch_sr <- sets_sr %>%
  group_by(YEAR) %>%
  summarise(catch = sum(EST_COMBINED_WT, na.rm = TRUE)) %>%
  mutate(catch = catch/1000)

sr_stats <- vessels_sr %>%
  left_join(stations_sr, by = "YEAR")
#statified Random summary
sr_stats <- sr_stats %>%
  left_join(catch_sr, by = "YEAR")

#commercial index
vessels_ci <- sets_ci %>%
  group_by(YEAR) %>%
  summarise(n_vessels = n_distinct(VESS_ID))

stations_ci <- sets_ci %>%
  group_by(YEAR) %>%
  summarise(n_station = n())

catch_ci <- sets_ci %>%
  group_by(YEAR) %>%
  summarise(catch = sum(EST_COMBINED_WT, na.rm = TRUE)) %>%
  mutate(catch = catch/1000)

ci_stats <- vessels_ci %>%
  left_join(stations_ci, by = "YEAR")
#commercial index summary
ci_stats <- ci_stats %>%
  left_join(catch_ci, by = "YEAR")
```

monthly distribution of stations
```{r}
sets_fs <- sets_fs %>%
  mutate(MONTH = substr(DATE_TIME3, 6,7))
sets_sr <- sets_sr %>%
  mutate(MONTH = substr(DATE_TIME3, 6,7))
sets_ci <- sets_ci %>%
  mutate(MONTH = substr(DATE_TIME3, 6,7))

fs_monthly <- sets_fs %>%
  group_by(YEAR, MONTH) %>%
  summarise(n=n())
fs_monthly <- spread(fs_monthly, MONTH, n)

sr_monthly <- sets_sr %>%
  group_by(YEAR, MONTH) %>%
  summarise(n=n())
sr_monthly <- spread(sr_monthly, MONTH, n)

ci_monthly <- sets_ci %>%
  group_by(YEAR, MONTH) %>%
  summarise(n=n())
ci_monthly <- spread(ci_monthly, MONTH, n)
##coming up in december...because bad data (basically (9998-12-31) when no data available I assume...)
```

observed by NAFO
```{r}
#commercial index separating at-sea observed vs. port-sample observed
aso_ci <- sets_ci %>%
  filter(TRIPCD_ID == 7057)
pso_ci <- sets_ci %>%
  filter(TRIPCD_ID == 7058)

aso_nafo_year <- aso_ci %>%
  group_by(YEAR, NAFO) %>%
  summarise(n_station = n())
aso_nafo_year <- spread(aso_nafo_year, NAFO, n_station)

pso_nafo_year <- pso_ci %>%
  group_by(YEAR, NAFO) %>%
  summarize(n_station = n())
pso_nafo_year <- spread(pso_nafo_year, NAFO, n_station)
```

Stratified Random sets by strata by year
```{r}
sr_strata_yr <- sets_sr %>%
  group_by(YEAR, STRATUM_ID) %>%
  summarize(n_station = n())
sr_strata_yr <- spread(sr_strata_yr, STRATUM_ID, n_station)
```

