---
title: "HS 2017 and 2018 Temperature"
output: word_document
---

```{r}
library(ggplot2)
library(geosphere)
library(dplyr)
library(marmap)
library(rgdal)
library(raster)
library(PBSmapping)
library(reshape2)
# library(SpatialHub)
library(rgeos)
library(metR)
library(tidyverse)
library(viridis)
library(scales)
```


Get the NOAA Bathy data to plot the station locations
```{r}
 survey_area <- getNOAA.bathy(lon1 = -45, lon2 = -69, lat1 = 39.5, lat2 = 48.5, resolution = 1)
```
Load in the NAFO regions (From: https://www.nafo.int/Data/GIS)
```{r}
 NAFO <- readOGR("C:/Users/denheyerc/Documents/Brendan/NAFO Divisions Shape File/Divisions.shp")
```

Crop to match the area of interest
```{r}
 NAFO_work <- crop(NAFO, y = c(-69, -48, 39.5, 48.5))
```

Make a pretty map

```{r}
tt <- autoplot(survey_area, geom=c("r", "c"), colour="grey", size=0.1) + scale_fill_etopo() + geom_contour(aes(z=z),
      breaks=c(-100, -200, -500, -1000, -2000, -4000),
      colour="grey90", size=0.01) +
        xlab("Degrees Longitude") +
        ylab("Degrees Latitude") + 
#  theme(legend.position = "none") +
#  geom_polygon(data = EEZ[EEZ$PID == 1, ], aes(x = X, y = Y), fill = NA, colour = "black", linetype = 2) + # Keyhole
#  geom_polygon(data = EEZ[EEZ$PID == 2, ], aes(x = X, y = Y), fill = NA, colour = "black", linetype = 2) + 
  geom_polygon(data = NAFO_work, aes(x = long, y = lat, group = group), fill = NA, colour = "black") +
  coord_cartesian(xlim = c(-69, -48), ylim = c(39.5, 48.5))
tt








```

Read in the 2017 temperature data

```{r}
temperature_2017 <- read.csv("C:/Users/denheyerc/Documents/Brendan/Work with collaborators/Charlotte Luo/Temperature Data/Temperature_hook_occupancy_2017.csv", header = TRUE, stringsAsFactors = FALSE)
temperature_2018 <- read.csv("C:/Users//denheyerc/Documents/Brendan/Work with collaborators/Charlotte Luo/Temperature Data/Temperature_hook_occupancy_2018.csv", header = TRUE, stringsAsFactors = FALSE)
```

Look at how many stations for 2017 dop not have associated temperature data
```{r}
length(which(is.na(temperature_2017$MED_TEMP)))
```

Plot and have  look at temperatures
```{r}
tt + geom_point(data = temperature_2017, aes(x = P1LONG, y = P1LAT, colour = MED_TEMP, size = 2), show.legend = TRUE) + scale_colour_viridis(option = "A")
```


Check how many stations don't have temperature for
```{r}
length(which(is.na(temperature_2018$MED_TEMP)))
```

Plot and have  look at temperatures
```{r}
tt + geom_point(data = temperature_2018, aes(x = P1LONG, y = P1LAT, colour = MED_TEMP, size = 2), show.legend = TRUE) + scale_colour_viridis(option = "A")
```



Get the set information for the survey stations
```{r}
require(RODBC)
ch <- odbcConnect("Bank", uid = "denheyer", pwd = "plb479a")
INF<-sqlQuery(ch,paste("SELECT YEAR, STATION, MIN(SDAYTIME) AS SDAYTIME FROM ISDB_HALIBUT.ISD_INF WHERE SETCD_ID IN (5) AND YEAR IN (2017,2018) GROUP BY YEAR, STATION")) # using group by to deal with split sets
odbcClose(ch) # head(INF)
INF$DATE<-substr(as.character(INF$SDAYTIME), 1,10)
```

Get the set information for the survey stations and write new file for Zeliang
```{r}
setwd('C://Users/denHeyerC/Documents/Students/Jiaxin/temperature')
temp2017<-read.csv(file="Temperature_hook_occupancy_2017.csv", header=T)
temp2018<-read.csv(file="Temperature_hook_occupancy_2018.csv", header=T)
dat2017<-temp2017[,c(3, 6,7,16)]
# head(dat2017)
dat2017$YEAR<-2017
dat2018<-temp2018[,c(3, 7,8,19)]
head(dat2018)
dat2018$YEAR<-2018
dat<-NULL
dat<-rbind(dat2017, dat2018)
newdat<-NULL
newdat<-merge(INF, dat, by = c('YEAR', 'STATION'))

# Comment out the write command so not to overwrite the data
#  write.csv(newdat, "HalSRTempData.csv")
```


Pull in the tempaerature data from the Zeliang
```{r}
modtemp<-read.table(file="data_obs_mod.csv", header=T)
summary(modtemp)
str(modtemp)
```


```{r}
with(modtemp, (plot(MED_TEMP, MOD_TEMP, col=YEAR, xlab="Observed Temperature", ylab="Model Temperature")))
modtemp$residtemp<-modtemp$MED_TEMP - modtemp$MOD_TEMP
hist(modtemp$residtemp, xlab="Observed - Model", main = "")  # by year
summary(modtemp$MED_TEMP)
summary(modtemp$residtemp)
```

```{r}
require(dichromat)
dat<-modtemp
dat$COLOR_VALUE<-dat$residtemp
my.colors<-colorRampPalette(c("yellow2", "goldenrod", "darkred")) #creates a function my.colors which interpolates n colors 
# my.colors<-colorRampPalette(c("green", "white", "red")) #creates a function my.colors which interpolates n colors between blue, white and red

color.df<-data.frame(COLOR_VALUE=seq(-8,8,0.01), color.name=my.colors(1601)) #generates 2001 colors from the color ramp
dat2<-merge(dat, color.df, by="COLOR_VALUE")
```


Plot and have  look at temperatures
```{r}

par(mfcol=c(2,1),mar=c(6,4,2,2))

dat3<-dat2[dat2$YEAR==2017,]
tt + geom_point(data = dat3, aes(x = P1LONG, y = P1LAT, colour = residtemp, size = 0.1), show.legend = TRUE)  +
  ggtitle("2017 Halibut SR Survey Temperature Residuals") + scale_color_gradient2(low = muted("yellow"), mid = "white",
  high = muted("red"), midpoint = 0, space = "Lab",
  na.value = "grey50", guide = "colourbar", aesthetics = "colour")

dat3<-dat2[dat2$YEAR==2018,]
tt + geom_point(data = dat3, aes(x = P1LONG, y = P1LAT, colour = residtemp, size = 0.1), show.legend = TRUE)  +
  ggtitle("2018 Halibut SR Survey Temperature Residuals") + scale_color_gradient2(low = muted("yellow"), mid = "white",
  high = muted("red"), midpoint = 0, space = "Lab",
  na.value = "grey50", guide = "colourbar", aesthetics = "colour")


```

# Pull in the temperature data from the Glory data from Zeliang.
```{r}
modtemp2<-read.table(file="data_obs_glorys.csv", header=T)
summary(modtemp2)

```



```{r}
with(modtemp2, (plot(MED_TEMP,GLORYS_TEMP, col=YEAR, xlab="Observed Temperature", ylab="Model Temperature")))
modtemp2$residtemp<-modtemp2$MED_TEMP - modtemp2$GLORYS_TEMP
hist(modtemp2$residtemp, xlab="Observed - Model", main = "")  # by year
summary(modtemp2$MED_TEMP)
summary(modtemp2$GLORYS_TEMP)
summary(modtemp2$residtemp)
```

```{r}
require(dichromat)
dat<-modtemp2
dat$COLOR_VALUE<-dat$residtemp
my.colors<-colorRampPalette(c("yellow2", "goldenrod", "darkred")) #creates a function my.colors which interpolates n colors 
# my.colors<-colorRampPalette(c("green", "white", "red")) #creates a function my.colors which interpolates n colors between blue, white and red

color.df<-data.frame(COLOR_VALUE=seq(-8,8,0.01), color.name=my.colors(1601)) #generates 2001 colors from the color ramp
dat2<-merge(dat, color.df, by="COLOR_VALUE")
```


Plot and have  look at temperature deviations.  Not there is some spatial pattern in the deviations.  Map the sets that do not have the temperature data for comparison.
```{r}

par(mfcol=c(2,1),mar=c(6,4,2,2))

dat3<-dat2[dat2$YEAR==2017,]
dat4<-dat[is.na(dat$MED_TEMP)&dat$YEAR==2017,]
tt + geom_point(data = dat3, aes(x = P1LONG, y = P1LAT, colour = residtemp, size = 0.1), show.legend = TRUE)  +
  ggtitle("2017 Halibut Survey Temperature Residuals") + scale_color_gradient2(low = muted("yellow"), mid = "white",
  high = muted("red"), midpoint = 0, space = "Lab",
  na.value = "grey50", guide = "colourbar", aesthetics = "colour")+ geom_point(data = dat4, aes(x = P1LONG, y = P1LAT))

dat3<-dat2[dat2$YEAR==2018,]
dat4<-dat[is.na(dat$MED_TEMP)&dat$YEAR==2018,]
dat5<-dat[dat$STATION=='580'&dat$YEAR==2018,]
tt + geom_point(data = dat3, aes(x = P1LONG, y = P1LAT, colour = residtemp, size = 0.1), show.legend = TRUE)  +
  ggtitle("2018 Halibut Survey Temperature Residuals") + scale_color_gradient2(low = muted("yellow"), mid = "white",
  high = muted("red"), midpoint = 0, space = "Lab",
  na.value = "grey50", guide = "colourbar", aesthetics = "colour") + geom_point(data = dat4, aes(x = P1LONG, y = P1LAT)) + geom_point(data = dat5, colour="Red", aes(x = P1LONG, y = P1LAT))

```